### 2022-10-31 Gedoen.tex

\documentclass[a4paper,10pt]{article}
\usepackage{subfiles}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{enumitem}
\usepackage{amsfonts}
\usepackage{kantlipsum,graphicx}
\usepackage{graphicx} 
\usepackage{comment}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{comment}
\usepackage{booktabs}
\usepackage{xfrac}
\usepackage{floatrow}
\usepackage{multicol}
\usepackage{float}
\usepackage{tikz}
\definecolor{light-gray}{gray}{0.8}
\usepackage{graphicx}
\setlength\parindent{0pt}
\graphicspath{{images/2dwave/}{images/2dBeam/}}
\usepackage{xcolor}
\usepackage{pgfplots}

\begin{document}
\section*{Gedoen}

\subsection{Chapter 1}
\begin{itemize}
	\item Het 1.4.21 reg gemaak. Moet nog deur gaan en seker maak ek het dit ander plekke reg.
\end{itemize}

\subsection{Chapter 2}
\begin{itemize}
	\item Het verbeteringe aangebring.
\end{itemize}

\subsection{Appendix}
\begin{itemize}
	\item $L_2$ in gesit.
\end{itemize}

\end{document}