### 2022-10-31 References.tex

\documentclass[main.tex]{subfiles}
\begin{document}
\begin{thebibliography}{AAA}
\bibitem[AL97]{An3} S S Antman and M Lanza de Christoforis, {\it Peculiar
instabilities due to the clamping of shearable rods}, Int. J.
Non-linear Mechanics $\mathbf{23}$ (31-54) 1997.

\bibitem[An76]{An1} S S Antman, {\it Ordinary differential eqautions of
non-linear elasticity~I: Foundations of the theories of non-linear
elastic rods and shells}, Archive for Rational Mechanics and
Analysis, $\mathbf{61}$, (307-351), 1976.

\bibitem[An96]{An2} S S Antman, {\it Dynamical Problems for Geometrically
Exact Theories of Nonlinearly Viscoelastic rods}, J. Nonlinear Sci.,
$\mathbf{6}$, 1-18, 1996.

\bibitem[BOC20]{Bochicchio2020} Bochiccho, I., Campo, M., Fernandez, J.R., Grazia Naso, M.,\ \textit{Analysis of a Thermoelastic Timoshenko Beam Model}, Acta Mechanica., (2020).

\bibitem[Bh09]{Bh} A Bhaskar, {\it Elastic waves in Timoshenko beams: the
`lost and found' of an eigenmode}, Proc. R. Soc. A $\mathbf{465}$
(239-255) 2009.

\bibitem[Co66]{Co} G R Cowper, {\it The shear coefficient in Timoshenko's beam
theory}, Journal of Applied Mechanics $\mathbf{33}$ (335-340) 1966.

\bibitem[CVV18]{VanRensburg18} Civin, D., van Rensburg, N.F.J., van der Merwe A.J.,\ \textit{Using energy methods to compare linear vibration models}, Applied Mathematics and Computation, \textbf{321}, (2018), 602-613.

\bibitem[Fung65]{Fung1965} Fung, Y C.,\ \textit{Foundations of Solid Mechanics}, Prentice-Hall Inc., Englewood Cliffs, (1965).

\bibitem[LC81]{LL} M Levinson and D W Cooke, {\it On the two frequency spectra of Timoshenko Beams}, Journal of
Sound and Vibration, $\mathbf{84(3)}$ (319-326), 1982.

\bibitem[LL91]{LL} J E Lagnese and G Leugering, {\it Uniform stabilization of
a nonliner beam by nonliner boundary feedback}, Journal of
Differential Equations, $\mathbf{91}$ (355-388), 1991.

\bibitem[LVV09]{LVV} A Labuschagne, N F J van Rensburg and A J van der Merwe,
{\it Comparison of linear  beam theories}, Mat. Comput. Model.,
$\mathbf{49}$,(20-30), 2009.

\bibitem[McK99]{McKenna97} McKenna, P.J.,\ \textit{Large Torsional Oscillations in Suspension Bridges Revisited: Fixing an Old Approximation}, American Mathematical Monthly, \textbf{106}(1), (1999), 1-18.

\bibitem[Rud53]{McKenna97} Walter, R.,\ \textit{Principles of Mathematical Analysis}, McGraw-Hill Book Company, Inc., New York-Toronto-London, (1953).

\bibitem[Sadd05]{Sadd} M H Sadd, {\it Elasticity. Theory, Applications and Numerics}, Academic Press, Burlington, (123-138), 2005

\bibitem[SR79]{SR} M Sapir and L Reiss, {\it Dynamic Buckling of a
Nonlinear Timoshenko Beam}, Siam J. Appl. Math., Vol
$\mathbf{37}$(2), (290-301), 1979.s

\bibitem[SV86]{SP6} J C Simo and L Vu-Quoc, {\it On the dynamics of
flexible beams under large overall motions - the plane case: Part
I}, Journal of Applied Mechanics {\bf 53}, (849-854), 1986.

\bibitem[SV87]{SP7} J C Simo and L Vu-Quoc, {\it The role of
nonlinear theories in transient dynamic analysis of flexible
structures}, Journal of Sound and Vibration {\bf 119}(3), 487-508,
1987.

\bibitem[SP06]{SP} N G Stephen and S Puchegger, {\it On the valid frequency
range of Timoshenko beam theory} J. Sound Vib. $\mathbf{297}$
(1082-1087) 2006.

\bibitem[Ti21]{Ti1} S Timoshenko *******

\bibitem[Ti37]{Ti2} S Timoshenko, {\it Vibration problems in engineering}
2nd Ed, 1937.

\bibitem[VV06]{VanRensburg06} van Rensburg, N.F.J., van der Merwe A.J.,\ \textit{Natural frequencies and modes of a Timoshenko beam}, Wave Motion, \textbf{44}, (2006), 58-69.

\bibitem[WU06]{Wu2006} Wu, Shen R.,\ \textit{Lumped Mass matrix in explicit finite element method for transient dynamics of elasticity}, Computer methods in applied mechanics and engineering, \textbf{195}, (2006), 5983-5994.

\end{thebibliography}
\end{document}