### 2022-10-31 files/Chapter6/2D_FEM.tex

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Introduction}
In this chapter, the Finite Element Method (FEM) is applied to the two- and three-dimensional cantilever beams as well as a Reissner-Mindlin cantilever plate. The aim of this chapter is to obtain a numerical formula to calculate the eigenvalues and eigenvectors for these models. This chapter serves as preparation for Chapter 6.\\

The outline of the chapter is as follows.
\begin{itemize}
	\item[] Section \ref{sec:FEM:2D} is the FEM for a two-dimensional cantilever body.
	\item[] Section \ref{sec:FEM:3D} is the FEM for a three-dimensional cantilever body. 
	\item[] Section \ref{sec:FEM:Plate} is the FEM for A Reissner-Mindlin cantilever plate model. 
\end{itemize}

\section{Two-dimensional cantilever body} \label{sec:FEM:2D}
Consider a two-dimensional cantilever elastic body, with a rectangular cross-section.

\subsubsection{Reference configuration for rectangular cross-section}
Let $\{e_1,e_2\}$ be a right-handed orthonormal basis for $R^2$. Denote the elastic body by $\Omega \in R_2$, with $(0,0)$ as the reference point. For a rectangular cross-section, $\Omega$ can be described as
\begin{eqnarray*}
	\Omega = \left \{ x \in E_2 \ | \ 0 \leq x_1 \leq 1, \ -\frac{h}{2} \leq x_2 \leq \frac{h}{2} \right \},
\end{eqnarray*} where $\partial \Omega$ denotes the boundary of $\Omega$. The boundary $\partial \Omega$ can be divided into the four distinct lines as follows:

\noindent\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\Sigma_0:& \quad x_1 &= 0\\
		\Sigma_1:& \quad x_1 &= 1
	\end{eqnarray*}
\end{minipage}%
\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\Gamma_1:& \quad x_2 &= -{h}/{2}\\
		\Gamma_2:& \quad x_2 &= {h}/{2}
	\end{eqnarray*}
\end{minipage}

\subsubsection{Cantilever elastic body}
Consider a two-dimensional elastic body clamped rigidly to a surface at $x_1 = 0$. The body is free-hanging on the other boundaries This is referred to as Problem 2D-1 in section \ref{ssec:2D_Model:ModelProblem}.
\FloatBarrier
\begin{figure}[h!]
	\centering
	\begin{tikzpicture}

		\draw[line width = 0.4mm] (0,0.5) -- (7,0.5);
		\draw[line width = 0.4mm] (0,-0.5) -- (7,-0.5);
		\draw[line width = 0.4mm] (7,-0.5) -- (7,0.5);
		\draw[line width = 0.4mm] (0,-0.5) -- (0,0.5);

		\draw[line width = 0.1mm] (0,1.5) -- (0,-1.5);

		\draw[line width = 0.1mm] (0,1.5) -- (-0.2,1.3);
		\draw[line width = 0.1mm] (0,1.3) -- (-0.2,1.1);
		\draw[line width = 0.1mm] (0,1.1) -- (-0.2,0.9);
		\draw[line width = 0.1mm] (0,0.9) -- (-0.2,0.7);
		\draw[line width = 0.1mm] (0,0.7) -- (-0.2,0.5);
		\draw[line width = 0.1mm] (0,0.5) -- (-0.2,0.3);
		\draw[line width = 0.1mm] (0,0.3) -- (-0.2,0.1);
		\draw[line width = 0.1mm] (0,0.1) -- (-0.2,-0.1);
		\draw[line width = 0.1mm] (0,-0.1) -- (-0.2,-0.3);
		\draw[line width = 0.1mm] (0,-0.3) -- (-0.2,-0.5);
		\draw[line width = 0.1mm] (0,-0.5) -- (-0.2,-0.7);
		\draw[line width = 0.1mm] (0,-0.7) -- (-0.2,-0.9);
		\draw[line width = 0.1mm] (0,-0.9) -- (-0.2,-1.1);
		\draw[line width = 0.1mm] (0,-1.1) -- (-0.2,-1.3);
		\draw[line width = 0.1mm] (0,-1.3) -- (-0.2,-1.5);

		\node at (7.3,0) {$h$};
		\draw[-stealth] (7.3,0.2) -- (7.3,0.5);
		\draw (7.2,0.5) -- (7.4,0.5);
		\draw[-stealth] (7.3,-0.2) -- (7.3,-0.5);
		\draw (7.2,-0.5) -- (7.4,-0.5);

		\node at (3.5,-0.8) {$1$};
		\draw[stealth-] (0,-0.8) -- (3.3,-0.8);
		\draw[stealth-] (7,-0.8) -- (3.7,-0.8);
		\draw (7,-0.9) -- (7,-0.7);

		\draw[line width = 0.1mm,->] (9,1) -- (10.4,1);
		\draw[line width = 0.1mm,->] (9,1) -- (9,2.1);
		\node at (9.7,0.8) {$e_1$};
		\node at (8.8,1.5) {$e_2$};

		\node at (-0.6,0) {$(0,0)$};
		\node at (0,0)[circle,fill,inner sep=1.5pt]{};

	\end{tikzpicture}
	\caption{A cantilever two-dimensional elastic body.}
\end{figure} 
\FloatBarrier

In section \ref{ssec:2D_Model:VariationalForm} the variational problem for the two dimensional cantilever model is defined by Problem 2D-1V. For convenience, the relevant results from section \ref{ssec:2D_Model:VariationalForm} are repeated.
\subsubsection{Problem 2D-1V}\label{sssec:2D_Model:Problem2D1VChp5}
Find a function $u$ such that for all $t>0$, $u \in T(\Omega)$ and 
\begin{align}
	c(u,\phi) = -b(u,\phi) + (Q,\phi) \label{eq:2D_Model:Problem2D1VEqChp5}
\end{align}
for all $\phi \in T(\Omega)$.\\

With the test function space 
\begin{eqnarray*}
	T(\Omega) & = & \left\{ \phi \in C^1(\bar{\Omega}) \ | \ \phi = 0 \ \textrm{ on } \ \Gamma \right\}.
\end{eqnarray*}
The bilinear forms and integral function defined by
\begin{eqnarray}
	b(u,\phi) & = & \int_\Omega~c_1\textrm{Tr}({\cal E}\Phi)+ c_2\textrm{Tr}({\cal E})
	\textrm{Tr}(\Phi) ~dA, \label{eq:2D_Model:BilinearChp5}\\
	c(\partial_t^2 u,\phi) & = & \int_\Omega~ (\partial^2_t u) \cdot \phi~dA, \label{eq:2D_Model:Bilinear_cChp5}\\
	(f,g) &=& \int_{\Omega} f\cdot g \ dA, \label{eq:2D_Model:Bilinear_intChp5}
\end{eqnarray}
with $\displaystyle c_1 = \frac{1}{\gamma(1+\nu)}$ and $\displaystyle c_2 = \frac{\nu}{\gamma(1-\nu^2)}$.\\

Using the definition of the reference configuration, the constitutive equations and the bilinear form $b$ can be rewritten as follows.

\subsubsection{Constitutive equations:}
\begin{eqnarray}
	\sigma_{11} & = & \frac{1}{\gamma(1-\nu^2)}(\partial_1 u_1 + \nu \partial_2 u_2) \label{CE1} \\
	\sigma_{22} & = & \frac{1}{\gamma(1-\nu^2)}(\partial_2 u_2 + \nu \partial_1 u_1) \label{CE2} \\
	\sigma_{12} & = & \frac{1}{2\gamma(1+\nu)}(\partial_1 u_2 + \partial_2 u_1) \label{CE3}
\end{eqnarray}

\subsubsection{Bilinear form:}
\begin{align}
	b(u,\phi) & = & \frac{1}{\gamma(1-\nu^2)}\int_{\Omega} (\partial_1 u_1 \partial_1 \phi_1 + \partial_2 u_2 \partial _2 \phi_2 + \nu\partial_1 u_1 \partial_2\phi_2 + \nu \partial_2 u_2 \partial_1 \phi_1 ) \ dA \nonumber\\
	& + & \frac{1}{2\gamma(1+\nu)}\int_{\Omega} (\partial_1 u_2 \partial_1 \phi_2 + \partial_1 u_2 \partial_2 \phi_1 + \partial_2 u_1 \partial_1\phi_2 + \partial_2 u_1 \partial_2\phi_1) \ dA.
\end{align}

This form of the bilinear form $b$ is more useful than the general form in \eqref{eq:2D_Model:BilinearChp5} to apply the Finite Element Method.

\subsection{Weak variational form}
Define the inertia space $V$ as the closure of $T(\Omega)$ in $H := H^1(0,1)\times H^1(0,1)$. Denote $X = L^2(0,1)\times L^2(0,1)$. The inertia space is $W  = X$ with norm $||\cdot||_W = \sqrt{c(\cdot,\cdot)}$.

\subsubsection{Problem 2D-1WV}
Find a function $u$ such that $\forall \ t > 0$, $u \in V$ with $\partial_t^2 u \in W$ and
\begin{eqnarray*}
	c(u,v) + b(u,v) & = & (Q,v)
\end{eqnarray*}
for all $v\in V$.

\subsection{Galerkin approximation}\label{2d_FEM_G}
Divide the reference configuration $\Omega$ into a rectangular grid of elements, such that there are $n = n_1\times n_2$ nodes.\\

Define a set of $n$-dimensional linear independent basis functions. For the two dimensional model, the basis functions can be defined by the set $$\left\{\langle\phi_1, 0\rangle, \langle\phi_2, 0\rangle,...,\langle\phi_{n}, 0 \rangle,\langle 0,\phi_1\rangle,\langle 0 ,\phi_2\rangle,...,\langle 0,\phi_{n}\rangle \right\}.$$ For this dissertation, we use bi-cubic functions $\phi_i$, however simpler bi-linear functions can also be used. For easier notation, denote each of these basis functions by $\delta_i$.\\

The functions in this set that satisfies the conditions of the test function space $T(\Omega)$ are called the admissible basis functions. The admissible basis functions for this model can be renumbered and expressed as the set $\left \{\delta_1, \delta_2,..., \delta_k \right\}$ for some $k \leq 2n$.\\

Define the space
\begin{eqnarray*}
S^h & = & \textrm{span}\left(\left\{\delta_i \ | \ i = 1,2,...,k \right\} \right)
\end{eqnarray*}

For each function $u^h \in S^h$, $u^h$ can be expressed as
\begin{eqnarray*}
	u^h = \sum_{i = 1}^{k} u_i(t) \delta_{i}(x)
\end{eqnarray*}

Problem 2D-1V can now be rewritten into a Galerkin Approximation, denoted by Problem 2D-1G.

\subsubsection{Problem 2D-1G}
Find a function $u^h$ such that for all $t>0$, $u^h \in S^h$ and
\begin{eqnarray*}
	(u^h, \phi_i) & = & -b(u^h,\phi_i) + (Q^I \cdot \phi_i)
\end{eqnarray*} for $i = 1,2,...,k$.\\

\textbf{Remarks:}
\begin{itemize}
	\item[] $Q^I$ is scalar vector that is the result of interpolating the function $Q$ over the rectangular grid. i.e. $Q^I_{i,j} = Q(x_i,x_j)$ for $i = 1,2,...,n_1$ and $j = 1,2,...,n_2$.
\end{itemize}

\subsection{System of differential equations}\label{ssec:2DFEM:DE}

Consider the following standard Finite Element Method matrices
\subsubsection{FEM matrices}
\noindent\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\mathbf{M}_{j,i} & = & \int_{\Omega} \phi_i \phi_j ~dA \\
		\mathbf{{K}_{11}}_{j,i} & = & \int_{\Omega} \partial_1\phi_i \partial_1\phi_j~dA\\
		\mathbf{{K}_{22}}_{j,i} & = & \int_{\Omega} \partial_2\phi_i \partial_2\phi_j~dA
	\end{eqnarray*}
\end{minipage}%
\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\mathbf{{K}_{12}}_{j,i} & = & \int_{\Omega} \partial_2\phi_i \partial_1\phi_j~dA\\
		\mathbf{{K}_{21}}_{j,i} & = & \int_{\Omega} \partial_1\phi_i \partial_2\phi_j~dA
	\end{eqnarray*}
\end{minipage}
for $i,j = 1,2,...,k$.\\

And 
\begin{eqnarray*}
	{M_{F}}_{j,i} = \int_{\Omega}  \phi_i \phi_j~dA
\end{eqnarray*} for $i = 1,2,...,k$ and for $j =1,2,...,2n$.\\

Define the following matrices:
\begin{eqnarray*}
	K_1 & = & \frac{1}{\gamma(1-\nu^2)} \mathbf{K_{11}} + \frac{1}{2\gamma(1+\nu)}\mathbf{K_{22}} \label{eq:2DFEM:K1} \\
	K_2 & = & \frac{\nu}{\gamma(1-\nu^2)} \mathbf{K_{21}} + \frac{1}{2\gamma(1+\nu)}\mathbf{K_{12}}\label{eq:2DFEM:K2}\\
	K_3 & = & \frac{\nu}{\gamma(1-\nu^2)} \mathbf{K_{12}} + \frac{1}{2\gamma(1+\nu)}\mathbf{K_{21}}\label{eq:2DFEM:K3}\\
	K_4 & = & \frac{1}{\gamma(1-\nu^2)} \mathbf{K_{22}} + \frac{1}{2\gamma(1+\nu)}\mathbf{K_{11}}\label{eq:2DFEM:K4}
\end{eqnarray*}

Using the standard FEM matrices and matrices $K_1$-$K_4$, the following concatenated matrices are defined.
\begin{eqnarray}
	\begin{aligned}
		K = 
		\begin{bmatrix}
			K1 & K2\\
			K3 & K4
		\end{bmatrix}
	\end{aligned}
	\ \ \ \ \ \ \ \ \
	\begin{aligned}
		M_f = 
		\begin{bmatrix}
			M_{F} & O_{F}\\
			O_{F} & M_{F}
		\end{bmatrix}
	\end{aligned}\label{eq:2DFEM:K+M}
\end{eqnarray}

\begin{eqnarray}
	M = 
	\begin{bmatrix}
		\mathbf{M} & O \\
		O & \mathbf{M}
	\end{bmatrix}\label{eq:2DFEM:M}
\end{eqnarray}

The matrices ${O}$ and ${O_f}$ are the zero matrices of the same size as $\mathbf{M}$ and ${M_f}$ respectively.\\

Using \eqref{eq:2DFEM:K+M} and \eqref{eq:2DFEM:M}, Problem 2D-1G is rewritten as a system of ordinary differential equations. This system is referred to as Problem 2D-1ODE.

\subsubsection{Problem 2D-1ODE}
Find a function $\bar{u} \in S^h$ such that
\begin{eqnarray}
	M\ddot{\bar{u}} & = & K\bar{u} + M_{f}Q^I.
\end{eqnarray} With $\bar{u}$ in the form $\bar{u} = \langle u, \partial_1 u, \partial_2 u, \partial_{12} u \rangle$.

\subsubsection{Remark}
This form of $\bar{u}$ is determined by the Hermite cubic basis functions.

\subsection{Eigenvalue problem}\label{2dFEM_EP}
For the eigenvalue problem, assume that $M_{f}Q^I = 0$ so that 
\begin{eqnarray}
		M\ddot{\bar{u}} & = & K\bar{u}.\label{eq:2DFEM:M2}
\end{eqnarray}
It is known that a system of ordinary differential equations have a general solution of the form $e^{rt}$. Suppose that $\bar{w} = e^{\lambda t} \bar{u}$ is a solution for \eqref{eq:2DFEM:M2}. In this solution, $\lambda$ is an eigenvalue and $\bar{u}$ a corresponding eigenfunction. Substitution into \eqref{eq:2DFEM:M2} results in
\begin{eqnarray*}
	M\lambda e^{\lambda t}\bar{u} & = & Ke^{\lambda t}\bar{u}.
\end{eqnarray*}
Since $e^{\lambda t} > 0$ for all values of $\lambda t$, we can cancel it from both sides of the equation and formulate the eigenvalue problem Problem 2D-1E.

\subsubsection{Problem 2D-1E}
Find a real number $\lambda$ and a function $\bar{u} \in S^h$ such that
\begin{eqnarray}
	M\lambda{\bar{u}} & = & K\bar{u}.
\end{eqnarray}

Now the eigenvalue problem can be solve using a finite difference method.

\end{document}

Divide the reference configuration $\Omega$ into $n_1\times n_2$ nodes in a rectangular grid.\\

Consider a set of 
For this dissertation, we consider bi-cubic basis functions, however bi-linear basis functions can also be used.

Consider a $n$-dimensional set of linearly independent two-dimensional scalar valued functions, denoted $\delta_i$. In this dissertation, these functions are chosen as bi-cubic functions, however simpler bi-linear functions can also be used.\\

The functions in the set that satisfies the conditions of the test function space $T(\Omega)$ are called the admissible basis functions. Number these admissible basis functions $\delta_1, \delta_2,...,\delta_k$ for some $k \leq n$.

Let $S^h$ be a $2k$-dimensional subset of $T(\Omega)$ such that
\begin{eqnarray*}
	S^h & = & \textrm{span}\left \{[\delta_1 \ 0], [\delta_2 \ 0],...,[\delta_k \ 0],[0 \ \delta_1],[0 \ \delta_2],...,[0 \ \delta_k] \right\}.
\end{eqnarray*}

Let $S^h$ be a $2k$-dimensional subset of $T(\Omega)$ such that
\begin{eqnarray*}
	S^h & = & \textrm{span}\left \{[\delta_1 \ 0], [\delta_2 \ 0],...,[\delta_k \ 0],[0 \ \delta_1],[0 \ \delta_2],...,[0 \ \delta_k] \right\}.
\end{eqnarray*}

he body $\Omega$ is divided into an $n_1\times n_2$ rectangular grid. Let $x_i^* = \langle x_1, x_2 \rangle \in \Omega$ where $\delta_{i}(x_1,x_2) = 1$. Define $\bar{x} = \langle x_1^*, x_2^*,...,x_k^* \rangle$. For each solution $u$ of Problem 2D-1V let $u^h$ denote the projection of $u$ into the finite dimensional space $S^h$. Each function $u^h=\langle u_1^h,u_2^h\rangle$ in S$^h$ can be expressed as
\begin{eqnarray*}
	u_1^h(\bar{x},t) = \sum_{i = 1}^{k} u_{1}(x_i^*,t) \delta_{i}(\bar{x}) \ \ \textrm{ and } \ u_2^h(\bar{x},t) = \sum_{i = 1}^{k} u_{2}(x_i^*,t) \delta_{i}(\bar{x}).
\end{eqnarray*}