### 2022-10-31 files/Chapter7/Beam-Main_Rig_2.tex

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Validity of the Timoshenko beam model}
This section is a literature study of the article \cite{LVV09}. The goal of the authors is to show that for practical applications, the solution of the Timoshenko beam model is close to the solution of a two-dimensional beam model. This is referred to as the validity of the Timoshenko beam model in this dissertation.\\

For practical applications, it would be preferred to compare the Timoshenko beam model to a more realistic three-dimensional model. But the authors of \cite{LVV09} explain that this will introduce complexities. \textcolor{red}{What complexities?} To mitigate some of the complexities, the two-dimensional model is used as an intermediate step.\\

Where \cite{LVV09} differs from most other research, is that the boundary conditions are also taken into consideration. The authors of \cite{LVV09} use a cantilever beam configuration for their investigation.\\

To compare the two models, the authors compare the natural frequencies of the models. From modal analysis, it is known that a solution of a second order hyperbolic type problem is a linear combination of the eigenvalues and eigenfunctions of the model. Therefore it is sufficient to compare the eigenvalues and eigenfunctions to determine the validity of the different dimensional models.\\

\subsection{The models}
Consider the cantilever Timoshenko beam model problem Problem T-2 from section \ref{ssec:1D_Model:ModelProblems} and the cantilever two-dimensional beam model problem Problem 2D-1 from section \ref{ssec:2D_Model:ModelProblem}.\\

Figure 6.1 shows the two beams side-by-side for a visual guide.

\begin{figure}[h!]\label{fig:compare:1D+2D}
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\caption{Side by side comparison of the beams.}
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\begin{center}
					\begin{tikzpicture}
					\draw[line width = 0.4mm] (0,0) -- (6,0);

					\node at (2.7,-0.25) {$\ell = 1$};

					\draw[line width = 0.1mm] (0,-1.5) -- (0,1.5);
					\draw[line width = 0.1mm] (0,1.5) -- (-0.2,1.4);
					\draw[line width = 0.1mm] (0,1.3) -- (-0.2,1.2);
					\draw[line width = 0.1mm] (0,1.1) -- (-0.2,1);
					\draw[line width = 0.1mm] (0,0.9) -- (-0.2,0.8);
					\draw[line width = 0.1mm] (0,0.7) -- (-0.2,0.6);
					\draw[line width = 0.1mm] (0,0.5) -- (-0.2,0.4);
					\draw[line width = 0.1mm] (0,0.3) -- (-0.2,0.2);
					\draw[line width = 0.1mm] (0,0.1) -- (-0.2,0);

					\draw[line width = 0.1mm] (0,-0.1) -- (-0.2,-0.2);
					\draw[line width = 0.1mm] (0,-0.3) -- (-0.2,-0.4);
					\draw[line width = 0.1mm] (0,-0.5) -- (-0.2,-0.6);
					\draw[line width = 0.1mm] (0,-0.7) -- (-0.2,-0.8);
					\draw[line width = 0.1mm] (0,-0.9) -- (-0.2,-1);
					\draw[line width = 0.1mm] (0,-1.1) -- (-0.2,-1.2);
					\draw[line width = 0.1mm] (0,-1.3) -- (-0.2,-1.4);
					\draw[line width = 0.1mm] (0,-1.5) -- (-0.2,-1.6);

					%\node at (3.2,-0.2) {$\ell = 1$};
					\end{tikzpicture}
				\end{center}
				\subcaption{Timoshenko Cantilever Beam}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\begin{center}
					\begin{tikzpicture}
						\draw[line width = 0.4mm] (0,0.2) -- (6,0.2);
						\draw[line width = 0.4mm] (0,-0.2) -- (6,-0.2);
						\draw[line width = 0.4mm] (6,-0.2) -- (6,0.2);
						%\draw[line width = 0.4mm] (0,-0.2) -- (0,0.2);

						%\draw[line width = 0.1mm,->] (6,2) -- (7.4,2);
						%\draw[line width = 0.1mm,->] (6,2) -- (6,0.9);
						%\node at (6.7,1.8) {$e_1$};
						%\node at (5.8,1.4) {$e_2$};

						\node at (6.2,0) {$h$};
						\node at (2.7,-0.45) {$\ell = 1$};

						\draw[line width = 0.1mm] (0,-1.5) -- (0,1.5);
						\draw[line width = 0.1mm] (0,1.5) -- (-0.2,1.4);
						\draw[line width = 0.1mm] (0,1.3) -- (-0.2,1.2);
						\draw[line width = 0.1mm] (0,1.1) -- (-0.2,1);
						\draw[line width = 0.1mm] (0,0.9) -- (-0.2,0.8);
						\draw[line width = 0.1mm] (0,0.7) -- (-0.2,0.6);
						\draw[line width = 0.1mm] (0,0.5) -- (-0.2,0.4);
						\draw[line width = 0.1mm] (0,0.3) -- (-0.2,0.2);
						\draw[line width = 0.1mm] (0,0.1) -- (-0.2,0);

						\draw[line width = 0.1mm] (0,-0.1) -- (-0.2,-0.2);
						\draw[line width = 0.1mm] (0,-0.3) -- (-0.2,-0.4);
						\draw[line width = 0.1mm] (0,-0.5) -- (-0.2,-0.6);
						\draw[line width = 0.1mm] (0,-0.7) -- (-0.2,-0.8);
						\draw[line width = 0.1mm] (0,-0.9) -- (-0.2,-1);
						\draw[line width = 0.1mm] (0,-1.1) -- (-0.2,-1.2);
						\draw[line width = 0.1mm] (0,-1.3) -- (-0.2,-1.4);
						\draw[line width = 0.1mm] (0,-1.5) -- (-0.2,-1.6);

						%\node at (3.2,-0.3) {$\ell = 1$};

					\end{tikzpicture}
				\end{center}
				\subcaption{Two-Dimensional Cantilever Beam}
			\end{minipage}
		}
	}
\end{figure}

The parameter $h$ is the height of the two-dimensional beam. Recall the dimensionless constant $\displaystyle \alpha = \frac{A \ell^2}{I}$ from section \ref{sssec:1D_Model:DimensionlessForm}. For the Timoshenko beam model with a rectangular cross-section, the area moment of inertial is 
\begin{eqnarray*}
	I = \frac{h^3b}{12}.
\end{eqnarray*}

After substitution, the following relation ship between $\alpha$ and $h$ is found
\begin{eqnarray*}
	h = \sqrt{\frac{12}{\alpha}}.
\end{eqnarray*}
The parameter $\alpha$ can be chosen in such a way that the beams have equal dimensions.

\subsection{Accuracy of the numerical eigenvalues}
For the eigenvalues and eigenfunctions of the Timoshenko beam, the authors of \cite{LVV09} reference an another article \cite{VV06}. In this article \cite{VV06}, a method is described to obtain the eigenvalues and eigenfunctions of a (cantilever) Timoshenko beam. This method also allows for the eigenvalues to be obtained to a desired level of accuracy (using a numerical method such as interval division). In this dissertation, \cite{VV06} is discussed in section \ref{sec:Timo:EigenvalueProblem}.\\

For the two-dimensional beam, the Finite Element Method is used to calculate the eigenvalues and eigenfunctions. In \cite{LVV09} the authors used piecewise bi-cubic basis functions and a grid size large enough to ensure that the eigenvalues are accurate to 3 significant digits.\\

\begin{figure}[H]
	\begin{center}
		\scalebox{1.4}{
			\includegraphics[scale=0.4]{Convergence.png}
			\caption{Maximum relative error of the first 10 eigenvalues of the two-dimensional beam.}
			\label{fig:conv}
		}
	\end{center}	
\end{figure}

Figure \ref{fig:conv} show the maximum relative error of the first 10 eigenvalues of the two-dimensional beam. In \cite{LVV09}, the authors use a grid size of $40 \times 10$ with $h= 1/10$ (or $\alpha = 1200$) and $h = 1/5$ (or $\alpha = 300$).\\

 Figure \ref{fig:conv} show that this can be improved significantly. For each case of $h$, the grid size is calculated as $n \times \max\left\{\lceil\frac{n}{h}\rceil,10\right\}$. This ensures that the nodes are distributed somewhat evenly relative to the length-width ratio. For this chapter, the results of the Timoshenko beam and two-dimensional beam are given accurate up to 5 significant digits.

\subsection{Comparing the shape of the eigenfunctions}
Before the eigenvalues of the beams can be compared, a method is required to be able to match the eigenvalues. In \cite{LVV09} the authors compare the modal shapes of the Timoshenko beam model to the modal shapes of the two-dimensional model.\\

Similar modal shapes implies that the eigenvalues can be matched up and compared. Eigenvalues relating to the shapes of the Timoshenko beams modes are called beam-type eigenvalues by \cite{LLV09}.

\subsubsection{Shapes relating to beam-type eigenvalues}
The following are examples of beam-type mode shapes for the displacement $w$ for the cantilever two-dimensional beam and cantilever Timoshenko beam. The figures are presented side-by-side.

\begin{figure}[h!]
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{Beam1.jpg}
				\subcaption{2D Beam Type - $\lambda_6 = 21.911$}
				\label{fig:minipage2}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{1DBeam2.png}
				\subcaption{Timoshenko - $\lambda_5 = 21.794$}
				\label{fig:minipage1}
			\end{minipage}
	}}
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{Beam2.jpg}
				\subcaption{2D Beam Type - $\lambda_7 = 45.711$}
				\label{fig:minipage2}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{1DBeam1.png}
				\subcaption{Timoshenko - $\lambda_6 = 45.390$}
				\label{fig:minipage1}
			\end{minipage}
			\caption{Modal shapes of the displacement $w$ for the beam-type 2D body and the Timoshenko beam with $\alpha = 4800$ ($h = 1/20$).}
		}
	}
\end{figure}
\FloatBarrier

\subsubsection{Shapes relating to non-beam type eigenvalues}
The following are examples of non beam-type mode shapes for the displacement $w$ of the cantilever two-dimensional beam.
\begin{figure}[h!]
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{NotBeam1.jpg}
				\subcaption{Non-Beam Type - $\lambda_4 = 7.7077$}
				\label{fig:minipage2}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{NotBeam2.jpg}
				\subcaption{Non-Beam Type - $\lambda_8 = 69.344$}
				\label{fig:minipage1}
			\end{minipage}
			\caption{Modal shapes of the displacement $w$ for the non-beam type 2D body with $h = 1/20$.}
		}
	}
\end{figure}
\FloatBarrier

Using this method all the relevant eigenvalues in this section were matched. The non-beam type eigenvalues are highlighted in gray in the tables.

\subsubsection{Shapes of the cross-section}
The Timoshenko beam theory differs from other one-dimensional beam theories by also including shear. It is assumed that the cross-sections need not remain perpendicular to the neutral axis of the beam. The cross-sections remain straight.\\

For the two-dimensional body, the shape of the cross-section of beam type eigenfunctions can deform into an a s-shape. Figure \ref{fig:s} shows an exaggerated representation of the s-shaped deformation. This is also given and explained in \cite{LVV09}.
\FloatBarrier
\begin{figure}[h!]
	\centering
	\begin{tikzpicture}[scale=1.2]

	\draw[line width = 0.4mm] (0,1) -- (7,1);
	\draw[line width = 0.4mm] (0,-1) -- (7,-1);
	\draw[line width = 0.4mm] (7,-1) -- (7,1);
	\draw[line width = 0.4mm] (0,-1) -- (0,1);

	\draw[line width = 0.1mm,dashed] (3.5,1) -- (3.5,-1);

	\draw[thick] plot [smooth,tension=0.9] coordinates{(3.3,1) (3.7,0.5) (3.5,0) (3.3,-0.5) (3.7,-1)};

	\draw[line width = 0.2mm,red] (3.3,1) -- (3.7,-1);

\end{tikzpicture}
	\caption{S-shape deformation and average rotation.}\label{fig:s}
\end{figure}
The red line in figure \ref{fig:s} show how the average rotation of the cross-section is calculated in \cite{LVV09}.
\FloatBarrier

\subsubsection{Direct comparison of mode shapes}
Figure \ref{fig:w} shows the displacement $w$ of the cantilever Timoshenko beam and the cantilever two-dimensional beam directly compared.

\begin{figure}[ht!]
	\scalebox{.8}{
		\makebox[\textwidth]{
			\centering
			\centering
			\includegraphics[scale=0.4]{1Dv2D.png}
			\caption{Comparison of the displacement $w$ mode shape corresponding to $\lambda_{10}$ for the 2D beam and $\lambda_8$ for the Timoshenko beam with $\alpha = 4800$ ($h=1/20$)}
			\label{fig:w}}}
\end{figure}
\FloatBarrier

Similarly Figure \ref{fig:phi} shows the direct comparison of the angle of the cross-section of the cantilever Timoshenko beam and the two-dimensional beam. The average rotation of the cross-section of the two-dimensional beam's mode shape is calculated as shown in figure \ref{fig:s}.

\FloatBarrier
\begin{figure}[ht!]
	\scalebox{.8}{
		\makebox[\textwidth]{
			\centering
			\centering
			\includegraphics[scale=0.4]{1Dv2Dphi.png}
			\caption{Comparison of the angle $\phi$ (best fit for 2D Beam) mode shape corresponding to $\lambda_{10}$ for the 2D beam and $\lambda_8$ for the Timoshenko beam with $\alpha = 4800$ ($h=1/20$)}
			\label{fig:phi}}}
\end{figure}
\FloatBarrier

Figures \ref{fig:w} and \ref{fig:phi} show that for $\alpha = 4800$, the mode shapes are very close, even for the eigenvalues 8 (for Timoshenko beam) and 10 (for the two-dimensional beam). The authors of \cite{LVV09} obtained similar results for $\alpha = 300$ for the eigenvalues 6 (for Timoshenko beam) and 9 (for the two-dimensional beam). Therefore a direct comparison of the mode shapes are feasible for short and long beams.

\subsubsection{Remark}
Due to the fact that any multiple of an eigenfunction is still an eigenfunction, the figures can be scaled. This is why no scale is indicated on the y-axis in \cite{LVV09}. Figures \ref{fig:w} and \ref{fig:phi} were scaled by eye.

\subsection{Comparing the eigenvalues}
\subsubsection{Results of [LVV09]}
The following table contains the results form \cite{LVV09} as well as the corresponding results obtained in this dissertation.
\begin{table}[!ht]
	\centering
	\caption{Results from [LVV09] and results obtained in this dissertation. $0^a$ indicates a 0 as a result of rounding. $\alpha = 1200$}
	\begin{tabular}{|c|c|c|c||c|c|c|c|}
		\hline
		\multicolumn{4}{|c||}{Results from [LVV09]} & \multicolumn{4}{c|}{Dissertation} \\ \hline \hline
		~ & 2D & Timo & RE & ~ & 2D & Timo & RE  \\ \hline
		1 & 0.0317 & 0.0316 & 0.00315 & 1 & 0.031713 & 0.031639 & 0.0023407  \\ 
		2 & 1.14 & 1.14 & $0^a$ & 2 & 1.1413 & 1.1365 &  0.0042050 \\ 
		3 & 7.72 & - & ~ & 3 & 7.7161 & - &   \\ 
		4 & 7.92 & 7.86 & 0.00758 & 4 & 7.918 & 7.8617 & 0.0071116    \\ 
		5 & 26.2 & 25.9 & 0.0115 & 5 & 26.148 & 25.869 & 0.010669  \\ 
		6 & 60.8 & 59.9 & 0.0148 & 6 & 60.816 & 59.946 & 0.014303  \\ 
		7 & 69.3 & - & ~ & 7 & 69.344 & - &    \\ 
		8 & 115 & 113 & 0.0174 & 8 & 115.28 & 113.23 &  0.017787  \\ 
		9 & 192 & 188 & 0.0208 & 9 & 191.57 & 187.55 &  0.020999  \\ 
		10 & 192 & - &  & 10 & 192.03 & - &   \\ 
		11 & 291 & 284 & 0.0241 & 11 & 290.76 & 283.81 &  0.023899  \\ \hline
	\end{tabular}\label{Results_LVV09}
\end{table}

In Table \ref{Results_LVV09}, the results from \cite{LVV09} and the respective results from this dissertation is given. This table shows that the results of \cite{LVV09} can be replicated with more significant digits.\\

Interestingly, with the improved accuracy over \cite{LVV09}, the relative errors remain similar in Table \ref{Results_LVV09}. It is possible that the higher level of accuracy will only benefit larger eigenvalues. Larger eigenvalues are however not of interest for practical applications. The 2nd relative error of \cite{LVV09} however could not be duplicated.

\subsection{Comparison of the eigenvalues}
In \cite{LVV09}, two values for $\alpha$ are considered. For $\alpha = 1200$, the Timoshenko beam has a length to width ratio of $1:10$. For $\alpha = 300$, the Timoshenko beam has length to width ratio of $1:5$. These two values represent a ``long slender'' beam and a ``short'' beam.\\

In Table \ref{tab:Timo}, the eigenvalues of the Timoshenko beam are compared to the first 20 eigenvalues of the two-dimensional beam. The eigenvalues are calculated for different values of $\alpha$.\\

\begin{table}[h!]
	\scalebox{.8}{
	\makebox[\textwidth]{
	\caption{Eigenvalues of a Timoshenko cantilever beam vs the eigenvalues of a cantilever two-dimensional elastic body. RE - Relative Error}
	\begin{tabular}{|cccc||cccc||cccc||cccc|}
		\hline
		\multicolumn{16}{|c|}{Comparison of Eigenvalues} \\
		\hline\hline
		\multicolumn{4}{|c||}{ $h = 1/5$ or $\alpha = 300$}       & \multicolumn{4}{c||}{$h =1/10$ or $\alpha = 1200$}      & \multicolumn{4}{c||}{$h = 1/20$ or $\alpha = 4800$}      & \multicolumn{4}{c|}{$h = 1/30$ or $\alpha = 10800$} \\
		\hline
		{i} & {2D} & {j} & {Timo} & {i} & {2D} & {j} & {Timo} & {i} & {2D} & {j} & {Timo} & {i} & {2D} & {j} & {Timo} \\
		\hline
		{1}&0.12151&1&0.12092&1&0.031713&1&0.031639&1&0.008013&1&0.008004&1&0.003568&1&0.003565\\
		{2}&3.5460&2&3.5071&2&1.1413&2     & 1.1365 & 2     & 0.30756 & 2     & 0.30705 & 2     & 0.13869 & 2     & 0.13855 \\
		\cellcolor{lightgray}{3} & \cellcolor{lightgray}{7.7311} &       & {-} & \cellcolor{lightgray}{3}     & \cellcolor{lightgray}{7.7161} &       & {-} & 3     & 2.3273 & 3     & 2.3213 & 3     & 1.0698 & 3     & 1.0683 \\
		{4} & 20.225 & 3     & 19.869 & 4     & 7.9180 & 3     & 7.8617 & \cellcolor{lightgray}{4}     & \cellcolor{lightgray}{7.7077} &       & {-} & 4     & 4.0140 & 4     & 4.0058 \\
		{5} & 56.109 & 4     & 54.766 & 5     & 26.148 & 4     & 25.869 & 5     & 8.5086 & 4     & 8.4762 & \cellcolor{lightgray}{5}     & \cellcolor{lightgray}{7.7047} &       & {-} \\
		\cellcolor{lightgray}{6} & \cellcolor{lightgray}{69.164} &       & {-} & 6     & 60.816 & 5     & 59.946 & 6     & 21.911 & 5     & 21.794 & 6     & 10.655 & 5     & 10.625 \\
		{7} & 114.03 & 5     & 110.75 & \cellcolor{lightgray}{7}     & \cellcolor{lightgray}{69.344} &       & {-} & 7     & 45.711 & 6     & 45.390 & 7     & 22.975 & 6     & 22.890 \\
		{8} & {189.17} &  6     & {186.50} & 8     & 115.28 & 6     & 113.23 & \cellcolor{lightgray}{8}     & \cellcolor{lightgray}{69.344} &       & {-} & 8     & 43.113 & 7     & 42.909 \\
		\cellcolor{lightgray}{9} & \cellcolor{lightgray}{192.61} &      &  & {9}     & {191.57} &   7  & {187.55} & 9     & 82.887 & 7     & 82.154 & \cellcolor{lightgray}{9}     & \cellcolor{lightgray}{69.331} &       & {-} \\
		{10} & 285.85 & 7     & 277.64 & \cellcolor{lightgray}{10}    & \cellcolor{lightgray}{192.03} &      & {-} & 10    & 136.03 & 8     & 134.58 & 10    & 73.230 & 8     & 72.803 \\
		{11} & 328.40 & 8     & 330.29 & 11    & 290.76 & 8     & 283.81 & \cellcolor{lightgray}{11}    & \cellcolor{lightgray}{192.48} &       & {-} & 11    & 115.41 & 9     & 114.61 \\
		\cellcolor{lightgray}{12} & \cellcolor{lightgray}{357.08} &       & {-} & \cellcolor{lightgray}{12}    & \cellcolor{lightgray}{374.45} &       & {-} & 12    & 207.29 & 9     & 204.69 & 12    & 171.61 & 10    & 170.20 \\
		{13} & 397.33 & 9     & 394.02 & 13    & 413.20 & 9     & 402.27 & 13    & 298.38 & 10    & 294.10 & \cellcolor{lightgray}{13}    & \cellcolor{lightgray}{192.52} &       & {-} \\
		{14} & 442.00   & 10    & 439.52 & 14    & 558.67 & 10    & 542.65 & \cellcolor{lightgray}{14}    & \cellcolor{lightgray}{376.83} &       & {-} & 14    & 243.56 & 11    & 241.26 \\
		\cellcolor{lightgray}{15} & \cellcolor{lightgray}{533.71} &       & {-} & \cellcolor{lightgray}{15}    & \cellcolor{lightgray}{614.11} &       & {-} & 15    & 410.63 & 11    & 404.01 & 15    & 332.83 & 12    & 329.28 \\
		{16} & 538.97 & 11    & 541.55 & 16    & 726.26 & 11    & 704.15 & 16    & 545.03 & 12    & 535.32 & \cellcolor{lightgray}{16}    & \cellcolor{lightgray}{377.16} &       & {-} \\
		\cellcolor{lightgray}{17} & \cellcolor{lightgray}{596.06} &       & {-} & \cellcolor{lightgray}{17}    & \cellcolor{lightgray}{906.28} &       & {-} & \cellcolor{lightgray}{17}    & \cellcolor{lightgray}{621.95} &       & {-} & 17    & 440.77 & 13    & 435.51 \\
		{18} & 602.77 & 12    & 596.09 & 18    & 913.69 & 12    & 884.92 & 18    & 702.30 & 13    & 688.64 & 18    & 568.51 & 14    & 561.04 \\
		\cellcolor{lightgray}{19} & \cellcolor{lightgray}{657.87} &       & {-} & 19    & 1113.7 & 13    & 1080.1 & 19    & 882.95 & 14    & 864.40 & \cellcolor{lightgray}{19}    & \cellcolor{lightgray}{623.05} &       & {-} \\
		{20} & 717.37 & 13    & 731.74 & \cellcolor{lightgray}{20}    & \cellcolor{lightgray}{1218.0}  &       & {-} & \cellcolor{lightgray}{20}    & \cellcolor{lightgray}{927.18} &       & {-} & 20    & 717.04 & 15    & 706.74 \\
		\hline
		\multicolumn{2}{|c}{Max RE:} & \multicolumn{2}{c||}{3.1718\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c||}{3.1486\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c||}{2.1018\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c|}{1.4361\%} \\
		\hline
	\end{tabular}%
	\label{tab:Timo}%
}
}
\end{table}%
\FloatBarrier

The authors of \cite{LVV09} show that for $\alpha = 300$ the eigenvalues compare very well up to the 20th two-dimensional eigenvalue. This is also apparent for other values of $\alpha$ as Table \ref{tab:Timo} shows. In fact, the accuracy of the eigenvalues increases as $\alpha$ increases. The authors of \cite{LVV09} also mention that for $\alpha = 100$, the results are very good for the first 10 eigenvalues.\\

Interestingly, the number of non-beam type eigenvalues within the first few eigenvalues decreases as $\alpha$ increases (or more relevant for the two-dimensional beam: as $h$ decreases).\\

Overall, the relevant conclusions of \cite{LVV09} is verified. For practical applications of beam models, the Timoshenko beam can be used. Preferably for long slender beams as Table \ref{tab:Timo} shows. The authors also mention that nature of the disturbance should also be taken into consideration, mainly because the models are linear.\\

 The short comings of \cite{LVV09} is that damping is not considered. This is mentioned by the authors of \cite{LVV09}, but will not be addressed in this dissertation. Further work is to extend \cite{LVV09} to a three-dimensional comparison. This is investigated further in this chapter.\\

\section{Validity of the two-dimensional elastic model}
This section continues the work of the article \cite{LVV09}. As mentioned by the authors of \cite{LVV09}, a comparison using a three-dimensional beam is desired. The two-dimensional beam model is an intermediate step to be able to investigate the validity of the Timoshenko beam model using a three-dimensional beam model as a reference. As mentioned by the authors, this will avoid some complexities.\\

The approach for this section remain similar to the previous section and the work of \cite{LVV09}. The eigenvalues for both models are calculated using FEM. The eigenvalues are then matched by sketching and comparing the mode shapes. Similar shapes indicate the the eigenvalues are comparable. The eigenvalues are then compared to determine how close the solutions of the models are.\\

As seen in the previous section, the two-dimensional model has eigenvalues not relating to beam type eigenvalues. There will also be eigenvalues not relating to beam type eigenvalues in this section. The results will include some of these eigenvalues, but the focus of our investigation remain with beam problems.

\subsection{The models}
Consider a cantilever two-dimensional beam, referred to as Problem 2D-1 in section \ref{ssec:2D_Model:ModelProblem} and a cantilever three-dimensional beam, referred to as Problem 3D-1 in section \ref{ssec:3D_Model:ModelProblems}.\\

Figure \ref{fig:2Dv3D} shows the two beams side-by-side.

\FloatBarrier
\begin{figure}[h!]
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\caption{Side-by-side comparison of the beams.}
			\label{fig:2Dv3D}
			\centering
			\begin{minipage}[b]{0.7\linewidth}
				\begin{tikzpicture}
					\draw[line width = 0.3mm] (0,1) -- (7,1);
					\draw[line width = 0.3mm] (0,-1) -- (7,-1);
					\draw[line width = 0.3mm] (7,-1) -- (7,1);
					\draw[line width = 0.3mm] (0,-1) -- (0,1);

					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+4});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+3.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+3});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+2.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+2});

					\draw[scale=0.5, domain=-3:0, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1.5});
					\draw[scale=0.5, domain=-3:0, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1});

					\draw[scale=0.5, domain=1:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1.5});
					\draw[scale=0.5, domain=2:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1});

					\draw[scale=0.5, domain=-3:0, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+0.5});
					\draw[scale=0.5, domain=-3:0, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x});
					\draw[scale=0.5, domain=-3:0, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-0.5});
					\draw[scale=0.5, domain=-3:0, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-1});
					\draw[scale=0.5, domain=-3:0, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-1.5});
					\draw[scale=0.5, domain=-3:0, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-2});
					\draw[scale=0.5, domain=-3:1, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-2.5});
					\draw[scale=0.5, domain=-3:2, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-3});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-3.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-4});

					\node at (7.2,0) {$h$};
					\node at (3.2,-1.2) {$\ell = 1$};
				\end{tikzpicture}
				\subcaption{Two-Dimensional Elastic Body}
			\end{minipage}
			\begin{minipage}[b]{0.7\linewidth}
				\begin{tikzpicture}
					\draw[line width = 0.3mm] (-0.1,1) -- (6.9,1);
					\draw[line width = 0.3mm] (-0.1,-1) -- (6.9,-1);
					\draw[line width = 0.3mm] (6.9,-1) -- (6.9,1);
					\draw[line width = 0.3mm] (-0.1,-1) -- (-0.1,1);

					\draw[line width = 0.3mm] (0.1,1.1) -- (7.1,1.1);
					\draw[line width = 0.3mm] (7.1,-0.9) -- (7.1,1.1);

					\draw[line width = 0.3mm] (-0.1,1) -- (0.1,1.1);
					\draw[line width = 0.3mm] (6.9,1) -- (7.1,1.1);
					\draw[line width = 0.3mm] (6.9,-1) -- (7.1,-0.9);

					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+4});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+3.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+3});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+2.5});

					\draw[scale=0.5, domain=-3:-0.1, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+2});
					\draw[scale=0.5, domain=-3:-0.1, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1.5});
					\draw[scale=0.5, domain=-3:-0.1, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1});

					\draw[scale=0.5, domain=0.5:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+2});
					\draw[scale=0.5, domain=1.5:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1.5});
					\draw[scale=0.5, domain=2.5:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1});

					\draw[scale=0.5, domain=-3:-0.1, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+0.5});
					\draw[scale=0.5, domain=-3:-0.1, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x});
					\draw[scale=0.5, domain=-3:-0.1, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-0.5});
					\draw[scale=0.5, domain=-3:-0.1, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-1});
					\draw[scale=0.5, domain=-3:-0.1, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-1.5});
					\draw[scale=0.5, domain=-3:0, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-2});
					\draw[scale=0.5, domain=-3:1, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-2.5});
					\draw[scale=0.5, domain=-3:2, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-3});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-3.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-4});

					\node at (6.95,1.25) {$b$};
					\node at (6.78,0) {$h$};
					\node at (3.2,-1.2) {$\ell = 1$};

				\end{tikzpicture}
				\subcaption{Three-Dimensional Elastic Body}

			\end{minipage}
		}
	}
\end{figure}
\FloatBarrier

For both models the parameter $h$ represents the width of the beams. The three-dimensional model also have depth $b$. Intuitively, the two-dimensional beam should compare well the the three-dimensional beam if $b$ is small.

\subsection{Accuracy of the numerical eigenvalues}
In the previous section, the accuracy of the two-dimensional beam's eigenvalues were investigated. This can be done similarly for the three-dimensional beam. The eigenvalues are also calculated using the Finite Element Method. Using piecewise tri-cubic basis functions, the following results were obtained.\\

****************INSERT GRAPHIC**************\\

\subsection{Comparing the mode shapes}
As in \cite{LVV09}, the shape of the modes are compared. Similar mode shapes implies that the eigenvalues are comparable. Results for both beam and non-beam type eigenvalues are included. The focus still remains on beam type eigenvalues.

\subsubsection{Mode shapes relating to beam type eigenvalues.}
Figure \ref{fig:beam-2dv3d} show some examples of beam type mode shapes for the displacement $u$. These are shapes that can be obtained from the Timoshenko beam model.

\begin{figure}[h!]
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{3D12.png}
				\subcaption{3D Beam Type - $\lambda_{12} = 2.3293$}
				\label{fig:minipage2}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{2D3.png}
				\subcaption{2D Beam Type - $\lambda_3 = 2.3273$}
				\label{fig:minipage1}
			\end{minipage}
	}}
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{3D22.png}
				\subcaption{3D Beam Type - $\lambda_{24} = 21.929$}
				\label{fig:minipage2}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{2D6.png}
				\subcaption{2D Beam Type - $\lambda_6 = 21.911$}
				\label{fig:minipage1}
			\end{minipage}
			\caption{Mode shapes of the displacement $w$ with $h=1/20$.}
			\label{fig:beam-2dv3d}
	}}
\end{figure}
\FloatBarrier

\subsubsection{Mode shapes relating to non-beam type eigenvalues that are present in the two-dimensional model.}
Figure \ref{fig:nonbeam-2dv3d} show examples of mode shapes relating to non-beam type eigenvalues for the displacement $u$. These mode shapes are the non-beam type eigenvalues found in both the two-dimensional beam and the three-dimensional beam model.

\FloatBarrier
\begin{figure}[h!]
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{3D33.png}
				\subcaption{3D Non-Beam Type - $\lambda_{33} = 69.374$}
				\label{fig:minipage2}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{2D8.png}
				\subcaption{2D Non-Beam Type - $\lambda_8 = 69.344$}
				\label{fig:minipage1}
			\end{minipage}
			\caption{Mode shapes of the displacement $w$ with $h=1/20$.}
			\label{fig:nonbeam-2dv3d}
	}}
\end{figure}
\FloatBarrier

\subsubsection{Mode shapes relating to non-beam type eigenvalues that are not present in the two-dimensional model.}
Figure \ref{fig:nonbeam-2dv3d} show examples of mode shapes relating to non-beam type eigenvalues for the displacement $u$ which are not present in the two-dimensional model. These mode shapes only appear in the three-dimensional beam.
\FloatBarrier
\begin{figure}[h!]
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{3DNonBeam10.png}
				\subcaption{Non-2D Type - $\lambda_{10} = xx$}
				\label{fig:minipage2}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{3DNonBeam11.png}
				\subcaption{Non-2D Type - $\lambda_{11} = xx$}
				\label{fig:minipage11}
			\end{minipage}
			\caption{Mode shapes of the displacement $w$ with $\alpha = 4800$.}

	}}
\end{figure}
\FloatBarrier

Figure \ref{fig:minipage11} looks similar to the beam type eigenvalues, however the shape is in the $x-z$ axis.\\

\subsection{Comparing the eigenvalues}
The eigenvalues can now be compared. In the upcoming tables, the eigenvalues of the two-dimensional beam are compared to the eigenvalues of the three-dimensional beam. For each case of $h$, different values of $b$ are given.\\

Eigenvalues relating to non-beam type eigenvalues that are present in both the two-dimensional and three-dimensional beam, are highlighted in grey. The other non-beam type eigenvalues that are only found in the three-dimensional beam, are omitted from the tables. The original numbering is maintained.\\

The two cases for $h$ are $h = 1/5$ for a short beam and $1/20$ for a slender beam. For now, the investigation is restricted to values where $b \leq h$.

\begin{table}[htbp]
	\scalebox{.8}{
	\makebox[\textwidth]{
		\caption{Comparsion of Eigenvalues with $h = 1/5$.}
		\begin{tabular}{|cc|cc|cc|cc||cc|}
			\hline
			\multicolumn{10}{|c|}{Eigenvalues} \\
			\hline
			\hline
			i     & {b = h} & i     & {b = 1/2 h} & i     & {b = 1/4 h} & i     & {b = 1/8 h} & j     & 2D \\
			\hline
			{2} & 0.12307 & {2} & 0.12234 & {2} & 0.12198 & {3} & 0.12178 & 1     & 0.12151 \\
			{3} & 3.5773 & {5} & 3.5630 & {6} & 3.5558 & {8} & 3.5519 & 2     & 3.5460 \\
			\rowcolor{lightgray}{5} & 7.7799 & {6} & 7.7596 & {8} & 7.7471 & {11} & 7.7401 & 3     & 7.7311 \\
			{8} & 20.334 & {9} & 20.283 & {11} & 20.26 & {14} & 20.247 & 4     & 20.225 \\
			{10} & 56.247 & {12} & 56.173 & {15} & 56.156 & {22} & 56.142 & 5     & 56.109 \\
			\rowcolor{lightgray}{11} & 69.197 & {14} & 69.319 & {17} & 69.281 & {24} & 69.238 & 6     & 69.164 \\
			{14} & 114.03 & {16} & 114.01 & {20} & 114.05 & {29} & 114.06 & 7     & 114.03 \\
			\rowcolor{lightgray}{17} & 187.01 & {19} & 189.14 & {25} & 189.37 & {36} & 189.34 & 8     & 189.17 \\
			{18} & 192.21 & {20} & 192.41 & {26} & 192.58 & {37} & 192.63 & 9     & 192.61 \\
			{21} & 284.76 & {23} & 285.43 & {31} & 285.74 & {42} & 285.84 & 10    & 285.85 \\
			{23} & 327.57 & {26} & 328.24 & {35} & 328.37 & {46} & 328.40 & 11    & 328.40 \\
			\rowcolor{lightgray}{25} & 347.77 & {28} & 356.44 & {36} & 357.30 & {50} & 357.33 & 12    & 357.08 \\
			{27} & 393.69 & {30} & 396.84 & {38} & 397.28 & {53} & 397.37 & 13    & 397.33 \\
			{30} & 434.46 & {34} & 441.05 & {41} & 441.81 & {57} & 441.99 & 14    & 442.00 \\
			\rowcolor{lightgray}{31} & 523.65 & {36} & 534.04 & {43} & 534.17 & {63} & 534.03 & 15    & 533.71 \\
			{34} & 550.51 & {37} & 537.82 & {44} & 538.86 & {64} & 539.06 & 16    & 538.97 \\
			\rowcolor{lightgray}{37} & 590.86 & {41} & 587.43 & {48} & 594.17 & {65} & 595.58 & 17    & 596.06 \\
			{39} & 590.86 & {42} & 600.52 & {49} & 602.25 & {67} & 602.69 & 18    & 602.77 \\
			\rowcolor{lightgray}{42} & 646.21 & {44} & 657.22 & {50} & 658.04 & {71} & 658.06 & 19    & 657.87 \\
			{44} & 711.07 & {46} & 714.62 & {53} & 717.10 & {73} & 717.51 & 20    & 717.37 \\
			\hline
			\hline
			\multicolumn{2}{|c|}{Max RE: \  2.6069\%} &\multicolumn{2}{c|}{Max RE: \ 1.4469\%}  & \multicolumn{2}{c|}{Max RE: \  0.38192\%}  & \multicolumn{2}{c||}{Max RE: \ 0.22301\%}& \multicolumn{2}{c|}{-} \\
			\hline
		\end{tabular}%
		\label{tab:2v3_1}%
	}}
\end{table}%
\FloatBarrier
Table \ref{tab:2v3_1} show that the eigenvalues of the two-dimensional model compare very well to the three-dimensional model.  The maximum relative error decreases significantly as $b$ is decreased. Even with $b = h$ the maximum relative error is still less than $2.7\%$. \\

Table \ref{tab:2Dv3D_1_breakup} gives more information on the maximum relative error by separating the beam type and non-beam type eigenvalues.

% Table generated by Excel2LaTeX from sheet 'Sheet2'
\begin{table}[htbp]
	\centering
	\caption{****$h = 1/5$}
	\begin{tabular}{|c|cccc|}
		\hline
		\multicolumn{5}{|c|}{Maximum Relative Error} \\
		\hline
		\hline
		& {b = h} & {b = 1/2h} & {b = 1/4h} & {b = 1/8h} \\
		\hline
		Beam Type & 2.1420 \% & 0.6804 \% & 0.38192 \% & 0.22301 \% \\
		Non-Beam Type & 2.6069 \% & 1.4469 \% & 0.31546 \% & 0.11680 \% \\
		\hline
	\end{tabular}%
	\label{tab:2Dv3D_1_breakup}%
\end{table}%

From Table \ref{tab:2Dv3D_1_breakup}, it is clear that indeed for beam type eigenvalues, a short two-dimensional beam compares very well to a short three-dimensional beam.\\

\begin{table}[ht]
	\scalebox{.8}{
	\makebox[\textwidth]{
	\caption{Comparsion of Eigenvalues with $h = 1/20$. The non beam-type eigenvalues are highlighted or omitted.}
	\begin{tabular}{|cc|cc|cc|cc||cc|}
		\hline
		\multicolumn{10}{|c|}{Eigenvalues} \\
		\hline
		\hline
		i     & {b = h} & i     & {b = 1/2 h} & i     & {b = 1/4 h} & i     & {b = 1/8 h} & j     & 2D \\
		\hline
		{2} & 0.008043 & {2} & 0.008029 & {2} & 0.008023 & {3} & 0.00802 & {1} & {0.008013} \\
		{3} & 0.3087 & {4} & 0.30816 & {5} & 0.30794 & {7} & 0.30785 & {2} & {0.30757} \\
		{5} & 2.3357 & {8} & 2.3316 & {9} & 2.3300  & {12} & 2.3293 & {3} & {2.3273} \\
		\rowcolor{lightgray}{8} & 7.7217 & {10} & 7.7156 & {13} & 7.7124 & {16} & 7.7111 & {4} & {7.7077} \\
		{10} & 8.5387 & {11} & 8.5238 & {14} & 8.5182 & {18} & 8.516 & {5} & {8.5086} \\
		{11} & 21.986 & {14} & 21.948 & {18} & 21.934 & {24} & 21.929 & {6} & {21.911} \\
		{14} & 45.863 & {18} & 45.781 & {21} & 45.756 & {30} & 45.746 & {7} & {45.712} \\
		\rowcolor{lightgray}{17} & 69.444 & {19} & 69.408 & {25} & 69.385 & {33} & 69.374 & {8} & {69.344} \\
		{18} & 83.149 & {22} & 82.999 & {27} & 82.960 & {36} & 82.944 & {9} & {82.887} \\
		{21} & 136.44 & {25} & 136.19 & {31} & 136.14 & {42} & 136.12 & {10} & {136.03} \\
		\rowcolor{lightgray}{23} & 192.62 & {27} & 192.62 & {35} & 192.58 & {47} & 192.56 & {11} & {192.48} \\
		{25} & 207.87 & {29} & 207.5 & {36} & 207.43 & {48} & 207.41 & {12} & {207.29} \\
		{27} & 299.14 & {33} & 298.63 & {41} & 298.56 & {55} & 298.53 & {13} & {298.38} \\
		\rowcolor{lightgray}{30} & 376.68 & {35} & 377.01 & {44} & 377.01 & {58} & 376.98 & {14} & {376.83} \\
		{31} & 411.58 & {37} & 410.89 & {46} & 410.83 & {61} & 410.81 & {15} & {410.63} \\
		{34} & 546.15 & {40} & 545.27 & {50} & 545.24 & {66} & 545.23 & {16} & {545.03} \\
		\rowcolor{lightgray}{37} & 620.77 & {42} & 622.02 & {53} & 622.19 & {69} & 622.18 & {17} & {621.95} \\
		{39} & 703.55 & {44} & 702.49 & {54} & 702.29 & {71} & 702.53 & {18} & {702.31} \\
		{42} & 884.27 & {47} & 883.02 & {59} & 883.14 & {82} & 883.19 & {19} & {882.96} \\
		\rowcolor{lightgray}{44} & 923.68 & {49} & 926.88 & {60} & 927.43 & {86} & 927.50 & {20} & {927.18} \\
		\hline
		\hline
		\multicolumn{2}{|c|}{Max RE: \  0.37701\%} &\multicolumn{2}{c|}{Max RE: \ 0.19893\%}  & \multicolumn{2}{c|}{Max RE: \  0.12393\%}  & \multicolumn{2}{c||}{Max RE: \ 0.092843\%}& \multicolumn{2}{c|}{-} \\
		\hline
	\end{tabular}%
	\label{tab:2v3_2}%
}}
\end{table}%
\FloatBarrier

In Table \ref{tab:2v3_2}, the eigenvalues are again compared for $h = 1/20$. The maximum relative error shows very impressive results for this case. The maximum relative error is less than $0.4\%$ and the maximum relative error decreases as $b$ is decreased.\\

\begin{table}[htbp]
	\centering
	\caption{*****$h = 1/20$}
	\begin{tabular}{|c|cccc|}
		\hline
		\multicolumn{5}{|c|}{Maximum Relative Error} \\
		\hline
		\hline
		& {b = h} & {b = 1/2h} & {b = 1/4h} & {b = 1/8h} \\
		\hline
		Beam Type & 0.37701 \% & 0.19893 \% & 0.12393 \% & 0.092843 \% \\
		Non-Beam Type & 0.37682 \% & 0.10218 \% & 0.061213 \% & 0.043601 \% \\
		\hline
	\end{tabular}%
	\label{tab:2v3_2_split}%
\end{table}%
\FloatBarrier

Separating the errors of the beam-type eigenvalues and the non-beam type eigenvalues in Table \ref{tab:2v3_2_split}, just reaffirms the results.\\

Restricting the models where $b \leq h$ yielded great results. It has been shown that the two-dimensional beam compares very well the the three-dimensional beam. The comparison also improves for smaller values of $h$. This is a similar result as seen with the work of \cite{LVV09} in the previous section.

\subsection{$b > h$}
Now we can ask the question what happens if $b > h$. Of course for very large values of $b$, the question should be asked is if the beam model is still relevant. Therefore, the investigation will look at an incremental increase of $b$, without making assumptions on the relevancy of a beam model.\\

\FloatBarrier
\begin{table}[htbp]
	\scalebox{.8}{
	\makebox[\textwidth]{
		\caption{$h = 1/20$}
		\begin{tabular}{|cc|cc|cc||cc|}
			\hline
			\multicolumn{8}{|c|}{Eigenvalues} \\
			\hline
			\hline
			i     & {b = 2h} & i     & {b = 4h} & i     & {b = 8h} & j     & 2D \\
			\hline
			{2} & 0.008076 & {1} & 0.008162 & {1} & 0.008324 & 1     & 0.008013 \\
			{3} & 0.30995 & {3} & 0.31298 & {3} & 0.31738 & 2     & 0.30757 \\
			{5} & 2.3462 & {5} & 2.3737 & {6} & 2.4116 & 3     & 2.3273 \\
			\rowcolor{lightgray}{8} & 7.7312 & {8} & 7.7471 & {9} & 7.7711 & 4     & 7.7077 \\
			{10} & 8.5841 & {9} & 8.7082 & {10} & 8.7929 & 5     & 8.5086 \\
			{11} & 22.124 & {12} & 22.491 & {14} & 23.222 & 6     & 21.911 \\
			{14} & 46.195 & {14} & 47.003 & {18} & 47.921 & 7     & 45.712 \\
			\rowcolor{lightgray}{17} & 69.454 & {17} & 69.281 & {22} & 72.307 & 8     & 69.344 \\
			{18} & 83.822 & {18} & 85.218 & {24} & 80.607 & 9     & 82.887 \\
			{21} & 137.43 & {21} & 138.58 & {32} & 140.97 & 10    & 136.03 \\
			\hline
			\hline
			\multicolumn{2}{|c|}{Max RE: \  1.1289\%} &\multicolumn{2}{c|}{Max RE: \ 2.8261\%}  & \multicolumn{2}{c|}{Max RE: \  5.9821\%} & \multicolumn{2}{c|}{-} \\
			\hline
		\end{tabular}%
		\label{tab:b>h}%
	}}
\end{table}
\FloatBarrier
In Table \ref{tab:b>h}, $h = 1/20$ was chosen since it was the beam case investigated in the previous section and $b$ is increased by a factor of $2$. And here we see that the maximum relative error increases almost by a factor of $2$ for each case. Note also that this table only shows the first 10 eigenvalues.\\

There is also an increase of non-beam type eigenvalues that are only present in the three-dimensional model.\\

In Table \ref{tab:b>h-split}, the maximum relative error of the beam-type and non-beam type eigenvalues are separated.
\FloatBarrier
\begin{table}[htbp]
	\centering
	\caption{$\alpha = 4800$}
	\begin{tabular}{|c|ccc|}
		\hline
		\multicolumn{4}{|c|}{Maximum Relative Error} \\
		\hline
		\hline
		& {b = 2h} & {b = 4h} & {b = 8h} \\
		\hline
		Beam Type & 1.1289 \% & 2.8261 \% & 5.9821 \% \\
		Non-Beam Type & 0.30521 \% & 0.51096 \% & 4.2734 \% \\
		\hline
	\end{tabular}%
	\label{tab:b>h-split}%
\end{table}%
\FloatBarrier

From Table \ref{tab:b>h-split}, it is clear that as $b$ increases, the maximum relative error of the beam type eigenvalues increases and the two-dimensional model compares less to the three-dimensional model.\\

In this section, it has been shown that the two-dimensional beam compares very well to the three-dimensional beam, in practical beam problems. Practical beam problems imply that $b \leq h$. The models also compare much better if $h$ is smaller. For values of  $b > h$, the use of a beam model should be questioned.\\

For $b > h$, another model might be presented. In the next section, a Reisner-Mindlin plate model is investigated.
\end{document}

\begin{figure}[h!]\label{fig:compare:1D+2D}
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\caption{Side by side comparison of the beams.}
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\begin{tikzpicture}
					\draw[line width = 0.3mm] (0,0) -- (7,0);

					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+4});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+3.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+3});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+2.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+2});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1});	
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+0.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-0.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-1});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-1.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-2});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-2.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-3});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-3.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-4});

					%\node at (3.2,-0.2) {$\ell = 1$};
				\end{tikzpicture}
				\subcaption{Timoshenko Cantilever Beam}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\begin{tikzpicture}
					\draw[line width = 0.2mm] (0,0.1) -- (7,0.1);
					\draw[line width = 0.2mm] (0,-0.1) -- (7,-0.1);
					\draw[line width = 0.2mm] (7,-0.1) -- (7,0.1);
					\draw[line width = 0.2mm] (0,-0.1) -- (0,0.1);

					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+4});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+3.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+3});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+2.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+2});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+0.5});

					\draw[scale=0.5, domain=-3:0, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x});
					\draw[scale=0.5, domain=-3:0.5, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-0.5});
					\draw[scale=0.5, domain=-3:1.5, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-1});
					\draw[scale=0.5, domain=-3:2.5, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-1.5});

					\draw[scale=0.5, domain=0.5:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x});
					\draw[scale=0.5, domain=1.5:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-0.5});
					\draw[scale=0.5, domain=2.5:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-1});

					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-2});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-2.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-3});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-3.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-4});

					\draw[line width = 0.1mm,->] (6,2) -- (7.4,2);
					\draw[line width = 0.1mm,->] (6,2) -- (6,0.9);
					\node at (6.7,1.8) {$e_1$};
					\node at (5.8,1.4) {$e_2$};

					\node at (7.2,0) {$h$};
					%\node at (3.2,-0.3) {$\ell = 1$};

				\end{tikzpicture}
				\subcaption{Two-Dimensional Cantilever Beam}
			\end{minipage}
		}
	}
\end{figure}