### 2023-04-24 work_plan.tex
workplan

\documentclass[12pt, a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage{amsmath, amssymb}
\usepackage{titlesec}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{float}
\usepackage{hyperref}
\usepackage{parskip}
\usepackage{wasysym} % for checkboxes

\geometry{
 a4paper,
 total={170mm,257mm},
 left=20mm,
 top=20mm,
 }

\titleformat{\section}{\normalfont\Large\bfseries}{\thesection.}{1em}{}
\titleformat{\subsection}{\normalfont\large\bfseries}{\thesubsection}{1em}{}

\title{Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates}
\author{Rudi du Toit}
\date{}

\begin{document}

\maketitle

\noindent
\textbf{Supervisor:} \\
Prof. J. van Rensburg \\
Email: jvrnf2013@gmail.com \\

\noindent
\textbf{Co-supervisor:} \\
Dr. M. Labuschagne  \\
Email: madelein.labuschagne@up.ac.za \\

\section*{Current Status of the Thesis}
The dissertation is mainly complete, and in the review phase. There are a few sections in which the theory is being updated.

\section*{Work Plan}
\subsection*{22-04-2023 - 30-04-2023}
\subsubsection*{Article work}
\begin{itemize}[label=\Square]
  \item Task 0.1: Reading articles.
\end{itemize}

\subsubsection*{Chapter 1: Comparison of models}
\begin{itemize}[label=\Square]
  \item Task 1.1: Review equation and section references.
\end{itemize}

\subsubsection*{Chapter 2: Mathematical analysis of vibration problems}
\begin{itemize}[label=\Square]
  \item Task 2.1: General vibration problems
  \item Task 2.2: Add series solution to the eiegenvalue problem
\end{itemize}

\subsubsection*{Chapter 3: Finite element theory}
\begin{itemize}[label=\Square]
  \item Task 3.1: Apply suggestion from chapter 2 onto chapter 3 ahead of review.
\end{itemize}

\subsubsection*{Chapter 4: Timoshenko beam model}
\begin{itemize}[label=\Square]
  \item Task 4.1: Expand the general solutions of the eigenvalue problem and link it to chapter 2.
\end{itemize}

\subsubsection*{Chapter 5: Finite element method}
\begin{itemize}[label=\Square]
  \item Task 5.1: Self review of Chapter 5.
\end{itemize}

\subsubsection*{Chapter 6: Validity of cantilever beam and plate models}
\begin{itemize}[label=\Square]
  \item Task 6.1: Self review of Chapter 6.
  \item Task 6.2: Re-calculate the numerical results for the three-dimensional beam and two-dimensional plate model.
  \item Task 6.3: Finish the comparison of the three-dimensional beam and two-dimensional plate model.
\end{itemize}

\subsubsection*{Chapter 7: Conclusion}
\begin{itemize}[label=\Square]
  \item Task 6.1: Second draft of the conclusion and summary.
\end{itemize}

\section*{Additional Notes and Remarks}
This week is a focus week on completing some task that have been set a side for some time. It is also necessary just to read up on articles.

\section*{Submission Deadlines}
\begin{itemize}
\item First draft submission: July 2023
\item Feedback from supervisor and co-supervisor: June 2023
\item Final draft submission: TBC
\item Defense date: TBC
\end{itemize}

\end{document}