### 2023-05-15 files/Chapter7/Introduction.tex
Chapter 7

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Introduction}
This chapter is an investigation into the validity of a cantilever Timoshenko beam and a cantilever Reissner-Mindlin plate model for practical applications. Validity in this context refers to how well the solutions of the model compare to the solutions of more realistic higher-dimensional model.

The chapter consists of three main parts.

Section 7.2 is a discussion and a verification of numerical results of the work done in \cite{LVV09} In this article, the authors compare a cantilever Timoshenko beam to a cantilever two-dimensional beam. We also include some extra results for different values of the parameter $\alpha$.

Section 7.3 is a natural extension of the article \cite{LVV09} which is mentioned by the authors. In this section, the validity of the two-dimensional beam model use in Section 7.2 is investigated. The results of the two-dimensional beam are compared to a three-dimensional beam.

The two-dimensional beam is an intermediate step to validate the Timoshenko beam model. A direct comparison between the one and three dimensional would be preferred, but the authors of \cite{LVV09} mention that this will introduce complexities, which will be discussed in more detail in the numerical results.

Section 7.4 extends the work of \cite{LVV09} further to plate models. Specifically a comparison of a cantilever two-dimensional Reissner-Mindlin plate to a cantilever three-dimensional plate.

\subsubsection{Global parameters}
All the models in this dissertation are made of the same isotropic material and have square cross-section. Therefore for all the models, the elastic modulus $\displaystyle G = \frac{E}{2(1+\nu)}$. The shear correction factor is chosen as $\kappa^2 = 5/6$ and Poisson's ratio $\nu = 0.3$.\\

\subsubsection{Global Parameters}
The models in this Chapter (and this dissertation) assume the use of an identical isotropic material with, a rectangular cross-section. These are defined in global parameters shared by all the models. The parameters are as follows:

\begin{itemize}
    \item Elastic modulus ($G$): This is calculated using the formula $G = \frac{E}{2(1+\nu)}$, where $E$ is the modulus of elasticity and $\nu$ is Poisson's ratio.

    \item Shear correction factor ($\kappa^2$): This is set to $5/6$, which is common for rectangular cross-sections.

    \item Poisson's ratio ($\nu$): This is set to $0.3$, a typical value for materials like steel used in engineering.
\end{itemize}

The consistent use of these parameters across all models helps to ensure that any differences in the results can be attributed to the model structures themselves, rather than variations in the material or geometric properties.
\end{document}

This is the same idea as done in the article [LVV09]. To make the comparison between models of different dimensions, the eigenvalues and eigenfuctions of the models are considered. If the eigenvalues and eigenfunctions compare well, it can be assumed that the solution to the model problem would also compare well. The reason for this is explained in Section 1.5.5. And since the Fourier coefficients in the solution decreases very rapidly, only the first few eigenvalues and eigenfunctions need to be considered.\\

Section 6.2 looks at some of the results obtained in [LVV09]. The results for the Euler-Bernoulli beam theory is excluded a the article has sufficient results. The idea of Section 6.2 is to verify the conclusions obtained in [LVV09]. The results of the section also aggregate the work of the article by examining different values for the parameter $\alpha$.\\

Section 6.3 extends the work of [LVV09] by examining the comparison between a two-dimensional and three-dimensional cantilever beam model. The authors' opinion is that a comparison between a two-dimensional and three-dimensional model would be less problematic than comparing the one-dimensional (Timoshenko) beam model to a three-dimensional model. The two-dimensional model can be seen as an intermediate step.

\subsubsection{Parameters}
For all models, the parameters are chosen as $\nu = 0.3$, and $\kappa^2 = 5/6$. It is assumed that all models are made from the same isotropic material, hence $G = \frac{E}{2(1+\nu)}$. Since the beams have a rectangular cross-section, for the Timoshenko beam model the area moment of inertia can be calculated as
\begin{eqnarray*}
	I & = & \frac{h^3b}{12}.
\end{eqnarray*}
The dimensionless constant $\alpha$, given in Section 1.5.1, gives the following relationship between the beam's height, length and the value of $\alpha$:
\begin{eqnarray*}
	\frac{h}{\ell}  & = & \sqrt{\frac{12}{\alpha}}
\end{eqnarray*}
This equation show that the ratio of the height of the beam and the length is dependant on the value of the parameter $\alpha$. The parameter $\alpha$ only appears in the one-dimensional beam model. But for continuity will be used for all models in this chapter to determine the ratio of length to height of the models.

\subsubsection{Remark:}
\begin{itemize}
	\item[--] All results in this chapter are given to 3 significant digits.
	\item[--] ****b does not depend on $\alpha$
\end{itemize}

**********************Needs work******************\\
The comparisons in this Chapter are as follow:
\begin{itemize}
	\item[] Section 7.2: The Timoshenko cantilever beam (Problem T-2) and two-dimensional cantilever beam (Problem 2D-1).
	\item[] Section 7.3: The two-dimensional cantilever beam (Problem 2D-1) and three-dimensional cantilever beam(Problem 3D-1)
	\item[] Section 7.4: The Reisner-Mindlin cantilever plate (Problem P-1) and three-dimensional cantilever plate (Problem 3D-1)
\end{itemize}

For the first part of this chapter, the work of the authors of \cite{LVV09} are discussed. The authors of

This chapter is a discussion and an extension of the work of the authors of \cite{LVV09}. In \cite{LVV09}, the authors examined the validity of a cantilever Timoshenko beam using a cantilever two-dimensional beam as a reference.\\

After the discussion of \cite{LVV09}, the work of the authors is extended to investigate the validity of a cantilever two-dimensional beam using a cantilever three-dimensional beam as a reference. Finally the validity of a cantilever Reissner-Mindlin plate is investigated using a cantilever three-dimensional plate model as reference. The structure and idea of the authors of \cite{LVV09} is used a guide throughout this chapter.\\

The chapter is divided into the following sections:\\