### 2023-07-21 files/Chapter3/dynamicconvergence.tex

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Introduction}
In this chapter, two important convergence results of the Finite Element Method
(FEM) is discussed. The first result looks at the convergence of the Galerkin
approximation for second order hyperbolic type problems, as presented
in~\cite{BV13}. The second result, from the textbook~\cite{SF97}, examines the
convergence of eigenvalues and eigenfunctions for a one-dimensional vibration
problem, using the Finite Element Method.

\section{Convergence of the Galerkin approximation}\label{sec:convergence-of-the-galerkin-approximation}
In the article \cite{BV13}, the authors investigate the convergence of the
Galerkin approximation for second order hyperbolic type problems. The article
\cite{BV18} extends the work of \cite{BV13} by including general damping and
damping at the endpoints. For the models in Chapter 1, \cite{BV13} is
sufficient while \cite{BV18} provides more insight and improved notation.

In Section 2.2, Problem GVar is presented and identical to the problem in
\cite{BV13}. The problem is repeated here for convenience.

\subsubsection*{Problem GVar}
Given a function $f:J\rightarrow X$, find a function $u\in C(J,\ X)$ such that $u'$ is continuous at $0$ with respect to
$\Vert \cdot \Vert_{W}$ and for each $t\in J,\ u(t)\in V,\ u'(t) \in V,\ u''(t)\in W$ and
\begin{eqnarray}
	c(u''(t),v)+a(u'(t),v)+b(u(t),v)= (f(t),v)_{X} \ \ \ \ \textrm{for each} \ v \in V, \label{GV_PB1}
\end{eqnarray}
while $u(0)=u_{0}$, $u'(0)=u_{1}$.

Assume that the assumptions \textbf{A1}-\textbf{A4} from Section 2.2 are
satisfied ensuring that Problem GVar has a unique solution.

Before the theory of \cite{BV13} can be discussed, some preliminary work is
necessary. The structure of this section is as follows. First the Galerkin
approximation is for Problem GVar is derived. Then a equivalent system of
ordinary differential equations is derived using the Finite Element Method.
Finally, the convergence of the Galerkin approximation is discussed using the
work of the article \cite{BV13}.

Recall from Chapter 2 that $V$ is the closure of the test function space $T$ in
the Hilbert space $H$.

\subsubsection*{Problem GVar$^h$}
Given a function $f: J \rightarrow X$, find a function $u_h \in C^2(J)$ such that for each $t\in J$
\begin{eqnarray}
	c(u_h''(t),v)+a(u_h'(t),v)+b(u_h(t),v)= (f(t),v)_{X} \ \ \ \ \textrm{for each} \ v \in S^h, \label{DC_E2}
\end{eqnarray}
with the initial values $u_h(0)=u^h_{0}$ and $u'_h(0)=u^h_{1}$. The initial conditions $u^h_{0}$ and $u^h_{1}$ are projections of $u_0$ and $u_1$ into the finite dimensional space $S^h$.

If the function $f$ is continuous, then Problem GVar$^h$ is equivalent to a
system of second order differential equations. The following theorem from
\cite{BV13} proves that if $f$ is continuous, then problem Problem GVar$^h$ has
a unique solution.

\newtheorem{DC_Thm2}{Theorem}
\begin{DC_Thm2} \label{DC_THM2}
	If $f \in C(J,X)$, then there exists a unique solution $u_h \in C^2(J)$ for Problem GVar$^h$ for each $u_0^h$ and $u_1^h$ in $S^h$. If $f = 0$ then $u_h \in C^2((-\infty, \infty))$.
\end{DC_Thm2}

This theorem is presented and proved in \cite{BV13}.

\subsection{System of ordinary differential equations}
Consider the standard FEM matrices defined by
\begin{eqnarray*}
	K_{ij} & = & b(\phi_j, \phi_i)\\
	C_{ij} & = & a(\phi_j, \phi_i)\\
	M_{ij} & = & c(\phi_j, \phi_i)\\
	F_{i}(t) & = & b(f(t), \phi_i)
\end{eqnarray*}

Using these matrices, Problem GVar$^h$ is rewritten as a system of ordinary
differential equations denoted by Problem ODE. \subsubsection*{Problem ODE}
Find a function $\bar{u} \in S^h$ such that
\begin{eqnarray}
	M\bar{u}'' + C \bar{u}' + K\bar{u} = F(t) \ \ \ \textrm{ with } \bar{u}(0) = \bar{u}^h_0 \textrm{ and }  \bar{u}(1) = \bar{u}^h_1 \label{ODE}
\end{eqnarray}
The function $F$ is the interpolation of the function $f$ into the discrete space $S^h$.

The following propositions related to Problem ODE are given in \cite{BV13}.

\newtheorem{DC_Prop2}{Proposition}
\begin{DC_Prop2}
	If $F\in C(J)$, then Problem ODE has a unique solution for each pair of vectors $\bar{u}^h_0$ and $\bar{u}^h_1$
\end{DC_Prop2}

\newtheorem{DC_Prop3}[DC_Prop2]{Proposition}
\begin{DC_Prop3}
	Suppose $M, K, C, F,\bar{u}^h_0$ and $\bar{u}^h_1$ are defined as above. Then, the function $u_{h}$ is a solution of Problem GVar$^{h}$ if and only if the function $\bar{u}$ is a solution of Problem ODE.
\end{DC_Prop3}

Since the set $\{\phi_1, \ \phi_2,...,\phi_n\}$ forms a linearly independent
basis for $S^h$, the bilinear form $c$ is positive definite on $S^h$.
Consequently, the matrix $M$ is invertible. The equation \eqref{ODE} can
therefore be rewritten as $\bar{u}'' + M^{-1}C\bar{u}' + M^{-1}K \bar{u} =
	M^{-1}F(t)$. Using the theory of linear differential equations, a unique
solution for Problem ODE can be found, proving Proposition 1.

Proposition 2 provides a link between the solution of Problem ODE and the
solution of Problem GVar$^h$.

\subsection{Error estimates}\label{e_est}
Under the assumptions \textbf{A1}-\textbf{A4} of Section 2 and continuity of $f$, there exists unique solutions for Problem GVar and Problem GVar$^h$. The next step is to show that the solution of Problem GVar$^h$ converges to the solution of Problem GVar. This is proven the main result of \cite{BV13}.

Let $u$ be the solution of Problem GVar and $u^h$ be the solution of Problem
GVar$^h$. The authors of \cite{BV13} define the following error,

Let $u$ be the solution of Problem GVar and $u^h$ be the solution of Problem GVar$^h$.
 The authors of \cite{BV13} define the following error,
\begin{eqnarray}
	e^h(t) = u(t) - u^h(t). \label{Error}
\end{eqnarray}

In \cite{BV13}, a projection $P$ is defined so that $Pu \in V$. The error is
then split using this projection. Let $e(t) = Pu(t)- u_h(t)$ and $e_p(t) = u(t)
	- Pu(t)$ so that
\begin{eqnarray}
	e^h(t) = e_p(t) +e(t). \label{Error}
\end{eqnarray}

According to the authors of~\cite{BV13}, interpolation theory can be used to determine the error $e_p$. However to obtain an error estimate for $e$ is not as trivial. This is where the use of the Finite Element Analysis comes in.

To simplify the proof, the problem is divided into a semi-discrete problem and a fully discrete problem.

Consider the time interval $J = [0,T]$. Divide $J$ into $N$ steps with length $\tau = \frac{T}{N}$. Each interval can be expressed as $[t_{k-1}, \ t_k]$ for $k = 1,...,N$. Denote the approximation of $u_h$ on the interval $[t_{k-1}, \ t_k]$ by $u_k^h$, i.e. $u_h(t_k) = u_k^h$ for each $k$.

The error $e$, can now be expressed as
\begin{eqnarray}
	e(t_k) = u(t_k) - u^h_k & = & [u(t_k)-u_h(t_k)] + [u_h(t_k) - u^h_k]. \label{error_e}
\end{eqnarray}

In \eqref{error_e}, the authors divide the error $e$ into a semi-discrete term
and a fully discrete term. The two terms are investigated as seperate problems
in \cite{BV13}.\\

\subsection{Main result}

The main result proving the convergence of the solution of the Galerkin
Approximation is given in \cite{BV13} as follows.

\newtheorem*{DC_Thm5}{Theorem 2}
\begin{DC_Thm5}
	{Main Result}\\
	If $f\in C^{2}([0,T]4,X)$, then
	\begin{eqnarray*}
		\Vert u_{h}(t_{k})-u_{k}^{h}\Vert_{W}\ & \leq & 7T^{2}\tau^{2}\max\Vert u_{h}^{(4)}\Vert_{W}+7T\tau^{2}\max\Vert u_{h}'''\Vert_{W}\\
		& & + \sqrt{2K_{a}}\tau^{4}\max\Vert u_{h}'''\Vert_{W}
	\end{eqnarray*}
	for each $t\in(0,\ T)$.
\end{DC_Thm5}

The parameter $\tau$ is some number in $(0,T)$. The constant $K_\alpha$ is the
constant defined in assumption \textbf{A4} of Section 2.2.1.

\subsection{Obtaining the error estimate}
In this section we will show how the error estimate in Theorem 2 is obtained by
the authors of \cite{BV13}.\\

The following two assumptions are made in \cite{BV13}.
\subsubsection*{Assumptions}
\begin{itemize}
	\item[] \textbf{B1} - The solution $u \in C(J,V)$ of Problem GVar$^h$ has the property that $(Pu) \in C^2(J)$.

	\item[] \textbf{B2} - There exists a subspace $H$ of $V$ and a positive integer $\alpha$ such that if $w \in H$, then $$\inf_{v\in S^{h}}||w-v||_V \leq \hat{C} h^\alpha|||w|||_H,$$ where $|||w|||_H$ is a norm or a semi-norm for $H$.
\end{itemize}

The projection $P$ in section \eqref{e_est} is defined in \cite{BV13} as
\begin{eqnarray*}
	b(u-Pu,v) = 0 \ \ \ \textrm{ for all } v \in V.
\end{eqnarray*}

From assumption \textbf{B1}, $Pu \in C^2(J)$ and by the definition $P$, $P$ is
a projection with respect to the inner product $b$.\\

Subtracting \eqref{GV_PB1} from \eqref{DC_E2} gives
\begin{eqnarray*}
	c(u''(t)-u''_h(t),v) + a(u'(t)-u'_h(t),v) + b(u(t)-u_h(t),v) = 0
\end{eqnarray*}

Using the definition of the error $e^h$ while also adding and subtracting
$Pu(t)$ in the bilinear form $b$, we can write
\begin{eqnarray*}
	c(e_h''(t),v) + a(e_h'(t),v) + b(u(t) - Pu(t),v) + b(Pu(t) +u_h(t),v)- = 0.
\end{eqnarray*}

It then follows from the definitions of the error $e$ and the projection $P$
that
\begin{eqnarray*}
	c(e''(t),v) + a(e'(t),v) + b(e(t),v) = 0
\end{eqnarray*}

This equation \eqref{DC_E3} is presented in \cite{BV13} as a proposition.
\newtheorem{DC_Prop1}{Proposition}
\begin{DC_Prop1}
	If $u$ is the solution of Problem GVar and $u$ satisfies assumption \textbf{B1}, then
	\begin{eqnarray}
		c(e_h''(t),v) + a(e_h'(t),v) + b(e(t),v) = 0 \ \ \ \textrm{ for all } v \in V.
	\end{eqnarray}
\end{DC_Prop1}

Using this proposition, the authors of \cite{BV13} derive a generalized error
estimate for the error $e = Pu - u_h$.

\newtheorem{lem_1}{Lemma}
\begin{lem_1}
	If $u$ is the solution of Problem GVar and $u$ satisfies assumption \textbf{B1}, then
	\begin{eqnarray}
		\Vert e(t)\Vert_{W} &\leq& \sqrt{2} \biggl( \Vert e(0) \Vert_W + 3T\Vert e'_h(0) \Vert_W + 3 \int_o^T \Vert e'_p(t)\Vert_W\nonumber \\
		&& + 3K_\alpha T \Vert e_h(0) \Vert_W + 3K_\alpha \int_0^T \Vert e_p(t) \Vert_W \biggr). \label{fundamental_estimate}
	\end{eqnarray}
\end{lem_1}

In chapter 2, it was explained that the models of this dissertation have weak
damping, i.e. $a = 0$. Therefore, the error estimate in Lemma 1 can be
simplified. This is given in \cite{BV13} as a corollary.

\newtheorem{col_1}{Corollary}
\begin{col_1}
	If $u$ is the solution of Problem GVar with $a = 0$ and $u$ satisfies assumption \textbf{B1}, then
	\begin{eqnarray}
		\Vert e_(t)\Vert_{W} \leq \sqrt{2} \Vert e(0) \Vert_W + 2T||e'_h(0)||_W + 4\sqrt{T}\max_{t \in [0,T]} \Vert e'_p(t) \Vert_W.
	\end{eqnarray}
\end{col_1}

\begin{proof}
	Assume that there is no damping other than the weak damping, i.e. $a = 0$. The the equation in proposition 1 becomes
	\begin{eqnarray*}
		c(e_h''(t),v) + b(e(t),v) = 0 \ \ \ \textrm{ for all } v \in V.
	\end{eqnarray*}

	If $v \in S^h[0,T]$, then 
	\begin{eqnarray*}
		c(e_h''(t),v(t)) + b(e(t),v(t)) = 0.
	\end{eqnarray*}

	Let $v'(t) = c(t)$, then 
	\begin{flalign}
		\frac{d}{dt}\left[\frac{1}{2}c(e,e)-\frac{1}{2}b(v,v)-c(e_h,v) \right] &= c(e',e) - b(e,v)- c(e''_h, v)-c(e_h',v')\nonumber\\
		&= c(e'_p,e). \label{DC_E4}
	\end{flalign}

	Choose $v(\tau) = 0$ for some $\tau \in [0,T]$, and integrate \eqref{DC_E4} between $0$ and $\tau$ to obtain
	\begin{eqnarray*}
		a
	\end{eqnarray*}

\end{proof}

\subsubsection{Semi-discrete problem}
The estimate \eqref{fundamental_estimate} in Lemma 1 is equivalent to
\begin{eqnarray*}
	\Vert u(t) - u_h(t) \Vert_{W} &\leq& \sqrt{2} \biggl( \Vert u(0) - u_h(0) \Vert_W + 3T\Vert e'_h(0) \Vert_W + 3 \int_o^T \Vert e'_p(t)\Vert_W\nonumber \\
	&& + 3K_\alpha T \Vert e_h(0) \Vert_W + 3K_\alpha \int_0^T \Vert e_p(t) \Vert_W \biggr).
\end{eqnarray*}

Define the projection $P$ such that for any function $u$,
\begin{eqnarray*}
	(Pu)(t) = Pu(t) \ \ \ \textrm{ for all } t \in (0,T).
\end{eqnarray*}

From Theorem 1, it follows that the solution $u$ of Problem GVar satisfies $u
	\in C^1([0,T]; V) \cap C^2((0,T);V)$. \textcolor{red}{This is a weak assumption
	made in \cite{BV13}.}\\

\sout{The following is a weak assumption in \cite{BV13}: The solution $u$ of Problem GVar satisfies $u \in C^1([0,T]; V) \cap C^2((0,T);V)$. This follows from Theorem 1 and is a necessary assumption in [BSV17].}\\

Let $\Pi$ denote the interpolation operator such that $u_0^h = \Pi u_0$ and
$u_1^h = \Pi u_1$.

\newtheorem*{DC_Thm4}{Theorem 2}
\begin{DC_Thm4}
	Suppose Assumption E2 holds and $u_{0}^{h}=\Pi u_{0}$ and $u_{1}^{h}=\Pi u_{1}.$ If the solution $u$ of Problem $GVar$ satisfies Assumption El, $u(t)\in H$ and $u'(t)\in H$, then
	\begin{eqnarray*}
		\Vert u(t)-u_{h}(t)\Vert_{W} & \leq & C_{b} \hat{C}h^{\alpha}|||u(t)|||_{H}+\sqrt{2}C_{b}\hat{C}h^{\alpha}(3T\max|||u'(t)|||_{H}\\
		& & +3K_{a}T\max|||u(t)|||_{H}+(2+3K_{a}T)|||u_{0}|||_{H} \\
		& & +3T|||u_{1}|||_{H}) ,
	\end{eqnarray*} for each $t\in[0, T].$
\end{DC_Thm4}

\subsection{Fully discrete Galerkin finite element approximation}
Consider the time interval $J = [0,T]$. Divide $J$ into $N$ steps with length
$\tau = \frac{T}{N}$. Each interval can be expressed as $[t_{k-1}, \ t_k]$ for
$k = 1,...,N$. Denote the approximation of $u_h$ on the interval $[t_{k-1}, \
			t_k]$ by $u_k^h$, i.e. $u_h(t_k) = u_k^h$ for each $k$.\\

\sout{The goal the fully-discrete approximation as given by \cite{BV13} is to find an estimate between Problem GVar and Problem GVar$^h$, which can be expressed as}

The error $e(t)$, can now be expressed as
\begin{eqnarray}
	e(t_k) = u(t_k) - u^h_k & = & [u(t_k)-u_h(t_k)] + [u_h(t_k) - u^h_k]. \label{error_ek}
\end{eqnarray}

From Theorem 2 of the semi-discrete problem an error estimate for
$[u(t_k)-u_h(t_k)]$ with respect to the norm of $W$ was obtained. The authors
of \cite{BV13} then proceed to obtain an error estimate for $[u_h(t_k) -
			u^h_k]$. Since the dimension of $S^h$ is not fixed, the equivalence of norms
cannot be used, and therefore this error estimate for $[u_h(t_k) - u^h_k]$
should also be with respect to the norm of $W$.

\sout{In Theorem 2, Section 3.1.3 it is shown that \cite{BV13} obtained an error estimate for $[u(t_k)-u_h(t_k)]$ with respect to the norm of $W$. The authors explain that an estimate for $[u_h(t_k) - u^h_k]$ should also be with respect to the norm $W$ since the dimension of $S^h$ is not fixed and hence the equivalence of the relevant norms cannot be used.}\\

\subsubsection{Notation}
For any sequence $\{x_k\} \in R_n$,
For any sequence $\{x_k\} \in R_n$,
\begin{eqnarray*}
	\delta_{t}x_{k}& = &\tau^{-1}[x_{k+1}-x_{k}],\\
	x_{k+\frac{1}{2}} & = & \frac{1}{2}[x_{k+1}+x_{k}].
\end{eqnarray*}

Let Problem GVar$^h$-D denote the fully discrete form of Problem GVar$^h$.
\subsubsection*{Problem GVar$^\mathbf{{h}}$-D} Find a sequence
$\{u_{k}^{h}\}\subset S^{h}$ such that for $k=0,1,2$, . . . , $N-1,$
\begin{eqnarray}
	\delta_{t}u_{k}^{h}\ & = & v_{k+\frac{1}{2}},\\
	c(\delta_{t}v_{k},\ \psi)+a(v_{k+\frac{1}{2}},\ \psi)+b(u_{k+\frac{1}{2}}^{h},\ \psi) & = & \frac{1}{2}([f(t_{k})+f(t_{k+1})],\ \psi)_{X} \nonumber \\ \label{PP}
\end{eqnarray}
for each $\psi\in S^{h}$, while $u_{0}^{h}=u_{h}(0)=d^{h}$ and $v_{0}=u_{h}'(0)=v^{h}$.\\

\textcolor{red}{Haal uit************************************************}\\
The article \cite{BV13} proves the following proposition. This proposition proves that Problem GVar$^h$-D is well defined.

\newtheorem{DC_Prop4}[DC_Prop2]{Proposition}
\begin{DC_Prop4}
	Problem $GVar^{h}-D$ has a unique solution for any pair of vectors $d^{h}$ and $v^{h}$ in $S^{h}$.
\end{DC_Prop4}

In the proof of Proposition 3, the authors of \cite{BV13} show that \eqref{PP}
in Problem GVar$^h$-D can be rewritten as
\begin{eqnarray}
	c(v_k+1,\phi) + \frac{\tau}{2} a(v_{k+1},\phi) + \frac{\tau^2}{4} a(v_{k+1},\phi) & = & c(v_k,\phi) - \frac{\tau}{2} a(v_k,\phi) \nonumber \\  &&
	- \frac{\tau^2}{4}b(v_k,\phi) - \tau b(u_k,\phi) \nonumber\\
	&&+\frac{\tau}{2}([f(t_k)+f(t_{k+1})],\phi)_X.\nonumber \\ \label{PPP}
\end{eqnarray} for each $\phi \in S^h$.\\

Using \eqref{PPP} and the definition of the standard finite element matrices,
Problem GVar$^h$-D can be reformulated into a algorithm for the Finite Element
Method. This is the fully discrete form of Problem GVar and is represented in
Problem FD.\\
\textcolor{red}{******************************************************}\\

Further results and the standard Finite Element Matrices, let the authors of
\cite{BV13} rewrite Problem GVar$^\mathbf{{h}}$-D into an algorithm for the
Finite Element Method. This is represented as Problem FD

\subsubsection*{Problem FD}
Find a sequence $\{\bar{u}_{k}\}\subset R_{n}$ such that for each $k,$
\begin{eqnarray*}
	\bar{u}_{k+1}\ & = & \bar{u}_{k}+\tau\bar{v}_{k+\frac{1}{2}},\\
	(M + \frac{\tau}{2}C+\frac{\tau^{2}}{4}K)\bar{v}_{k+1} & = & (M- \frac{\tau}{2}C-\frac{\tau^{2}}{4}K)\bar{v}_{k}-\tau K\bar{u}_{k}+\frac{\tau}{2}[F(t_{k})+F(t_{k+1})]
\end{eqnarray*}
with $\bar{u}_{0}=\bar{d}$ and $\bar{v}_{0}=\bar{v}.$\\

In this form, the problem is fully discrete. This form also allow the authors
of \cite{BV13} to obtain an error estimate for the fully-discrete problem.

\subsubsection{Error estimate for the fully discrete problem}

\textcolor{red}{Haal uit*****************************}\\
There error estimate for $[u_h(t_k) - u^h_k]$ can now be determined. Recall that $\tau = \frac{T}{N}$. Subsituting $t = t_k$ and $t = t_k+1$ in Problem GVar$^h$ results in
\begin{eqnarray*}
	c( \tau^{-1}[v_{h}(t_{k+1})-v_{h}(t_{k})], \psi)+\frac{1}{2}a([v_{h}(t_{k+1})+v_{h}(t_{k})], \psi)+\frac{1}{2}b([u_{h}(t_{k+1})+u_{h}(t_{k})], \psi) \\
	=\frac{1}{2}([f(t_{k+1})+f(t_{k})], \psi)_{X}+c(\rho_{k}, \psi)
\end{eqnarray*}
where $v_{h}(t)=u_{h}'(t)$ and
\begin{eqnarray*}
	\rho_{k}=\tau^{-1}[v_{h}(t_{k+1})-v_{h}(t_{k})]-\frac{1}{2}[v_{h}'(t_{k+1})+v_{h}'(t_{k})].
\end{eqnarray*}

The authors in \cite{BV13} denote the following errors
\begin{eqnarray*}
	e_{k}=u_{h}(t_{k})-u_{k} \ \ \ \ \textrm{ and } \ \ \ \ q_{k}=u_{h}'(t_{k})-v_{k}.
\end{eqnarray*} In the authors approach, $e_0$ and $q_0$ are zero.

\newtheorem*{DC_Lem3}{Lemma 2}
\begin{DC_Lem3}
	Stability
	\begin{eqnarray*}
		\max\Vert e_{n}\Vert_{W}^{2}\leq 8T\tau\sum_{n=0}^{N-1}\Vert\epsilon_{n}\Vert_{W}^{2}+2\tau^{4}\Vert\rho_{0}\Vert_{W}^{2}+(8\tau^{2}+2\tau^{4}K_{a})\Vert\sigma_{0}\Vert_{W}^{2}.
	\end{eqnarray*}
\end{DC_Lem3}

Using this lemma, \cite{BV13} proves the following error estimate.\\
\textcolor{red}{*************************************}

Recall that $\displaystyle \tau = \frac{T}{N}$. The error estimate for the term
$[u_h(t_k) - u^h_k]$ with respect to the norm of $W$ as proven by the authors
of \cite{BV13} is presented here as Theorem 3.

Finally, Theorem 2 of the semi-discrete problem and Theorem 3 of the fully
discrete problem gives an error estimate for the error $e(t)$. Consequently,
the error estimate $e^h(t) = u(t) - u^h(t)$ is obtained.

The parameter $\tau$ is some number in $(0,T)$. The constant $K_\alpha$ is the constant defined in assumption \textbf{A4} of Section 2.2.1.

\end{document}

Define the error $e_h$ as the error between the solutions of Problem GVar and
Problem GVar$^h$, i.e.
\begin{eqnarray}
	e_h(t) = u(t)-u_h(t). \label{Error}
\end{eqnarray}

Since $u$ is unknown, the error $e_h$ cannot be calculated directly. Instead
the authors of \cite{BV13} derive estimate for this error using projections.***
\\

Define a projection operator $P_h$ by
\begin{eqnarray*}
	b(u - P_hu, v) = 0 \ \ \ \textrm{ for all } v \in S^h.
\end{eqnarray*}

Define the projection $P$ such that for any function $u$,
\begin{eqnarray*}
	(Pu)(t) = Pu(t) \ \ \ \textrm{ for all } t \in (0,T).
\end{eqnarray*}

Same a \cite{BV13}, $P_h$ is renamed to $P$ for convenience. $P_h$ will be used
when it is needed to differentiate between $P$ and $P_h$. Define the following
errors $e_p$ and $e$ by
\begin{eqnarray*}
	e(t) & = & Pu(t) - u_h(t),\\
	e_p(t) & = & u(t) - Pu(t).
\end{eqnarray*}
Then \eqref{Error} can be rewritten as
\begin{eqnarray*}
	e_h(t) = e(t) - e_p(t).
\end{eqnarray*}

Estimates for the norm of $e_p$ is obtained in \cite{BV13} using approximation
theory. It is only left to find an estimate for $e(t)$. The approach of
\cite{BV13} is to estimate the error $e$ using the projection error $e_p$ and
errors for the initial conditions.

\subsubsection*{Assumptions}
The following assumptions are made in \cite{BV13} for the convergence results.
\begin{itemize}
	\item[] \textbf{E1} - The solution $u$ of Problem GVar has the property that $(Pu) \in C^2(0,T)$.

	\item[] \textbf{E2} - The solution $u$ of Problem GVar satisfies $u \in C^1([0,T]; V) \cap C^2((0,T);V)$.
\end{itemize}

\subsubsection{Remark}
- Assumption \textbf{E2} is a weak assumption and follows from the existence theorem, Theorem 1 of Section 2.2. In [BSV17], \textbf{E2} is no longer a weak assumption but a necessary assumption. *****\\

From this Proposition in \cite{BV13} the general variational problem in terms
of the errors $e_h$ and $e$. \newtheorem{DC_Prop1}{Proposition}
\begin{DC_Prop1}
	If $u$ is a solution of Problem GVar and satisfies Assumption E1, then
	\begin{eqnarray*}
		c(e_h''(t),v)+a(e_h'(t),v)+b(e_h(t),v) = 0 \ \ \ \ \textrm{for each} \ v \in S^h.
	\end{eqnarray*}
\end{DC_Prop1}

\subsection{Fundamental estimate}
The authors of \cite{BV13} use the form of Proposition 1, to give a
generalization of an estimate for the error $e(t)= Pu(t)- u_h(t)$.

\newtheorem{DC_Lem2}{Lemma}
\begin{DC_Lem2}
	If the solution $u$ of Problem $GVAr$ satisfies Assumption El, then for $t\in[0,\ T],$
	\begin{eqnarray*}
		\Vert e(t)\Vert_{W}\ &\leq\ & \sqrt{2}(\Vert e(0)\Vert_{W}+3T\Vert e_{h}'(0)\Vert_{W}+3\int_{0}^{T}\Vert e_{p}'(t)\Vert_{W}\\
		& & +3K_{a}T\Vert e_{h}(0)\Vert_{W}+3K_{a}\int_{0}^{T}\Vert e_{p}(t)\Vert_{W})
	\end{eqnarray*}
\end{DC_Lem2}

Similarly, for $a = 0$ (weak damping as explained in Section 2.2), the
following corollary is obtained in a similar proof to Lemma 1.
\newtheorem{DC_Cor1}{Corollary}
\begin{DC_Cor1}
	If the solution $u$ of Problem $GVar$ with $a=0$ satisfies Assumption $El$, then for $t\in[0, T],$
	\begin{eqnarray*}
		\Vert e(t)\Vert_{W} & \leq  & \sqrt{2}\Vert e(0) \Vert_W + 2T\Vert e_h'(0) \Vert_W + 4 \sqrt{T} \max\limits_{t \in [0,T]} \Vert e_p'(t) \Vert_W.
	\end{eqnarray*}
\end{DC_Cor1}

\subsection{Convergence and error estimated for the semi-discrete approximation}
Consider Problem GVar$^h$ with discretised space and continuous time variables.
This is known as a semi-discrete problem. This is a natural intermediate step
between a continuous and a fully discrete problem.\\

\newtheorem*{DC_Thm3}{Theorem 2}
\begin{DC_Thm3}
	If the solution $u$ of Problem $G$ satisfies Assumption Al, then
	\begin{eqnarray}
		\Vert u(t)-u_{h}(t)\Vert_{W} & \leq & \Vert e_{p}(t)\Vert_{W}+\sqrt{2}(\Vert Pu_{0}-u_{0}\Vert_{W}+3T\Vert u_{1}-u_{1}^{h}\Vert_{W}+3\int_{0}^{T}\Vert e_{p}'(t)\Vert_{W} \nonumber \\
		& & +(1+3K_{a}T) \Vert u_{0}-u_{0}^{h}\Vert_{W}+3K_{a}\int_{0}^{T}\Vert e_{p}(t)\Vert_{W}) \label{DC_E7}
	\end{eqnarray}
	for each $t\in[0,\ T].$
\end{DC_Thm3}
\begin{comment}
\begin{proof}
	From (3.2), $\Vert u(t)-u_{h}(t)\Vert_{W}\leq\Vert e_{p}(t)\Vert_{W}+\Vert e(t)\Vert_{W}$.\\

	The result follows from Lemma 4.1 since $\Vert
		Pu_{0}-u_{0}^{h}\Vert_{W}\leq\Vert Pu_{0}-u_{0}\Vert+\Vert
		u_{0}-u_{0}^{h}\Vert.$
\end{proof}
\end{comment}

To be able to use Theorem in application, it is necessary to be able to
estimate an element in $V$ with an element in $S^h$. Suppose $h$ is related to
the dimension n of $S^h$ such that $h \rightarrow 0$ as $n \rightarrow
	\infty$.\\

Another assumption is introduced by \cite{BV13}.

\subsubsection*{Assumption E3}
There exists a subspace $H$ of $V$ and a positive integer $\alpha$ such that if $w\in H,$ then
\begin{eqnarray*}
	\inf_{v\in S^{h}}\Vert w-v\Vert_{V}\leq\hat{C}h^{\alpha}|||w|||_{H},
\end{eqnarray*}
where $|||w|||_{H}$ is a norm or semi-norm for $H.$\\

Using this assumption, the projection errors in Theorem 2 can be expressed as
\begin{eqnarray*}
	||e'_p(t)||_W & \leq & C_b \hat{C} h^\alpha |||u'(t)|||_H,\\
	||u_0 - u_0^h||_W & \leq & C_b \hat{C} h^\alpha |||u_0|||_H,
\end{eqnarray*} if $u^h_0$ and $u^h_1$ are chosen as $Pu_0$ and $Pu_1$. However choosing $u^h_0$ and $u^h_1$ in this form is not possible in general. Let $\Pi$ denote the interpolation operator such that $u_0^h = \Pi u_0$ and $u_1^h = \Pi u_1$.

\newtheorem*{DC_Thm4}{Theorem 3}
\begin{DC_Thm4}
	Suppose Assumption E3 holds and $u_{0}^{h}=\Pi u_{0}$ and $u_{1}^{h}=\Pi u_{1}.$ If the solution $u$ of Problem $GVar$ satisfies Assumption El, $u(t)\in H$ and $u'(t)\in H$, then
	\begin{eqnarray*}
		\Vert u(t)-u_{h}(t)\Vert_{W} & \leq & C_{b} \hat{C}h^{\alpha}|||u(t)|||_{H}+\sqrt{2}C_{b}\hat{C}h^{\alpha}(3T\max|||u'(t)|||_{H}\\
		& & +3K_{a}T\max|||u(t)|||_{H}+(2+3K_{a}T)|||u_{0}|||_{H} \\
		& & +3T|||u_{1}|||_{H}) ,
	\end{eqnarray*} for each $t\in[0, T].$
\end{DC_Thm4}