### 2023-09-30 files/Chapter6/Validity_Timo.tex
chapter 6v2

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Validity of a Cantilever Timoshenko beam} \label{sec:validity-of-a-cantilever-timoshenko-beam}

In this section, the validity of the Timoshenko beam model is investigated. Specifically, the validity of a cantilever Timoshenko beam model, using a cantilever two-dimensional beam model as a reference. This is a discussion and a replication of numerical results of the work done in the article \cite{LVV09}.

The article \cite{LVV09} is titled `Comparison of linear beam theories'. In this article, the authors compare a cantilever Timoshenko beam to a cantilever two-dimensional beam. \textcolor{red}{Give a brief summary of the article here, linking to whatever has already been discussed in the literature review.}

\subsection{The models}
The following two models are the focus of this section and the models that are used in the article \cite{LVV09}. From section \ref{ssec:1D_Model:ModelProblems}, the cantilever Timoshenko beam model, referred to as Problem T-2. From section \ref{ssec:2D_Model:ModelProblem}, the cantilever two-dimensional beam model, referred to as Problem 2D-1.\\

Figure \ref{fig:compare:1D+2D} shows the two beams side-by-side.

\begin{figure}[h!]
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\caption{Side by side comparison of the beams.}
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\begin{center}
					\begin{tikzpicture}
					\draw[line width = 0.4mm] (0,0) -- (6,0);

					\node at (2.7,-0.25) {$\ell = 1$};

					\draw[line width = 0.1mm] (0,-1.5) -- (0,1.5);
					\draw[line width = 0.1mm] (0,1.5) -- (-0.2,1.4);
					\draw[line width = 0.1mm] (0,1.3) -- (-0.2,1.2);
					\draw[line width = 0.1mm] (0,1.1) -- (-0.2,1);
					\draw[line width = 0.1mm] (0,0.9) -- (-0.2,0.8);
					\draw[line width = 0.1mm] (0,0.7) -- (-0.2,0.6);
					\draw[line width = 0.1mm] (0,0.5) -- (-0.2,0.4);
					\draw[line width = 0.1mm] (0,0.3) -- (-0.2,0.2);
					\draw[line width = 0.1mm] (0,0.1) -- (-0.2,0);

					\draw[line width = 0.1mm] (0,-0.1) -- (-0.2,-0.2);
					\draw[line width = 0.1mm] (0,-0.3) -- (-0.2,-0.4);
					\draw[line width = 0.1mm] (0,-0.5) -- (-0.2,-0.6);
					\draw[line width = 0.1mm] (0,-0.7) -- (-0.2,-0.8);
					\draw[line width = 0.1mm] (0,-0.9) -- (-0.2,-1);
					\draw[line width = 0.1mm] (0,-1.1) -- (-0.2,-1.2);
					\draw[line width = 0.1mm] (0,-1.3) -- (-0.2,-1.4);
					\draw[line width = 0.1mm] (0,-1.5) -- (-0.2,-1.6);

					%\node at (3.2,-0.2) {$\ell = 1$};
					\end{tikzpicture}
				\end{center}
				\subcaption{Timoshenko Cantilever Beam}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\begin{center}
					\begin{tikzpicture}
						\draw[line width = 0.4mm] (0,0.2) -- (6,0.2);
						\draw[line width = 0.4mm] (0,-0.2) -- (6,-0.2);
						\draw[line width = 0.4mm] (6,-0.2) -- (6,0.2);
						%\draw[line width = 0.4mm] (0,-0.2) -- (0,0.2);

						%\draw[line width = 0.1mm,->] (6,2) -- (7.4,2);
						%\draw[line width = 0.1mm,->] (6,2) -- (6,0.9);
						%\node at (6.7,1.8) {$e_1$};
						%\node at (5.8,1.4) {$e_2$};

						\node at (6.2,0) {$h$};
						\node at (2.7,-0.45) {$\ell = 1$};

						\draw[line width = 0.1mm] (0,-1.5) -- (0,1.5);
						\draw[line width = 0.1mm] (0,1.5) -- (-0.2,1.4);
						\draw[line width = 0.1mm] (0,1.3) -- (-0.2,1.2);
						\draw[line width = 0.1mm] (0,1.1) -- (-0.2,1);
						\draw[line width = 0.1mm] (0,0.9) -- (-0.2,0.8);
						\draw[line width = 0.1mm] (0,0.7) -- (-0.2,0.6);
						\draw[line width = 0.1mm] (0,0.5) -- (-0.2,0.4);
						\draw[line width = 0.1mm] (0,0.3) -- (-0.2,0.2);
						\draw[line width = 0.1mm] (0,0.1) -- (-0.2,0);

						\draw[line width = 0.1mm] (0,-0.1) -- (-0.2,-0.2);
						\draw[line width = 0.1mm] (0,-0.3) -- (-0.2,-0.4);
						\draw[line width = 0.1mm] (0,-0.5) -- (-0.2,-0.6);
						\draw[line width = 0.1mm] (0,-0.7) -- (-0.2,-0.8);
						\draw[line width = 0.1mm] (0,-0.9) -- (-0.2,-1);
						\draw[line width = 0.1mm] (0,-1.1) -- (-0.2,-1.2);
						\draw[line width = 0.1mm] (0,-1.3) -- (-0.2,-1.4);
						\draw[line width = 0.1mm] (0,-1.5) -- (-0.2,-1.6);

						%\node at (3.2,-0.3) {$\ell = 1$};

					\end{tikzpicture}
				\end{center}
				\subcaption{Two-Dimensional Cantilever Beam}
			\end{minipage}
		}
	}
\end{figure}\label{fig:compare:1D+2D}

As mentioned in the introduction, the height of the two beams can be set to be the same by considering the parameter $\alpha$ for the Timoshenko beam model and the parameter $h$ for the two-dimensional beam model.

\subsection{Calculating the eigenvalues and eigenvectors}
In section \ref{sec:Timo:EigenvalueProblem} a method for calculating the eigenvalues and eigenvectors of the Timoshenko beam is provided. And in section \ref{sec:Timo:Cantilever}, the eigenvalues and eigenvectors of a cantilever Timoshenko beam are calculated as an example. These can also be obtained to a desired level of accuracy.

For the two-dimensional beam, the eigenvalues and eigenvectors are calculated using the Finite Element Method. In section \ref{sec:FEM:2D}, the Finite Element Method for a cantilever two-dimensional beam is derived. Specifically, section \ref{2dFEM_EP} derived the eigenvalue problem for the two-dimensional beam using the Finite Element Method, called Problem 2D-1E.

\subsubsection{Problem 2D-1E}
Find a real number $\lambda$ and a function $\bar{u} \in S^h$ such that
\begin{eqnarray}
	K\bar{u} & = & M\lambda{\bar{u}},
\end{eqnarray} where $M$ and $K$ are the standard Finite Element Method matrices defined in section \ref{sec:FEM:2D}.

In this form, the eigenvalue problem is a system of ordinary differential equations, and can be solved using a numerical method.

\subsubsection{Accuracy of the eigenvalues}
Solving \textbf{Problem 2D-1E} with a numerical method, requires a quick investigation into the rate of convergence so that the accuracy of the eigenvalues can be established. Chapter \ref{ch:finite-element-theory} provides the necessary theory on the existence of solutions as well as the proof of convergence of the Finite Element Method.

\begin{figure}[H]
    \centering
    \begin{adjustbox}{center}
        \includegraphics[scale=0.7]{Convergence.png}
    \end{adjustbox}
    \caption{Rate of convergence of the first 20 eigenvalues.}
    \label{fig:conv_2d_eig}
\end{figure}

Figure \ref{fig:conv_2d_eig} shows the rate of convergence of the first 20 eigenvalues of the two-dimensional beam. Each color represents a spesific eigenvalue.

The number of elements are chosen so that at least the first 10 eigenvalues are accurate to 5 significant digits.

\subsection{Comparing the mode shapes}
Recall from the introduction, that the investigation is focussed on beam-type problems. The two-dimensional model is for an elastic body, and not specific to a beam, like the Timoshenko beam model. Therefore there are non-beam type behaviour that can be expected from the two-dimensional model. This is also be reflected in the eigenvalues of the two-dimensional model.

Therefore a method is required to compare and match the eigenvalues of the two models. In \cite{LVV09}, the authors compare the mode shapes of the two models, to match up the eigenvalues. Recall \textcolor{red}{****} that eigenvalues relating to the mode shapes similar to the mode shapes of the Timoshenko beam, are called beam type eigenvalues. The other eigenvalues are called non-beam type eigenvalues. This naming convention is used by the authors of \cite{LVV09}.

\subsubsection{Shapes relating to beam-type eigenvalues}
The following figures are examples of beam-type mode shapes for the displacement $w$.

\begin{figure}[h!]
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{Beam1.jpg}
				\subcaption{2D Beam Type - $\lambda_6 = 21.911$}
				\label{fig:minipage2}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{1DBeam2.png}
				\subcaption{Timoshenko - $\lambda_5 = 21.794$}
				\label{fig:minipage1}
			\end{minipage}
	}}
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{Beam2.jpg}
				\subcaption{2D Beam Type - $\lambda_7 = 45.711$}
				\label{fig:minipage2}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{1DBeam1.png}
				\subcaption{Timoshenko - $\lambda_6 = 45.390$}
				\label{fig:minipage1}
			\end{minipage}
			\caption{Modal shapes of the displacement $w$ for the beam-type 2D body and the Timoshenko beam with $\alpha = 4800$ ($h = 1/20$).}
		}
	}
\end{figure}
\FloatBarrier

\subsubsection{Shapes relating to non-beam type eigenvalues}
The following figures are examples of non beam-type mode shapes for the displacement $w$. These mode shapes are not present in the cantilever Timoshenko beam model and are not beam related. 
\begin{figure}[h!]
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{NotBeam1.jpg}
				\subcaption{Non-Beam Type - $\lambda_4 = 7.7077$}
				\label{fig:minipage2}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{NotBeam2.jpg}
				\subcaption{Non-Beam Type - $\lambda_8 = 69.344$}
				\label{fig:minipage1}
			\end{minipage}
			\caption{Modal shapes of the displacement $w$ for the non-beam type 2D body with $h = 1/20$.}
		}
	}
\end{figure}
\FloatBarrier

\subsubsection{Shape of the cross-sections}
The Timoshenko beam theory improve some one-dimensional beam theories such as the Euler-Bernoulli beam theory by also including the effect of shear. For Timoshenko beam theory it is assumed that the cross-sections need not remain perpendicular to the neutral axis of the beam. The cross-sections however remain a straight line.\\

For the two-dimensional beam, the shape of the cross-section can deform into a s-shape. This is explained in \cite{LVV09} and a similar figure in \cite{LVV09} is given here. This is a difference between the two models that are inherent to their definitions.\\

\FloatBarrier
\begin{figure}[h!]
	\centering
	\begin{tikzpicture}[scale=1.2]

	\draw[line width = 0.4mm] (0,1) -- (7,1);
	\draw[line width = 0.4mm] (0,-1) -- (7,-1);
	\draw[line width = 0.4mm] (7,-1) -- (7,1);
	\draw[line width = 0.4mm] (0,-1) -- (0,1);

	\draw[line width = 0.1mm,dashed] (3.5,1) -- (3.5,-1);

	\draw[thick] plot [smooth,tension=0.9] coordinates{(3.3,1) (3.7,0.5) (3.5,0) (3.3,-0.5) (3.7,-1)};

	\draw[line width = 0.2mm,red] (3.3,1) -- (3.7,-1);

\end{tikzpicture}
	\caption{S-shape deformation of the cross-section of the two-dimensional beam.}\label{fig:s}
\end{figure}
\FloatBarrier

Note that figure \ref{fig:s} shows an exaggerated example of the deformation of a cross-section. The red line shows how the authors of \cite{LVV09} calculated the average rotation of a cross-section of the two-dimensional beam. This average rotation can be used to compare the rotation of the cross-section of the two-dimensional beam to the rotation of the cross-section of the Timoshenko beam given by $\phi$.

\subsubsection{Direct comparison of mode shapes}
Figure \ref{fig:w} directly compares a mode shape of the Timoshenko beam to a mode shape of the two-dimensional beam. For the two-dimensional model, the center line of the displacement of the beam is shown. For the Timoshenko beam, the displacement $w$ is shown.

\begin{figure}[ht!]
	\scalebox{.8}{
		\makebox[\textwidth]{
			\centering
			\centering
			\includegraphics[scale=0.4]{1Dv2D.png}
			\caption{Comparison of the displacement $w$ mode shape corresponding to $\lambda_{10}$ for the 2D beam and $\lambda_8$ for the Timoshenko beam with $\alpha = 4800$ ($h=1/20$)}
			\label{fig:w}}}
\end{figure}
\FloatBarrier

Similarly, figure {fig:phi} directly compares the angle of the cross-section of the Timoshenko beam and the two-dimensional beam. The average rotation of the cross-section of the two-dimensional beam's mode shape is calculated as shown in figure \ref{fig:s}.

\FloatBarrier
\begin{figure}[ht!]
	\scalebox{.8}{
		\makebox[\textwidth]{
			\centering
			\centering
			\includegraphics[scale=0.4]{1Dv2Dphi.png}
			\caption{Comparison of the angle $\phi$ (best fit for 2D Beam) mode shape corresponding to $\lambda_{10}$ for the 2D beam and $\lambda_8$ for the Timoshenko beam with $\alpha = 4800$ ($h=1/20$)}
			\label{fig:phi}}}
\end{figure}
\FloatBarrier

These figures are examples that show how similar the mode shapes of the two models are. This specific example is for $\alpha = 4800$, which represents a long, narrow beam. The authors of \cite{LVV09} obtained similar results for $\alpha=300$, which represents a short and fat beam.

\subsubsection{Remark:} Note that the overall shape is more important than the actual fit. This is because any multiple of a eigenvector is still an eigenvector. The mode shapes were specifically scaled to obtain figures \ref{fig:w} and \ref{fig:phi}.

\subsection{Comparing the eigenvalues}
Using this method of comparing the mode shapes, the eigenvalues can now be matched up and compared.

\subsubsection{Some results from [LVV09]}
The following table contains results from \cite{LVV09} verbatim to 3 significant digits, as well as the replicated results to 5 significant digits.

\begin{table}[!ht]
	\centering
	\caption{Results from [LVV09] and results obtained in this dissertation. $0^*$ indicates a 0 as a result of rounding. $\alpha = 1200$.}
	\begin{tabular}{|c|c|c|c||c|c|c|c|}
		\hline
		\multicolumn{4}{|c||}{Results from [LVV09]} & \multicolumn{4}{c|}{Dissertation} \\ \hline \hline
		~ & 2D & Timo & RE & ~ & 2D & Timo & RE  \\ \hline
		1 & 0.0317 & 0.0316 & 0.00315 & 1 & 0.031713 & 0.031639 & 0.0023407  \\ 
		2 & 1.14 & 1.14 & $0^*$ & 2 & 1.1413 & 1.1365 &  0.0042050 \\ 
		3 & 7.72 & - & ~ & 3 & 7.7161 & - &   \\ 
		4 & 7.92 & 7.86 & 0.00758 & 4 & 7.918 & 7.8617 & 0.0071116    \\ 
		5 & 26.2 & 25.9 & 0.0115 & 5 & 26.148 & 25.869 & 0.010669  \\ 
		6 & 60.8 & 59.9 & 0.0148 & 6 & 60.816 & 59.946 & 0.014303  \\ 
		7 & 69.3 & - & ~ & 7 & 69.344 & - &    \\ 
		8 & 115 & 113 & 0.0174 & 8 & 115.28 & 113.23 &  0.017787  \\ 
		9 & 192 & 188 & 0.0208 & 9 & 191.57 & 187.55 &  0.020999  \\ 
		10 & 192 & - &  & 10 & 192.03 & - &   \\ 
		11 & 291 & 284 & 0.0241 & 11 & 290.76 & 283.81 &  0.023899  \\ \hline
	\end{tabular}\label{Results_LVV09}
\end{table}

This table shows that the results obtained in this dissertation are very similar to the results obtained in \cite{LVV09}. This is for a specific case where $\alpha = 1200$.

In the following table, the eigenvalues for different values of $\alpha$ are compared.

\begin{table}[h!]
	\scalebox{.8}{
	\makebox[\textwidth]{
	\caption{Eigenvalues of a Timoshenko cantilever beam vs the eigenvalues of a cantilever two-dimensional elastic body. *RE is the relative error.}
	\begin{tabular}{|cccc||cccc||cccc||cccc|}
		\hline
		\multicolumn{16}{|c|}{Comparison of Eigenvalues} \\
		\hline\hline
		\multicolumn{4}{|c||}{ $h = 1/5$ or $\alpha = 300$}       & \multicolumn{4}{c||}{$h =1/10$ or $\alpha = 1200$}      & \multicolumn{4}{c||}{$h = 1/20$ or $\alpha = 4800$}      & \multicolumn{4}{c|}{$h = 1/30$ or $\alpha = 10800$} \\
		\hline
		{i} & {2D} & {j} & {Timo} & {i} & {2D} & {j} & {Timo} & {i} & {2D} & {j} & {Timo} & {i} & {2D} & {j} & {Timo} \\
		\hline
		{1}&0.12151&1&0.12092&1&0.031713&1&0.031639&1&0.008013&1&0.008004&1&0.003568&1&0.003565\\
		{2}&3.5460&2&3.5071&2&1.1413&2     & 1.1365 & 2     & 0.30756 & 2     & 0.30705 & 2     & 0.13869 & 2     & 0.13855 \\
		\cellcolor{lightgray}{3} & \cellcolor{lightgray}{7.7311} &       & {-} & \cellcolor{lightgray}{3}     & \cellcolor{lightgray}{7.7161} &       & {-} & 3     & 2.3273 & 3     & 2.3213 & 3     & 1.0698 & 3     & 1.0683 \\
		{4} & 20.225 & 3     & 19.869 & 4     & 7.9180 & 3     & 7.8617 & \cellcolor{lightgray}{4}     & \cellcolor{lightgray}{7.7077} &       & {-} & 4     & 4.0140 & 4     & 4.0058 \\
		{5} & 56.109 & 4     & 54.766 & 5     & 26.148 & 4     & 25.869 & 5     & 8.5086 & 4     & 8.4762 & \cellcolor{lightgray}{5}     & \cellcolor{lightgray}{7.7047} &       & {-} \\
		\cellcolor{lightgray}{6} & \cellcolor{lightgray}{69.164} &       & {-} & 6     & 60.816 & 5     & 59.946 & 6     & 21.911 & 5     & 21.794 & 6     & 10.655 & 5     & 10.625 \\
		{7} & 114.03 & 5     & 110.75 & \cellcolor{lightgray}{7}     & \cellcolor{lightgray}{69.344} &       & {-} & 7     & 45.711 & 6     & 45.390 & 7     & 22.975 & 6     & 22.890 \\
		{8} & {189.17} &  6     & {186.50} & 8     & 115.28 & 6     & 113.23 & \cellcolor{lightgray}{8}     & \cellcolor{lightgray}{69.344} &       & {-} & 8     & 43.113 & 7     & 42.909 \\
		\cellcolor{lightgray}{9} & \cellcolor{lightgray}{192.61} &      &  & {9}     & {191.57} &   7  & {187.55} & 9     & 82.887 & 7     & 82.154 & \cellcolor{lightgray}{9}     & \cellcolor{lightgray}{69.331} &       & {-} \\
		{10} & 285.85 & 7     & 277.64 & \cellcolor{lightgray}{10}    & \cellcolor{lightgray}{192.03} &      & {-} & 10    & 136.03 & 8     & 134.58 & 10    & 73.230 & 8     & 72.803 \\
		{11} & 328.40 & 8     & 330.29 & 11    & 290.76 & 8     & 283.81 & \cellcolor{lightgray}{11}    & \cellcolor{lightgray}{192.48} &       & {-} & 11    & 115.41 & 9     & 114.61 \\
		\cellcolor{lightgray}{12} & \cellcolor{lightgray}{357.08} &       & {-} & \cellcolor{lightgray}{12}    & \cellcolor{lightgray}{374.45} &       & {-} & 12    & 207.29 & 9     & 204.69 & 12    & 171.61 & 10    & 170.20 \\
		{13} & 397.33 & 9     & 394.02 & 13    & 413.20 & 9     & 402.27 & 13    & 298.38 & 10    & 294.10 & \cellcolor{lightgray}{13}    & \cellcolor{lightgray}{192.52} &       & {-} \\
		{14} & 442.00   & 10    & 439.52 & 14    & 558.67 & 10    & 542.65 & \cellcolor{lightgray}{14}    & \cellcolor{lightgray}{376.83} &       & {-} & 14    & 243.56 & 11    & 241.26 \\
		\cellcolor{lightgray}{15} & \cellcolor{lightgray}{533.71} &       & {-} & \cellcolor{lightgray}{15}    & \cellcolor{lightgray}{614.11} &       & {-} & 15    & 410.63 & 11    & 404.01 & 15    & 332.83 & 12    & 329.28 \\
		{16} & 538.97 & 11    & 541.55 & 16    & 726.26 & 11    & 704.15 & 16    & 545.03 & 12    & 535.32 & \cellcolor{lightgray}{16}    & \cellcolor{lightgray}{377.16} &       & {-} \\
		\cellcolor{lightgray}{17} & \cellcolor{lightgray}{596.06} &       & {-} & \cellcolor{lightgray}{17}    & \cellcolor{lightgray}{906.28} &       & {-} & \cellcolor{lightgray}{17}    & \cellcolor{lightgray}{621.95} &       & {-} & 17    & 440.77 & 13    & 435.51 \\
		{18} & 602.77 & 12    & 596.09 & 18    & 913.69 & 12    & 884.92 & 18    & 702.30 & 13    & 688.64 & 18    & 568.51 & 14    & 561.04 \\
		\cellcolor{lightgray}{19} & \cellcolor{lightgray}{657.87} &       & {-} & 19    & 1113.7 & 13    & 1080.1 & 19    & 882.95 & 14    & 864.40 & \cellcolor{lightgray}{19}    & \cellcolor{lightgray}{623.05} &       & {-} \\
		{20} & 717.37 & 13    & 731.74 & \cellcolor{lightgray}{20}    & \cellcolor{lightgray}{1218.0}  &       & {-} & \cellcolor{lightgray}{20}    & \cellcolor{lightgray}{927.18} &       & {-} & 20    & 717.04 & 15    & 706.74 \\
		\hline
		\multicolumn{2}{|c}{Max RE:} & \multicolumn{2}{c||}{3.1718\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c||}{3.1486\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c||}{2.1018\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c|}{1.4361\%} \\
		\hline
	\end{tabular}%
	\label{tab:Timo}%
}
}
\end{table}%
\FloatBarrier

This table shows that the eigenvalues compare very well. The non-beam type eigenvalues are highlighted in grey. For a short fat beam ($\alpha = 300$), the maximum relative error for the first 20 two-dimensional eigenvalues is just over $3\%$, while for a long thin beam ($\alpha = 10800$), the maximum relative error is just over $1\%$. This shows that as the beam becomes longer and thinner, the Timoshenko beam is better approximation of the two-dimensional beam. But overall the Timoshenko beam compares very well.

This table also shows that as the beam gets more narrow, there are less non-beam type eigenvalues within the first few eigenvalues.This would also indicate that the Timoshenko beam would be a better approximation of the two-dimensional beam as the beam gets more narrow since the two-dimensional model behaves `more like a beam'.

\begin{figure}
    \centering
    \begin{tikzpicture}
        \begin{axis}[
            title={Relative Error of Eigenvalues for Different Values of Alpha},
            xlabel={Eigenvalue},
            ylabel={Relative Error (\%)},
            legend pos=north east,
            grid=major,
        ]
        % Alpha = 300
        \addplot coordinates {
            (1,0.4875) (2,1.1106) (3,0) (4,1.7895) (5,2.4560) (6,0) (7,2.9630) (8,1.4324) (9,0) (10,2.9530)
            (11,0.5735) (12,0) (13,0.8375) (14,0.5643) (15,0) (16,0.4757) (17,0) (18,1.1240) (19,0) (20,1.9740)
        };
        \addlegendentry{\(\alpha = 300\)}

        % Alpha = 1200
        \addplot coordinates {
            (1,0.2337) (2,0.4217) (3,0) (4,0.7198) (5,1.0738) (6,1.4507) (7,0) (8,1.6244) (9,2.1618) (10,0)
            (11,2.4560) (12,0) (13,2.7265) (14,0) (15,0) (16,5.2392) (17,0) (18,3.2823) (19,3.0864) (20,0)
        };
        \addlegendentry{\(\alpha = 1200\)}

        % Alpha = 4800
        \addplot coordinates {
            (1,0.1125) (2,0.1661) (3,0.2574) (4,0) (5,0.3824) (6,0.5373) (7,0.7065) (8,0) (9,0.9950) (10,1.0740)
            (11,0) (12,1.2680) (13,1.4507) (14,0) (15,0) (16,1.8120) (17,0) (18,1.9850) (19,2.1618) (20,0)
        };
        \addlegendentry{\(\alpha = 4800\)}

        % Alpha = 10800
        \addplot coordinates {
            (1,0.0841) (2,0.1009) (3,0.1287) (4,0.2052) (5,0.3824) (6,0.2817) (7,0.3725) (8,0.4712) (9,0) (10,0.5940)
            (11,0.7005) (12,0.8140) (13,0.9307) (14,0) (15,0) (16,1.8120) (17,1.2150) (18,1.9850) (19,2.1618) (20,2.8860)
        };
        \addlegendentry{\(\alpha = 10800\)}

        \end{axis}
    \end{tikzpicture}
    \caption{Plot of the relative error for each eigenvalue and each value of \(\alpha\).}
    \label{fig:relative_error_plot}
\end{figure}

\end{document}