### 2023-10-02 files/Chapter6/Validity_Plate.tex
chapter 6 finished excluding new data that is still needed

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Validity of a Reissner-Mindlin plate} \label{sec:validity-of-a-plate}

In this section, the validity of a cantilever Reisner-Mindlin plate model is investigated. The idea of the model comes from the beam models in the previous section failing to approximate a cantilever elastic body with a width much larger than it's height. The process of determining the validity of the model follows the same ideas and methodology as in the previous sections, that is established in \cite{LVV09}.

Similarly to how the Timoshenko beam theory improves on the Euler-Bernoulli beam theory by including the shear forces, so does the Reissner-Mindlin plate models improve on classic plate models. The Reissner-Mindlin plate model is a two-dimensional model, and therefore is it of value to investigate the validity of the model using a three-dimensional plate model as a reference.

\subsection{The models}
Form section \ref{ssec:P_Model:ModelProblems}, the cantilever Reissner-Mindlin plate model is given, and is referred to as Problem P-1. The three-dimensional model is the same as used in the previous section, and is referred to as Problem 3D-1 defined in section \ref{ssec:3D_Model:ModelProblems}.

Figure \ref{fig:plate_sbs} shows the two cantilever plates side by side.

\FloatBarrier
\begin{figure}[h!]
	\scalebox{.7}{
		\makebox[\textwidth][c]{
			\caption{Side by side visualization of the cantilever plates.}
			\label{fig:plate_sbs}
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\begin{tikzpicture}
					\draw[line width = 0.3mm] (-0.8,-0.1) -- (6.2,-0.1);
					\draw[line width = 0.3mm] (0.8,0.7) -- (7.8,0.7);
					\draw[line width = 0.3mm] (-0.8,-0.1) -- (0.8,0.7);
					\draw[line width = 0.3mm] (6.2,-0.1) -- (7.8,0.7);

					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+4});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+3.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+3});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+2.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+2});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1});

					\draw[scale=0.5, domain=-3:-1.5, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+0.5});
					\draw[scale=0.5, domain=2:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+0.5});

					\draw[scale=0.5, domain=-3:-0.5, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x});
					\draw[scale=0.5, domain=-3:0.5, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-0.5});
					\draw[scale=0.5, domain=-3:1.5, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-1});
					\draw[scale=0.5, domain=-3:2.5, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-1.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-2});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-2.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-3});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-3.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-4});

					\node at (7.2,0.1) {$b$};
					\node at (3.2,-0.27) {$\ell = 1$};

					\draw[line width = 0.2mm] (0.8,0.7) -- (7.8,0.7);
					\draw[line width = 0.2mm] (-0.8,-0.1) -- (0.8,0.7);
					\draw[line width = 0.2mm] (6.2,-0.1) -- (7.8,0.7);
				\end{tikzpicture}
				\subcaption{Reissner-Mindlin cantilever plate}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\begin{tikzpicture}
					\draw[line width = 0.3mm] (-0.8,0.1) -- (6.2,0.1);
					\draw[line width = 0.3mm] (-0.8,-0.1) -- (6.2,-0.1);
					\draw[line width = 0.3mm] (6.2,-0.1) -- (6.2,0.1);
					\draw[line width = 0.3mm] (-0.8,-0.1) -- (-0.8,0.1);

					\draw[line width = 0.3mm] (0.8,0.9) -- (7.8,0.9);
					\draw[line width = 0.3mm] (7.8,0.7) -- (7.8,0.9);

					\draw[line width = 0.3mm] (-0.8,0.1) -- (0.8,0.9);
					\draw[line width = 0.3mm] (6.2,0.1) -- (7.8,0.9);
					\draw[line width = 0.3mm] (6.2,-0.1) -- (7.8,0.7);

					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+4});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+3.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+3});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+2.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+2});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1});

					\draw[scale=0.5, domain=-3:-1.5, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+0.5});
					\draw[scale=0.5, domain=2.5:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+0.5});

					\draw[scale=0.5, domain=-3:-0.5, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x});
					\draw[scale=0.5, domain=-3:0.5, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-0.5});
					\draw[scale=0.5, domain=-3:1.5, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-1});
					\draw[scale=0.5, domain=-3:2.5, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-1.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-2});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-2.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-3});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-3.5});
					\draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-4});

					\node at (6.85,0.6) {$b$};
					\node at (6,0) {$h$};
					\node at (3.2,-0.3) {$\ell = 1$};

					% \draw[line width = 0.1mm,->] (9,1) -- (10,1.6);
					% \draw[line width = 0.1mm,->] (9,1) -- (10.4,1);
					% \draw[line width = 0.1mm,->] (9,1) -- (9,-0.1);
					% \node at (9.2,1.4) {$e_3$};
					% \node at (9.7,0.8) {$e_1$};
					% \node at (8.8,0.4) {$e_2$};
				\end{tikzpicture}
				\subcaption{Three-dimensional cantilever plate}

			\end{minipage}
		}
	}
\end{figure}
\FloatBarrier

\subsection{Calculating the eigenvalues}
The eigenvalues for both models are calculated using the Finite Element Method. For the plate model, the Finite Element Method is derived in \ref{sec:FEM:Plate} and for the three-dimensional model, the Finite Element Method is derived in \ref{sec:FEM:3D}.

\subsubsection{Problem P-1E}
Find a vector function $u$ and a number $\lambda$ such that
\begin{eqnarray}
	K{u} & = & M\lambda{u}, 
\end{eqnarray} where $K$ and $M$ are the standard Finite Element Method matrices for bi-cubic basis functions. The definition of the matrices are given in \ref{sec:FEM:Plate}.

\subsubsection{Problem 3D-1E}
Find a real number $\lambda$ and a function $\bar{u} \in S^h$ such that
\begin{eqnarray}
	K\bar{u} & = & M\lambda{\bar{u}},
\end{eqnarray} where $K$ and $M$ are the standard Finite Element Method matrices for tri-cubic basis functions. The definition of the matrices are given in section \ref{sec:FEM:3D}.

\subsubsection{Accuracy of the eigenvalues}
Figure \ref{fig:conv_3d_eig} show the rate of convergence of the first 20 eigenvalues of Problem 3D-1E.

\begin{figure}[H]
    \centering
    \begin{adjustbox}{center}
        \includegraphics[scale=0.7]{Convergence.png}
    \end{adjustbox}
    \caption{Rate of convergence of the first 20 eigenvalues.}
    \label{fig:conv_3d_eig}
\end{figure}
\textcolor{red}{Update this figure.}

The number of elements can be chosen so that at least the first 10 eigenvalues are accurate to 5 significant digits.

\subsection{Comparing the mode shapes}
Similar to the previous sections, the mode shapes of the two models are compared, to be able to match up the eigenvalues.

\subsubsection{Mode shapes relating to plate type eigenvalues}
Figure \ref{fig:plate_mode_shapes} shows examples of mode shapes relating to plate-type eigenvalues. The plate models has many different shapes.

\FloatBarrier
\begin{figure}[h!]
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{Plate2.png}
				\subcaption{Plate - $\lambda_5 = 0.643$}
				\label{fig:minipage2}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{3D2.png}
				\subcaption{3D Model - $\lambda_5 = 0.645$}
				\label{fig:minipage1}
			\end{minipage}
	}}
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}{.8\textwidth}
				\centering
				\includegraphics[width=1\linewidth]{Plate1.png}
				\subcaption{Plate - $\lambda_6 = 1.92$}
				\label{fig:minipage1}
			\end{minipage}
			\begin{minipage}{.8\textwidth}
				\centering
				\includegraphics[width=1\linewidth]{3D1.png}
				\subcaption{3D Model - $\lambda_7 = 1.92$}
				\label{fig:minipage2}
			\end{minipage}
	}}
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{Plate3.png}
				\subcaption{Plate - $\lambda_8 = 2.74$}
				\label{fig:minipage1}
			\end{minipage}
			\quad
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{3D3.png}
				\subcaption{3D Model - $\lambda_8 = 2.74$}
				\label{fig:minipage2}
			\end{minipage}
	}}
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{Plate4.png}
				\subcaption{Plate - $\lambda_{12} = 8.96$ (Side view)}
				\label{fig:minipage2}
			\end{minipage}
			\quad
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{3D4.png}
				\subcaption{3D Model - $\lambda_{14} = 9.01$ (Side view)}
				\label{fig:minipage2}
			\end{minipage}
			\caption{Comparison of mode shapes relating to plate-type models. $b = 1/20$, $d = 1$.}
	}}
\end{figure}\label{fig:plate_mode_shapes}

\FloatBarrier

\subsubsection{Mode shapes relating to non-plate type eigenvalues}
Figure \ref{fig:non_plate_mode_shapes} shows examples of mode shapes not relating to plate-type eigenvalues.

\begin{figure}[h!]
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}{.8\textwidth}
				\centering
				\includegraphics[width=1\linewidth]{3Dnp1.png}
				\subcaption{3D Model - $\lambda_{6} = 1.35$ (Top view)}
				\label{fig:minipage1}
			\end{minipage}
			\begin{minipage}{.8\textwidth}
				\centering
				\includegraphics[width=1\linewidth]{3Dnp2.png}
				\subcaption{3D Model - $\lambda_{13} = 7.80$ (Top view)}
				\label{fig:minipage2}
			\end{minipage}
	}}
	\caption{Comparison of mode shapes not relating to plate-type models. $b = 1/20$, $d = 1$.}
\end{figure} \label{fig:non_plate_mode_shapes}

\subsection{Comparing the eigenvalues}
The eigenvalues of the two plate models can now be compared. For both the models, there are two parameters to consider for the height $h$ and width $b$. For realistic applications, the width of the plate is larger than its height.

For the comparisons, the investigation will consider a plate with small width $b = 0.25$, a plate of width equal to lenght $b = 1$ and a wide plate $1.75$. For each case, decreasing values of the height $h$ will be considered.

\FloatBarrier
\begin{table}[htbp]
	\scalebox{.8}{
	\makebox[\textwidth]{
		\caption{Comparison of eigenvalues with $b = 0.25$, with decreasing values of $h$.}
		\begin{tabular}{|cccc||cccc||cccc||cccc|}
			\hline
			\multicolumn{16}{|c|}{Comparison of Eigenvalues, $b = 0.25$} \\
			\hline\hline
			\multicolumn{4}{|c||}{$h = 1/5$}       & \multicolumn{4}{c||}{$h =1/10$}      & \multicolumn{4}{c||}{$h = 1/20$}      & \multicolumn{4}{c|}{$h = 1/30$} \\
			\hline
			{i} & {3D} & {j} & {Plate} & {i} & {3D} & {j} & {Plate} & {i} & {3D} & {j} & {Plate} & {i} & {3D} & {j} & {Plate} \\
			\hline
			1     & 0.12348 & 1     & {0.12249} & 1     & 0.032327 & 1     & 0.032184 & 1     & 0.008207 & 1     & 0.008186 & 1     & 0.003663 & 1     & 0.003656 \\
			2     & 0.18638 &       & -     & 2     & 0.18531 &       & {-} & 2     & 0.18476 &       & {-} & 2     & 0.14204 & 2     & 0.14173 \\
			3     & 2.4151 & 2     & {2.3954} & 3     & 1.161 & 2     & 1.1537 & 3     & 0.31436 & 2     & 0.31341 & 3     & 0.18456 &       & {-} \\
			4     & 3.5856 & 3     & {3.5368} & 4     & 1.3155 & 3     & 1.31  & 4     & 0.44696 & 3     & 0.44582 & 4     & 0.21514 & 3     & 0.21454 \\
			5     & 4.785 &       & -     & 5     & 4.7697 &       & {-} & 5     & 2.3862 & 4     & 2.3767 & 5     & 1.1001 & 4     & 1.0971 \\
			6     & 7.7889 &       & -     & 6     & 7.7674 & 4     & 7.9816 & 6     & 4.1984 & 5     & 4.1857 & 6     & 2.0374 & 5     & 2.0313 \\
			7     & 20.37 & 4     & {19.969} & 7     & 8.056 &       & {-} & 7     & 4.7615 &       & {-} & 7     & 4.1509 & 6     & 4.1367 \\
			8     & 21.722 & 5     & {21.537} & 8     & 12.034 & 5     & 11.972 & 8     & 7.7543 &       & {-} & 8     & 4.7585 &       & {-} \\
			9     & 25.181 &       & -     & 9     & 25.147 &       & {-} & 9     & 8.7609 & 6     & 8.7132 & 9     & 6.2078 & 7     & 6.1864 \\
			10    & 56.321 & 6     & {54.887} & 10    & 26.608 & 6     & 26.266 & 10    & 12.598 & 7     & 12.548 & 10    & 7.7495 &       & {-} \\
			11    & 60.267 & 7     & {59.71} & 11    & 34.438 & 7     & 34.197 & 11    & 22.619 & 8     & 22.456 & 11    & 11.064 & 8     & 11.015 \\
			12    & 65.513 &       & -     & 12    & 61.829 & 8     & 60.801 & 12    & 25.128 &       & {-} & 12    & 13.734 & 9     & 13.678 \\
			13    & 69.031 &       & -     & 13    & 65.518 &       & {-} & 13    & 27.297 & 9     & 27.155 & 13    & 23.854 & 10    & 23.724 \\
			14    & 114.09 & 8     & {110.7} & 14    & 69.133 & 9     & 69.548 & 14    & 47.123 & 10    & 46.695 & 14    & 25.121 &       & {-} \\
			15    & 117.88 & 9     & {116.66} & 15    & 70.211 &       & {-} & 15    & 50.515 & 11    & 50.174 & 15    & 26.054 & 11    & 25.927 \\
			16    & 127.05 &       & -     & 16    & 116.78 & 10    & 114.43 & 16    & 65.523 &       & {-} & 16    & 36.259 & 12    & 36.15 \\
			17    & 184.46 & 10    & {185.78} & 17    & 121.43 & 11    & 119.94 & 17    & 69.089 &       & {-} & 17    & 41.973 & 13    & 41.807 \\
			18    & 191.91 & 11    & {191.47} & 18    & 127.23 &       & {-} & 18    & 71.563 & 12    & 71.166 & 18    & 44.95 & 14    & 44.685 \\
			19    & 193.03 &       & -     & 19    & 178.27 & 12    & 175.79 & 19    & 81.444 & 13    & 80.894 & 19    & 45.798 & 15    & 45.507 \\
			20    & 193.71 &       & -     & 20    & 186.06 & 13    & 187.28 & 20    & 84.785 & 14    & 84.064 & 20    & 54.838 & 16    & 54.573 \\
			\hline
			\multicolumn{2}{|c}{Max RE:} & \multicolumn{2}{c||}{2.9764\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c||}{2.7585\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c||}{0.90845\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c|}{0.63525\%} \\
			\hline
		\end{tabular}%
		\label{tab:Table_plate_1}%
	}}
\end{table}%
\FloatBarrier

Table \ref{tab:Table_plate_1} shows that the narrow plate model compares well the to three-dimensional model. Also note that as the height $h$ decreases, the number of plate type eigenvalues within the first 20 eigenvalues increases.

The next table, table \ref{tab:Table_plate_2}, shows the comparison of the eigenvalues for a plate of width equal to its length, $b = 1$.

\FloatBarrier
\begin{table}[htbp]
	\scalebox{.8}{
	\makebox[\textwidth]{
		\caption{Comparison of eigenvalues with $b = 1$, with decreasing values of $h$.}
		\begin{tabular}{|cccc||cccc||cccc||cccc|}
			\hline
			\multicolumn{16}{|c|}{Comparison of Eigenvalues, $b = 1$} \\
			\hline\hline
			\multicolumn{4}{|c||}{$h = 1/5$}       & \multicolumn{4}{c||}{$h =1/10$}      & \multicolumn{4}{c||}{$h = 1/20$}      & \multicolumn{4}{c|}{$h = 1/30$} \\
			\hline
			{i} & {3D} & {j} & {Plate} & {i} & {3D} & {j} & {Plate} & {i} & {3D} & {j} & {Plate} & {i} & {3D} & {j} & {Plate} \\
			\hline
			1     & 0.12869 & 1     & {0.12743} & 1     & 0.033784 & 1     & 0.033626 & 1     & 0.008565 & 1     & 0.008543 & 1     & 0.003817 & 1     & {0.00381} \\
			2     & 0.62189 & 2     & {0.61703} & 2     & 0.18635 & 2     & 0.18562 & 2     & 0.049829 & 2     & 0.04957 & 2     & 0.022511 & 2     & {0.022405} \\
			3     & 1.3638 &       & -     & 3     & 1.1603 & 3     & 1.153 & 3     & 0.31428 & 3     & 0.31315 & 3     & 0.14192 & 3     & {0.14156} \\
			4     & 3.5808 & 3     & {3.5289} & 4     & 1.3579 &       & {-} & 4     & 0.50842 & 4     & 0.50676 & 4     & 0.23035 & 4     & {0.22969} \\
			5     & 5.8165 & 4     & {5.7694} & 5     & 1.864 & 4     & 1.8577 & 5     & 0.64595 & 5     & 0.64297 & 5     & 0.29502 & 5     & {0.29394} \\
			6     & 6.6048 & 5     & {6.5216} & 6     & 2.2928 & 5     & 2.2792 & 6     & 1.3555 &       & {-} & 6     & 0.8928 & 6     & {0.88735} \\
			7     & 7.8455 &       & -     & 7     & 6.4973 & 6     & 6.4546 & 7     & 1.9293 & 6     & 1.9157 & 7     & 1.156 & 7     & {1.1527} \\
			8     & 9.8027 &       & -     & 8     & 7.8167 &       & {-} & 8     & 2.5089 & 7     & 2.4975 & 8     & 1.265 & 8     & {1.261} \\
			9     & 17.03 & 6     & {16.802} & 9     & 8.4533 & 7     & 8.3671 & 9     & 2.7486 & 8     & 2.7369 & 9     & 1.3543 &       & - \\
			10    & 21.225 & 7     & {20.743} & 10    & 9.3453 & 8     & 9.2873 & 10    & 3.3087 & 9     & 3.2903 & 10    & 1.5352 & 9     & {1.529} \\
			11    & 23.935 & 8     & {23.552} & 11    & 9.8009 &       & {-} & 11    & 5.5365 & 10    & 5.4914 & 11    & 2.598 & 10    & {2.5803} \\
			12    & 24.694 &       & -     & 12    & 10.925 & 9     & 10.827 & 12    & 5.9696 & 11    & 5.9246 & 12    & 2.8161 & 11    & {2.7997} \\
			13    & 27.187 & 9     & {26.69} & 13    & 17.599 & 10    & 17.447 & 13    & 7.8024 &       & {-} & 13    & 4.2654 & 12    & {4.2474} \\
			14    & 28.844 &       & -     & 14    & 18.596 & 11    & 18.407 & 14    & 9.0173 & 12    & 8.9573 & 14    & 4.6674 & 13    & {4.6415} \\
			15    & 32.373 &       & -     & 15    & 24.729 &       & {-} & 15    & 9.8006 & 13    & 9.838 & 15    & 4.9397 & 14    & {4.9143} \\
			16    & 41.288 & 10    & {40.519} & 16    & 27.4  & 12    & 27.021 & 16    & 9.9172 &       & {-} & 16    & 5.7243 & 15    & {5.6803} \\
			17    & 42.082 & 11    & {41.246} & 17    & 28.81 &       & {-} & 17    & 10.365 & 13    & 10.286 & 17    & 6.6585 & 16    & {6.6025} \\
			18    & 51.427 &       & -     & 18    & 30.737 & 13    & 30.413 & 18    & 11.934 & 14    & 11.817 & 18    & 7.2935 & 17    & {7.2474} \\
			19    & 56.963 & 12    & {56.1} & 19    & 30.83 & 14    & 30.413 & 19    & 13.915 & 15    & 13.767 & 19    & 7.7966 &       & - \\
			20    & 57.741 &       & -     & 20    & 32.414 &       &       & 20    & 15.109 & 16    & 14.974 & 20    & 9.7996 &       & - \\
			\hline
			\multicolumn{2}{|c}{Max RE:} & \multicolumn{2}{c||}{2.2695\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c||}{1.3816\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c||}{1.0657\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c|}{0.84102\%} \\
			\hline
		\end{tabular}%
		\label{tab:Table_plate_2}%
	}}
\end{table}%
\FloatBarrier

Table \ref{tab:Table_plate_2} shows that the plate model compares slightly less well than the case where $b=0.25$, but the comparison is still very good. Again it seems that as $h$ decreases, the number of plate-type eigenvalues within the first 20 eigenvalues increases, and in this case there are more than the previous case.

\end{document}