### 2023-11-07 table_of_symbols.tex

\documentclass[main.tex]{subfiles}
\begin{document}

\chapter*{List of Symbols}

The following table provides a list of symbols that will be used throughout this dissertation along with their descriptions:

\noindent
\begin{tabularx}{\textwidth}{cXr}
\toprule
Symbol & Description & First Appearance (Page) \\ 
\midrule
\endhead % Header to be repeated on every page
\( R^n \) & n-dimensional space of real numbers & \pageref{sym:Rn} \\
\( u/w \) & displacement vector & \pageref{sym:u} \\
\( \Omega \) & subset of $R^n$ & \pageref{sym:Omega} \\
\( \rho \) & density & \pageref{sym:rho} \\
\( \partial_x^n \) & n-th partial derivative with respect to $x$ & \pageref{sym:partial_diff} \\
\( \sigma_{ij} \) & element of the stress tensor T & \pageref{sym:sigmaij} \\
\( Q \) & force per unit volume & \pageref{sym:Q} \\
\( \textrm{div}T \) & divergence of T & \pageref{sym:divT} \\
\( \textrm{Tr}(T) \) & Trace of T & \pageref{sym:TrT} \\
\( \mathcal{E} \) & infinitesimal strain tensor & \pageref{sym:mathcalE} \\
\( \varepsilon_{ij} \) & element of the infinitesimal strain tensor $\mathcal{E}$ & \pageref{sym:varepsilonij} \\
\( E \) & Young's modulus & \pageref{sym:E} \\
\( \nu \) & Poisson's ratio & \pageref{sym:nu} \\
\( t \) & time & \pageref{sym:t} \\
\( G \) & shear modulus of elasticity & \pageref{sym:G} \\
\( \tau \) & dimensionless time & \pageref{sym:tau} \\
\( \xi \) & dimensionless space & \pageref{sym:xi} \\
\( Q^* \) & dimensionless force per unit volume & \pageref{sym:Qstar} \\
\( \sigma^*_{ij} \) & dimensionless stress tensor element & \pageref{sym:sigmaijstar} \\
\( \ell \) & dimension representing length & \pageref{sym:ell} \\
\( \kappa^2 \) & some dimensionless constant/shear correction factor & \pageref{sym:kappa2}/\pageref{sym:kappa2T} \\
\( I \) & identity matrix & \pageref{sym:I} \\
\( \gamma \) & a dimensionless constant & \pageref{sym:gamma} \\
\( n \) & a normal vector & \pageref{sym:n} \\
\( \Sigma/\Gamma \) & distinct parts of $\Omega$ & \pageref{sym:SigmaGamma} \\
\( \phi \) & arbitrary vector/rotation of cross-section of Timoshenko beam & \pageref{sym:phi}/\pageref{sym:phiT} \\
\( C^n \) & set of n-times continuously differentiable functions & \pageref{sym:Cn} \\
\( a(\cdot,\cdot),b(\cdot,\cdot),c(\cdot,\cdot) \) & bilinear forms & \pageref{sym:a},\pageref{sym:b},\pageref{sym:c} \\
\( T(\Omega) \) & test function space on $\Omega$ & \pageref{sym:TOmega} \\
\( dV \) & volume integral meassure & \pageref{sym:dV} \\
\( dS \) & surface integral meassure & \pageref{sym:dS} \\
\( dA \) & area integral meassure & \pageref{sym:dA} \\
\( ds \) & line integral meassure & \pageref{sym:ds} \\
\( A \) & area of a cross-section & \pageref{sym:A} \\
\( V \) & shear force & \pageref{sym:V} \\
\( I \) & area moment of inertia & \pageref{sym:Iinertia} \\
\( M \) & moment & \pageref{sym:M} \\
\( \alpha/\beta \) & dimensionless constants & \pageref{sym:alphabeta} \\
\( e_n \) & unit vector in the direction n & \pageref{sym:en} \\
\( \mathbf{Q} \) & constitutive equation for the Reissner-Mindlin plate model & \pageref{sym:bfQ} \\
\( L^n \) & space of n-power Lebesgue integrable functions & \pageref{sym:Ln} \\
\( H^n \) & space of functiontions with weak derivatives up to order n (n'th dimensional Sobolev space) & \pageref{sym:Hn} \\
\( X \) & global space & \pageref{sym:Xspace} \\
\( W \) & inertia space & \pageref{sym:Wspace} \\
\( V \) & energy space & \pageref{sym:Vspace} \\
\( ||\cdot||_X \) & norm in the space $X$ & \pageref{sym:norm} \\
\( C^\infty \) & space of smooth functions & \pageref{sym:Cinfty} \\
\( C^\infty_0 \) & space of smooth functions with compact support & \pageref{sym:Cinftyzero} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\
\( u \) & Displacement vector & \pageref{sym:u} \\

\bottomrule
\end{tabularx}

\end{document}