### 2023-11-12 table_of_symbols.tex

\documentclass[main.tex]{subfiles}
\begin{document}

\chapter*{List of Symbols}

% Vector Spaces and Related Concepts
\noindent
\textbf{Vector Spaces and Related Concepts}\\
\begin{tabularx}{\textwidth}{cXr}
\toprule
Symbol & Description & First Appearance (Page) \\ 
\midrule
\( \mathbb{N} \) & set of natural numbers \( j \) & \pageref{sym:natural} \\
\( R^n \) & n-dimensional space of real numbers & \pageref{sym:Rn} \\
\( \Omega \) & subset of \( R^n \), usually representing a body/reference configuration & \pageref{sym:Omega} \\
\( \partial \Omega \) & the boundary of $\Omega$ & \pageref{sym:boundaryOmega} \\
\( \bar{\Omega} \) & the boundary of $\Omega$ & \pageref{sym:baromega} \\
\( C^n \) & set of n-times continuously differentiable functions & \pageref{sym:Cn} \\
\( C^\infty \) & space of smooth functions & \pageref{sym:Cinfty} \\
\( C^\infty_0 \) & space of smooth functions with compact support & \pageref{sym:Cinftyzero} \\
\( L^n \) & space of n-power Lebesgue integrable functions & \pageref{sym:Ln} \\
\( H^n \) & space of functions with weak derivatives up to order n (n'th dimensional Sobolev space) & \pageref{sym:Hn} \\
\( X \) & global space & \pageref{sym:Xspace} \\
\( V \) & inertia space & \pageref{sym:Wspace} \\
\( W \) & energy space & \pageref{sym:Vspace} \\
\( T(\Omega) \) & test function space on \( \Omega \) & \pageref{sym:TOmega} \\
\( S^h \) & finite dimensional subspace & \pageref{sym:Sh1} \\
\( \mathcal{P}_j \) & set of all polynomials of degree at most \( j \) & \pageref{sym:poly} \\
\( E_n \) & space spanned by the orthonormal basis vectors $e_i$ & \pageref{sym:En} \\

\bottomrule
\end{tabularx}

% Mathematical Measures and Operations
\noindent
\textbf{Mathematical Measures and Operations}\\
\begin{tabularx}{\textwidth}{cXr}
\toprule
Symbol & Description & First Appearance (Page) \\ 
\midrule
\( a(\cdot,\cdot), b(\cdot,\cdot), c(\cdot,\cdot) \) & bilinear forms & \pageref{sym:a}, \pageref{sym:b}, \pageref{sym:c} \\
\( (\cdot, \cdot)_X \) & innerproduct of X & \pageref{sym:innerproduct} \\
\( ||\cdot||_X \) & norm in the space \( X \) & \pageref{sym:norm} \\
\( \partial_x^nf \) & n-th partial derivative of f with respect to \( x \) & \pageref{sym:partial_diff} \\
\( \textrm{div} X \) & divergence of the matrix \( X \) & \pageref{sym:divT} \\
\( \textrm{Tr}(X) \) & Trace of the matrix \( X \) & \pageref{sym:TrT} \\
\( \textrm{det}(M) \) & determinant of M & \pageref{sym:determinant} \\
\( \textrm{span}(\cdot) \) & span of a set & \pageref{sym:span} \\
\( dV \) & volume integral measure & \pageref{sym:dV} \\
\( dS \) & surface integral measure & \pageref{sym:dS} \\
\( dA \) & area integral measure & \pageref{sym:dA} \\
\( ds \) & line integral measure & \pageref{sym:ds} \\
\( \mathcal{E}(\cdot) \) & energy function & \pageref{sym:Energy} \\
\( R(\cdot) \) & Rayleigh quotient & \pageref{sym:Rayleigh} \\
\( \Pi \) & interpolation operator & \pageref{sym:interpolation} \\
\( \bar{u} \) & another form explicity showing u is a vector & \pageref{sym:baru} \\
\( \textrm{RE} \) & abbriviation for the relative error & \pageref{sym:RE} \\
\bottomrule
\end{tabularx}

% Physical Quantities and Parameters
\noindent
\textbf{Physical Quantities and Parameters}\\
\begin{tabularx}{\textwidth}{cXr}
\toprule
Symbol & Description & First Appearance (Page) \\ 
\midrule
\( \lambda \) & eigenvalue & \pageref{sym:lambda} \\
\( u \) or \( w \) & displacement vector & \pageref{sym:u} \\
\( \phi \) & arbitrary vector/rotation of cross-section of Timoshenko beam & \pageref{sym:phi} \\
\( Q \) & force per unit volume & \pageref{sym:Q} \\
\( \rho \) & density & \pageref{sym:rho} \\
\( T \) & stress tensor & \pageref{sym:T} \\
\( \sigma_{ij} \) & element of the stress tensor \( T \) & \pageref{sym:sigmaij} \\
\( \mathcal{E} \) & infinitesimal strain tensor & \pageref{sym:mathcalE} \\
\( \varepsilon_{ij} \) & element of the infinitesimal strain tensor \( \mathcal{E} \) & \pageref{sym:varepsilonij} \\
\( E \) & Young's modulus & \pageref{sym:E} \\
\( \nu \) & Poisson's ratio & \pageref{sym:nu} \\
\( t \) & time & \pageref{sym:t} \\
\( \ell \) & dimension representing length & \pageref{sym:ell} \\
\( h \) & dimension representing height & \pageref{sym:height} \\
\( b \) &  dimension representing width & \pageref{sym:width} \\
\( G \) & shear modulus of elasticity & \pageref{sym:G} \\
\( A \) & area of a cross-section & \pageref{sym:A} \\
\( V \) & shear force & \pageref{sym:V} \\
\( I \) & area moment of inertia & \pageref{sym:Iinertia} \\
\( M \) & moment & \pageref{sym:M} \\
\( f^* \) & dimensionless form of f & \pageref{sym:sigmaijstar} \\
\( \tau \) & dimensionless time & \pageref{sym:tau} \\
\( \xi \) & dimensionless space & \pageref{sym:xi} \\
\( \alpha/\beta \) & dimensionless constants & \pageref{sym:alphabeta} \\
\( \kappa^2 \) & some dimensionless constant/shear correction factor & \pageref{sym:kappa2} \\
\( I \) & identity matrix & \pageref{sym:I} \\
\( \gamma \) & a dimensionless constant & \pageref{sym:gamma} \\
\( n \) & a normal vector & \pageref{sym:n} \\
\( \Sigma/\Gamma \) & distinct parts of \( \Omega \) & \pageref{sym:SigmaGamma} \\
\( \mu \) & eigenvector & \pageref{sym:mu} \\
\( e_i \) & orthonormal basis vector & \pageref{sym:e_i} \\

\bottomrule
\end{tabularx}

\end{document}