### 2023-11-26 files/Chapter3/dynamicconvergence.tex
Chapter 3

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Introduction}

In this chapter, two important convergence results of the Finite Element Method
(FEM) are discussed. The first result looks at the convergence of the Galerkin
approximation for second order hyperbolic type problems (or general vibation problems). The papers considered for the first result are \cite{BV13} and \cite{BSV17}. The second result is from the textbook~\cite{SF73} and examines the
convergence of eigenvalues and eigenfunctions for a vibration
problem, using the Finite Element Method.

\section{Galerkin approximation for second order hyperbolic type problems}\label{sec:convergence-of-the-galerkin-approximation}
In the article \cite{BV13}, the authors investigate the convergence of the
Galerkin approximation for second order hyperbolic type problems. The article
\cite{BSV17} extends the work of \cite{BV13} by including general damping and
damping at the endpoints. For the models in Chapter 1, \cite{BV13} is
sufficient while \cite{BSV17} provides more insight and improved notation.

In Section 2.2 of this dissertation, Problem GVar is presented. It is identical to the problem in
\cite{BV13}. For convenience, the problem is repeated here.

\subsection{Formulation of the Galerkin approximation}
Recall the spaces $V$, $W$ and $X$ from Section \ref{sec:existence:ExistenceAndUniqueness} where $V \subset W \subset X$.

\subsubsection*{Problem GVar}
Given a function $f:J\rightarrow X$, find a function $u\in C(J,\ X)$ such that $u'$ is continuous at $0$ with respect to
$\Vert \cdot \Vert_{W}$ and for each $t\in J,\ u(t)\in V,\ u'(t) \in V,\ u''(t)\in W$ and
\begin{eqnarray}
	c(u''(t),v)+a(u'(t),v)+b(u(t),v)= (f(t),v)_{X} \ \ \ \ \textrm{for each} \ v \in V, \label{GV_PB1}
\end{eqnarray}
while $u(0)=u_{0}$, $u'(0)=u_{1}$.

Assume that the assumptions \textbf{A1}-\textbf{A4} from Section 2.2 are
satisfied ensuring that Problem GVar has a unique solution.

Before the theory of \cite{BV13} can be discussed, some preliminary work is
necessary. The structure of this section is as follows. First the Galerkin
approximation for Problem GVar is derived. Then an equivalent system of
ordinary differential equations is derived using the Finite Element Method.
Finally, the convergence of the Galerkin approximation is discussed using the
work of the article \cite{BV13}.

\textcolor{red}{***************** NEW ********************************}

\subsubsection{*** Example of the Galerkin approximation of a cantilever Timoshenko beam ***}
Consider the example of the cantilever Timoshenko beam model from Section \ref{sec:existence}. In this section the variational problem Problem T-2V is given in terms of bilinear forms.

\subsubsection{Problem T-2V}
Find a function $u \in T[0,1]$ such that for all $t \in [0,1]$,
\begin{eqnarray*}
	c(u''(t),v) + b(u(t),v) & = & (Q(t),v),
\end{eqnarray*} for each $v \in T[0,1]$.

The interval $[0,1]$ is divided into $n$ equal subintervals $[x_{i}, x_{i+1}]$, each of length $\displaystyle h = \frac{1}{n}$, such that $x_i = ih$ for $i = 0,1,...,n$.

Consider a a set of $n+1$ linear independent, piecewise linear basis functions $\delta_i \in T[0,1]$. The subset of these functions that satisfies the boundary conditions are called admissible basis functions. For Problem 2-T, the admissible basis functions are ${\delta_i}$ for $i = 1,2,...,n$. Define the space $S^h$ as the space spanned by the admissible basis functions, i.e.
\begin{eqnarray*}
	S^h = \textrm{span}\{\delta_1, \delta_2, ..., \delta_n\}.
\end{eqnarray*}

This space $S^h$ is a finite dimensional subspace of $T[0,1]$. Define the following functions $w^h \in S^h$ and $\phi^h \in S^h$ as
\begin{eqnarray*}
	w^h(t) & = & \sum_{i=1}^{n} w(x_i^*,t) \delta_i(t),\\
	\phi^h(t) & = & \sum_{i=1}^{n} w(x_i^*,t) \delta_i(t),
\end{eqnarray*} where $x^*_i \in [x_i, x_{i+1}]$.

Using these functions, the Galerkin approximation for Problem T-2, referred to as Problem T-2G, can be derived.

\subsubsection{Problem T-2G}
Find a function $u_h \in S^h$ such that for all $t \in [0,1]$,
\begin{eqnarray*}
	c(u''_h(t),v) + b(u_h(t),v) & = & (Q^I(t),v),
\end{eqnarray*} for each $v \in S^h$, where $u_h = \langle w^h, \phi^h \rangle$. $Q^I$ is a vector of the values of $Q(x^*_i)$ for $i = 0,1,2,...,n$.

This example serves as an illustration of the derivation of the Galerkin approximation and the convention of symbols before the general case is presented. Piecewise linear basis functions are used for this example, but other basis functions can be used. In Chapter 5, the basis functions used are piecewise cubic Hermite polynomials.

For the general case presented below, $S^h$ is a finite dimensional subspace of $V$.

\textcolor{red}{***************** NEW ********************************}

\subsubsection*{Problem GVar$^h$}
Given a function $f: J \rightarrow X$, find a function $u_h \in C^2(J, S^h)$ such that for each $t\in J$
\begin{eqnarray}
	c(u_h''(t),v)+a(u_h'(t),v)+b(u_h(t),v)= (f(t),v)_{X} \ \ \ \ \textrm{for each} \ v \in S^h, \label{DC_E2}
\end{eqnarray}
with the initial values $u_h(0)=u^h_{0}$ and $u'_h(0)=u^h_{1}$. The initial conditions $u^h_{0}$ and $u^h_{1}$ are projections of $u_0$ and $u_1$ in the finite dimensional space $S^h$.

\subsection{System of ordinary differential equations}
Problem GVar$^h$ is equivalent to a system of second order differential equations. Consider the standard FEM matrices defined by
\begin{eqnarray*}
	K_{ij} & = & b(\phi_j, \phi_i),\\
	C_{ij} & = & a(\phi_j, \phi_i),\\
	M_{ij} & = & c(\phi_j, \phi_i),\\
	F_{i}(t) & = & c(f(t), \phi_i),
\end{eqnarray*} where $\phi_i$ and $\phi_j$ are admissible basis functions.

Using these matrices, Problem GVar$^h$ is rewritten as a system of ordinary
differential equations denoted by Problem ODE.

Recall that $u^h(t) = \sum_{k} u_k(t)\phi_k$ where $\bar{u}_k = \left( u_1(t), u_2(t), ...  , u_n(t)\right)$ where each $\phi_k$ corresponds to a node number $k$. More complex cases are treated in Chapter \ref{ch:finite-element-method}.

\subsubsection*{Problem ODE}
Find a function $\bar{u} \in S^h$ such that \label{sym:baru}
\begin{eqnarray}
	M\bar{u}'' + C \bar{u}' + K\bar{u} = F(t) \ \ \ \textrm{ with } \bar{u}(0) = \bar{u}_0 \textrm{ and }  \bar{u}(1) = \bar{u}_1 \label{ODE}
\end{eqnarray}

The following propositions related to Problem ODE are given in \cite{BV13}.

\newtheorem{DC_Prop2}{Proposition}
\begin{DC_Prop2}
	If $F\in C(J)$, then Problem ODE has a unique solution for each pair of vectors $\bar{u}_0$ and $\bar{u}_1$
\end{DC_Prop2}

\newtheorem{DC_Prop3}[DC_Prop2]{Proposition}
\begin{DC_Prop3}
	Suppose $M, K, C, F,\bar{u}_0$ and $\bar{u}_1$ are defined as above. Then, the function $u_{h}$ is a solution of Problem GVar$^{h}$ if and only if the function $\bar{u}$ is a solution of Problem ODE.
\end{DC_Prop3}

Proposition 2 provides a link between the solution of Problem ODE and the
solution of Problem GVar$^h$. Theorem 1 below follows.

\newtheorem{DC_Thm2}{Theorem}
\begin{DC_Thm2} \label{DC_THM2}
	If $f \in C(J,X)$, then there exists a unique solution $u_h \in C^2(J, S^h)$ for Problem GVar$^h$ for each $u_0^h$ and $u_1^h$ in $S^h$. If $f = 0$ then $u_h \in C^2((-\infty, \infty))$.
\end{DC_Thm2}

It is now required to find an approximation for the solution of $\bar{u}$ of Problem ODE.

Consider the time interval $J = [0,T]$. Divide $J$ into $N$ steps with length $\tau = \frac{T}{N}$. Each interval can be expressed as $[t_{k-1}, \ t_k]$ for $k = 1,...,N$. Denote the approximation of $u_h$ on the interval $[t_{k-1}, \ t_k]$ by $u_k^h$, i.e. $u_h(t_k)$ corresponds to $u_k^h$ for each $k$. (Recall that $u_h(t) = \sum u_i(t)\phi_i$.) A finite difference method is used to compute $u^h_k$ for each $k$.

\subsection{Error estimates}\label{e_est}
In this subsection we consider estimates for the error $u(t_k) - u^h_k$. To simplify the process, the error is divided into errors for the semi-discrete problem and the fully discrete problem.

Under the assumptions \textbf{A1}-\textbf{A4} of Section 2 and continuity of $f$, there exists a unique solution for Problem GVar$^h$. The next step is to show that the solution of Problem GVar$^h$ converges to the solution of Problem GVar.

Let $u$ be the solution of Problem GVar and $u^h$ be the solution of Problem GVar$^h$. The authors of \cite{BV13} define the following error,
\begin{eqnarray}
	e^h(t) = u(t) - u^h(t). \label{Error}
\end{eqnarray}

In \cite{BV13} it is assumed that there exists a subspace $H$ of $V$ and a positive integer $\alpha$ such that
\begin{eqnarray*}
	\inf_{v\in S^{h}}\Vert w-v\Vert_{V}\leq{C}h^{\alpha}|||w|||_{H},
\end{eqnarray*} for each $w \in H$ where $|||w|||_H$ is a norm or semi-norm for $H$.

\newtheorem*{DC_Thm6_1}{Theorem 2}
\begin{DC_Thm6_1}
	If $u(t) \in H$ and $u'(t) \in H$ for each $t$, then
	\begin{eqnarray*}
		||e^h(t)||_W & \leq & C h^\alpha \left( |||u(t)||_H + |||u'(t)||_H \right),
	\end{eqnarray*} for each $t$.
\end{DC_Thm6_1}

From Theorem 2 for the semi-discrete problem an error estimate for
$e(t) = [u(t)-u_h(t)]$ with respect to the norm of $W$ was obtained. The authors of \cite{BV13} then proceed to obtain an error estimate for $e_k = [u_h(t_k) - u^h_k]$.

The error can then be expressed as
\begin{eqnarray}
	e(t_k) = u(t_k) - u^h_k & = & [u(t_k)-u_h(t_k)] + [u_h(t_k) - u^h_k]. \label{error_e}
\end{eqnarray}

In \eqref{error_e}, the the error for the semi-discrete problem is the term $u(t_k)-u_h(t_k)$ and the term $u_h(t_k) - u^h_k$ is the error for the fully disrete appreximation of the semi-discrete approximation.

Since the dimension of $S^h$ is not fixed, the equivalence of norms
cannot be used, and therefore this error estimate for $e_k = [u_h(t_k) - u^h_k]$
should also be with respect to the norm of $W$. The local error $e_1$ can be estimated using Taylor polynomials, but then $e_k$ `grows' as $k$ increases.

A stability result is derived in \cite{BV13}. Recall that $\displaystyle \tau = \frac{T}{N}$.

\newtheorem*{DC_Lem3}{Lemma}
\begin{DC_Lem3}
	\begin{eqnarray*}
		\max\Vert e_{n}\Vert_{W}^{2}\leq KT\tau
	\end{eqnarray*}
	where $K$ depends on $u$, $u_h$ and their derivatives.
\end{DC_Lem3}

Using this lemma, \cite{BV13} prove the error estimate. The error estimate for the term $[u_h(t_k) - u^h_k]$ with respect to the norm of $W$ as proven by the authors of \cite{BV13} is presented here as Theorem 3.

\newtheorem*{DC_Thm5_1}{Theorem 3}
\begin{DC_Thm5_1}
	If $f\in C^{2}([0,T],X)$, then
	\begin{eqnarray*}
		\Vert u(t_{k})-u_{k}^{h}\Vert_{W}\ & \leq & K\tau^2
	\end{eqnarray*}
	for each $t\in(0,\ T)$.
\end{DC_Thm5_1}

\subsection{Main result}
Finally, Theorem 2 of the semi-discrete problem and Theorem 3 of the fully
discrete problem gives an error estimate for the error $e(t)$. Consequently,
the error estimate $e^h(t) = u(t) - u^h(t)$ is obtained.

The main result proving the convergence of the solution of the Galerkin
Approximation is given in \cite{BV13} as follows.
\newtheorem*{DC_Thm5}{Theorem 4}
\begin{DC_Thm5}
	{Main Result}\\
	If $f\in C^{2}([0,T],X)$, then
	\begin{eqnarray*}
		\Vert u(t_{k})-u_{k}^{h}\Vert_{W}\ & \leq & K\tau^2
	\end{eqnarray*}
	for each $t\in(0,\ T)$.
\end{DC_Thm5}

The constant $K$ depend on $u$, $u_h$ and their derivatives.

\end{document}

Define the error $e_h$ as the error between the solutions of Problem GVar and
Problem GVar$^h$, i.e.
\begin{eqnarray}
	e_h(t) = u(t)-u_h(t). \label{Error}
\end{eqnarray}

Since $u$ is unknown, the error $e_h$ cannot be calculated directly. Instead
the authors of \cite{BV13} derive estimate for this error using projections.***
\\

Define a projection operator $P_h$ by
\begin{eqnarray*}
	b(u - P_hu, v) = 0 \ \ \ \textrm{ for all } v \in S^h.
\end{eqnarray*}

Define the projection $P$ such that for any function $u$,
\begin{eqnarray*}
	(Pu)(t) = Pu(t) \ \ \ \textrm{ for all } t \in (0,T).
\end{eqnarray*}

Same a \cite{BV13}, $P_h$ is renamed to $P$ for convenience. $P_h$ will be used
when it is needed to differentiate between $P$ and $P_h$. Define the following
errors $e_p$ and $e$ by
\begin{eqnarray*}
	e(t) & = & Pu(t) - u_h(t),\\
	e_p(t) & = & u(t) - Pu(t).
\end{eqnarray*}
Then \eqref{Error} can be rewritten as
\begin{eqnarray*}
	e_h(t) = e(t) - e_p(t).
\end{eqnarray*}

Estimates for the norm of $e_p$ is obtained in \cite{BV13} using approximation
theory. It is only left to find an estimate for $e(t)$. The approach of
\cite{BV13} is to estimate the error $e$ using the projection error $e_p$ and
errors for the initial conditions.

\subsubsection*{Assumptions}
The following assumptions are made in \cite{BV13} for the convergence results.
\begin{itemize}
	\item[] \textbf{E1} - The solution $u$ of Problem GVar has the property that $(Pu) \in C^2(0,T)$.

	\item[] \textbf{E2} - The solution $u$ of Problem GVar satisfies $u \in C^1([0,T]; V) \cap C^2((0,T);V)$.
\end{itemize}

\subsubsection{Remark}
- Assumption \textbf{E2} is a weak assumption and follows from the existence theorem, Theorem 1 of Section 2.2. In [BSV17], \textbf{E2} is no longer a weak assumption but a necessary assumption. *****\\

From this Proposition in \cite{BV13} the general variational problem in terms
of the errors $e_h$ and $e$. \newtheorem{DC_Prop1}{Proposition}
\begin{DC_Prop1}
	If $u$ is a solution of Problem GVar and satisfies Assumption E1, then
	\begin{eqnarray*}
		c(e_h''(t),v)+a(e_h'(t),v)+b(e_h(t),v) = 0 \ \ \ \ \textrm{for each} \ v \in S^h.
	\end{eqnarray*}
\end{DC_Prop1}

\subsection{Fundamental estimate}
The authors of \cite{BV13} use the form of Proposition 1, to give a
generalization of an estimate for the error $e(t)= Pu(t)- u_h(t)$.

\newtheorem{DC_Lem2}{Lemma}
\begin{DC_Lem2}
	If the solution $u$ of Problem $GVAr$ satisfies Assumption El, then for $t\in[0,\ T],$
	\begin{eqnarray*}
		\Vert e(t)\Vert_{W}\ &\leq\ & \sqrt{2}(\Vert e(0)\Vert_{W}+3T\Vert e_{h}'(0)\Vert_{W}+3\int_{0}^{T}\Vert e_{p}'(t)\Vert_{W}\\
		& & +3K_{a}T\Vert e_{h}(0)\Vert_{W}+3K_{a}\int_{0}^{T}\Vert e_{p}(t)\Vert_{W})
	\end{eqnarray*}
\end{DC_Lem2}

Similarly, for $a = 0$ (weak damping as explained in Section 2.2), the
following corollary is obtained in a similar proof to Lemma 1.
\newtheorem{DC_Cor1}{Corollary}
\begin{DC_Cor1}
	If the solution $u$ of Problem $GVar$ with $a=0$ satisfies Assumption $El$, then for $t\in[0, T],$
	\begin{eqnarray*}
		\Vert e(t)\Vert_{W} & \leq  & \sqrt{2}\Vert e(0) \Vert_W + 2T\Vert e_h'(0) \Vert_W + 4 \sqrt{T} \max\limits_{t \in [0,T]} \Vert e_p'(t) \Vert_W.
	\end{eqnarray*}
\end{DC_Cor1}

\subsection{Convergence and error estimated for the semi-discrete approximation}
Consider Problem GVar$^h$ with discretised space and continuous time variables.
This is known as a semi-discrete problem. This is a natural intermediate step
between a continuous and a fully discrete problem.\\

\newtheorem*{DC_Thm3}{Theorem 2}
\begin{DC_Thm3}
	If the solution $u$ of Problem $G$ satisfies Assumption Al, then
	\begin{eqnarray}
		\Vert u(t)-u_{h}(t)\Vert_{W} & \leq & \Vert e_{p}(t)\Vert_{W}+\sqrt{2}(\Vert Pu_{0}-u_{0}\Vert_{W}+3T\Vert u_{1}-u_{1}^{h}\Vert_{W}+3\int_{0}^{T}\Vert e_{p}'(t)\Vert_{W} \nonumber \\
		& & +(1+3K_{a}T) \Vert u_{0}-u_{0}^{h}\Vert_{W}+3K_{a}\int_{0}^{T}\Vert e_{p}(t)\Vert_{W}) \label{DC_E7}
	\end{eqnarray}
	for each $t\in[0,\ T].$
\end{DC_Thm3}
\begin{comment}
\begin{proof}
	From (3.2), $\Vert u(t)-u_{h}(t)\Vert_{W}\leq\Vert e_{p}(t)\Vert_{W}+\Vert e(t)\Vert_{W}$.\\

	The result follows from Lemma 4.1 since $\Vert
		Pu_{0}-u_{0}^{h}\Vert_{W}\leq\Vert Pu_{0}-u_{0}\Vert+\Vert
		u_{0}-u_{0}^{h}\Vert.$
\end{proof}
\end{comment}

To be able to use Theorem in application, it is necessary to be able to
estimate an element in $V$ with an element in $S^h$. Suppose $h$ is related to
the dimension n of $S^h$ such that $h \rightarrow 0$ as $n \rightarrow
	\infty$.\\

Another assumption is introduced by \cite{BV13}.

\subsubsection*{Assumption E3}
There exists a subspace $H$ of $V$ and a positive integer $\alpha$ such that if $w\in H,$ then
\begin{eqnarray*}
	\inf_{v\in S^{h}}\Vert w-v\Vert_{V}\leq\hat{C}h^{\alpha}|||w|||_{H},
\end{eqnarray*}
where $|||w|||_{H}$ is a norm or semi-norm for $H.$\\

Using this assumption, the projection errors in Theorem 2 can be expressed as
\begin{eqnarray*}
	||e'_p(t)||_W & \leq & C_b \hat{C} h^\alpha |||u'(t)|||_H,\\
	||u_0 - u_0^h||_W & \leq & C_b \hat{C} h^\alpha |||u_0|||_H,
\end{eqnarray*} if $u^h_0$ and $u^h_1$ are chosen as $Pu_0$ and $Pu_1$. However choosing $u^h_0$ and $u^h_1$ in this form is not possible in general. Let $\Pi$ denote the interpolation operator such that $u_0^h = \Pi u_0$ and $u_1^h = \Pi u_1$.

\newtheorem*{DC_Thm4}{Theorem 3}
\begin{DC_Thm4}
	Suppose Assumption E3 holds and $u_{0}^{h}=\Pi u_{0}$ and $u_{1}^{h}=\Pi u_{1}.$ If the solution $u$ of Problem $GVar$ satisfies Assumption El, $u(t)\in H$ and $u'(t)\in H$, then
	\begin{eqnarray*}
		\Vert u(t)-u_{h}(t)\Vert_{W} & \leq & C_{b} \hat{C}h^{\alpha}|||u(t)|||_{H}+\sqrt{2}C_{b}\hat{C}h^{\alpha}(3T\max|||u'(t)|||_{H}\\
		& & +3K_{a}T\max|||u(t)|||_{H}+(2+3K_{a}T)|||u_{0}|||_{H} \\
		& & +3T|||u_{1}|||_{H}) ,
	\end{eqnarray*} for each $t\in[0, T].$
\end{DC_Thm4}