### 2024-04-07 final_seminar/title_and_abstract.tex

\documentclass{article}
\usepackage{lipsum} % for generating dummy text

\title{Finite element analysis of multi-dimensional and simplified models for beams and plates}
\author{Rudi du Toit}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
    Beams are fundamental structural elements in engineering. They are used in a wide range of applications, from bridges to buildings. Simplified models are often used in vibration problems due to their reduced complexity and computational demand. However, the accuracy of these models in practical applications is not always certain. In this presentation, we look at the validity of the Timoshenko beam model as well as the validity of the two-dimensional model. We use modal analysis and the Finite Element Method (FEM) to compare the models' natural frequencies and mode shapes to those of a more realistic three-dimensional model.
\end{abstract}

\end{document}