### 2024-04-15 final_seminar/main.tex
order change and added datat

\documentclass[8pt]{beamer}
\usetheme{metropolis} % Use the Metropolis theme
\usepackage{pgfplots}
\usepackage{caption,subcaption}
\usepackage[style=authoryear,backend=biber]{biblatex}
\usepackage[table]{xcolor}
\addbibresource{../references.bib}

\title{Finite element analysis of multi-dimensional and simplified models for beams}
\author{Rudi du Toit}
\date{\today}
\institute{University of Pretoria}

\newcommand{\footfullcitenonumber}[1]{
  \footnotetext{\fullcite{#1}}
}

\begin{document}

% Redefine the title page template
\setbeamertemplate{title page}{
    \begin{center}
        \usebeamercolor[fg]{titlegraphic}\inserttitlegraphic\par
        \vskip0.25em%
        {\fontsize{14}{16}\selectfont\usebeamerfont{title}\usebeamercolor[fg]{title}\inserttitle\par}% Increase the font size of the title
        \ifx\insertsubtitle\@empty%
        \else%
        \vskip0.25em%
        {\fontsize{12}{14}\selectfont\usebeamerfont{subtitle}\usebeamercolor[fg]{subtitle}\insertsubtitle\par}% Increase the font size of the subtitle
        \fi%     
        \vskip1em\par
        \usebeamerfont{author}\insertauthor\par
        \vskip1em\par
        \usebeamerfont{institute}\insertinstitute\par
        \vskip1em%
        \includegraphics[width=0.2\linewidth]{logo-up.jpg}\par % Include the logo here
        \vskip1em\par
        \usebeamerfont{date}\insertdate\par
        \vfill
    \end{center}
}

\begin{frame}[plain]
\titlepage
\end{frame}

\section{Introduction}
    \begin{frame}{Introduction}
        Beams are fundamental structural elements in engineering. They are used in a wide range of applications, from bridges to buildings.
        \pause

        Simplified models are often used in vibration problems due to their reduced complexity and computational demand. However, the accuracy of these models in practical applications is not always certain.
        \pause

        In this presentation, we look at the validity of the Timoshenko beam model as well as the validity of the two-dimensional model. We use modal analysis and the Finite Element Method (FEM) to compare the models' natural frequencies and mode shapes to those of a more realistic three-dimensional model.
    \end{frame}

    \begin{frame}{Structure of the presentation}
        \tableofcontents
    \end{frame}

\section{Literature study}
    \subsection{Comparison of linear beam theories}
        \begin{frame}{Literature study - Comparison of linear beam theories}
            Consider the article \textit{Comparison of linear beam theories} [\cite{LVV09}]

            \footfullcite{LVV09}
        \end{frame}

        \begin{frame}{Detailed Comparison: Timoshenko Beam Model vs. Two-Dimensional Model}

            \textbf{Objective of Comparison:}
            \begin{itemize}
              \item To validate the efficacy of the Timoshenko beam theory against a more comprehensive two-dimensional model, which accounts for plane stress conditions and detailed stress distribution across the beam's cross-section.
            \end{itemize}

            \textbf{Methodology:}
            \begin{itemize}
              \item Utilized finite element methods to calculate natural frequencies and modes, providing a direct comparison of results from both models.
              \item Both models analyzed under the same boundary conditions and loading scenarios to ensure consistency in comparison.
            \end{itemize}

            \textbf{Key Findings:}
            \begin{itemize}
              \item \textbf{Close Agreement}: Timoshenko model results closely match those from the two-dimensional model, with most eigenvalues showing less than 3\% deviation.
              \item \textbf{Model Reliability}: Demonstrates high reliability of the Timoshenko model for predicting beam behavior under various loading conditions, particularly for lower vibration modes.
            \end{itemize}

        \end{frame}

        \begin{frame}
            \textbf{Model Advantages:}
            \begin{itemize}
              \item \textbf{Timoshenko Model}: Offers significant improvements in accuracy over simpler models like Euler-Bernoulli, especially in scenarios where beam lengths are relatively short and shear deformations cannot be ignored.
              \item \textbf{Two-Dimensional Model}: Provides a more rigorous analysis by accounting for two-dimensional stress effects, which are important for comprehensive structural analyses.
            \end{itemize}

            \textbf{Practical Implications:}
            \begin{itemize}
              \item \textbf{Engineering Applications}: Timoshenko model is particularly advantageous for practical engineering applications, providing a good balance between computational efficiency and accuracy.
              \item \textbf{Advanced Analyses}: For critical engineering tasks where detailed stress distribution is crucial, the two-dimensional model serves as a more suitable analytical tool.
            \end{itemize}

            \textbf{Conclusions and Recommendations:}
            \begin{itemize}
              \item \textbf{Preferred Use of Timoshenko Model}: Recommended for most practical engineering applications involving beams, where it strikes an optimal balance between accuracy and simplicity.
              \item \textbf{When to Use Two-Dimensional Model}: Advisable for in-depth structural analyses where precise stress distribution is required, such as in complex assemblies or where materials with varying properties are involved.
              \item This comparison serves as an intermediate step towards the more comprehensive analysis involving the comparison of two-dimensional and three-dimensional models.
            \end{itemize}

            \end{frame}

    \subsection{Empirical comparison of beams}
        \begin{frame}{Literature study - Empirical comparison of beams}
            Consider the article \textit{On the valid frequency range of Timoshenko beam theory} [\cite{SP06}]\footfullcite{SP06}
        \end{frame}

        \begin{frame}{Setup and Testing Method}
            \textbf{Objective:}
            \begin{itemize}
                \item To validate the frequency predictions of Timoshenko beam theory (TBT) for different end conditions and frequencies.
            \end{itemize}

            \textbf{Methodology:}
            \begin{itemize}
                \item Experimental setup involved a short free-free beam of near-square cross-section.
                \item Beam dimensions: Length = 40.047 mm, Breadth = 10.097 mm, Depth = 10.047 mm.
                \item Beam material: Aluminum alloy, with properties determined via resonant ultrasound spectroscopy (RUS).
                \item Comparisons made with experimental observations, commercial finite element (ANSYS) predictions, and RUS technique predictions.
                \item Finite element analysis used 20-node brick elements, and TBT predictions were calculated using specific frequency equations.
            \end{itemize}

            \textbf{Key Setup Features:}
            \begin{itemize}
                \item The experimental beam was suspended and excited by two carbon-fibre loops attached to piezoelectric transducers.
                \item Resonant frequencies detected by sweeping the frequency range with a network analyzer.
            \end{itemize}
        \end{frame}

        \begin{frame}{Numerical Results and Discussion}
            \textbf{Results Summary:}
            \begin{itemize}
                \item Excellent agreement between measured, ANSYS, and RUS frequencies below the cut-off frequency.
                \item Above the cut-off frequency, TBT predictions deviate increasingly from measured and simulated frequencies.
            \end{itemize}

            \textbf{Discussion:}
            \begin{itemize}
                \item For free-free end conditions, discrepancies arise above the cut-off frequency, suggesting limitations in the TBT's applicability.
                \item TBT predictions were increasingly inaccurate for higher modes (n >= 8), exceeding the typical engineering tolerance.
            \end{itemize}

            \textbf{Conclusion:}
            \begin{itemize}
                \item The study suggests a cut-off frequency as a reasonable upper bound for TBT predictions.
                \item This research serves as an intermediate step, setting the stage for further comparison between two-dimensional and three-dimensional models.
            \end{itemize}
        \end{frame}

\section{Models}
    \subsection{Timoshenko cantilever beam}
        \begin{frame}{Models - Timoshenko cantilever beam}
            \begin{columns}[T] % align columns
                \begin{column}{.6\textwidth}
                    \only<1->{ % Frame 1: Equations of motion
                    Equations of motion
                    \begin{eqnarray*}
                        \partial_{t}^{2} w &=& \partial_{x}V + Q,\\
                        \frac{1}{\alpha} \partial_{t}^{2} \phi &=& V + \partial_{x}M.
                    \end{eqnarray*}
                    % Frame 1: Constitutive equations
                    Constitutive equations
                    \begin{eqnarray*}
                        M &=& \frac{1}{\beta}\partial_x \phi,\\
                        V &=& \partial_x w-\phi.
                    \end{eqnarray*}
                    }
                \end{column}

                \begin{column}{.4\textwidth}
                    \only<2->{% Frame 2: Parameter descriptions
                    \textbf{Description:}
                    \begin{itemize}
                        \item[-] \( w \): Displacement
                        \item[-] \( \phi \): Angle of rotation of the cross-sections
                        \item[-] \( M \): Bending moment
                        \item[-] \( V \): Shear force
                        \item[-] \( Q \): External distributed load
                        \item[-] \( \alpha, \beta \): Dimensionless constants
                    \end{itemize}}
                \end{column}
            \end{columns}
        \end{frame}

        \begin{frame}{Models - Timoshenko cantilever beam}
            \begin{figure}[h!]
                \centering
                \begin{tikzpicture}

                    \draw[line width = 0.4mm] (0,0) -- (6,0);
                    \node at (2.7,-0.25) {$\ell = 1$};

                    \draw[line width = 0.1mm] (0,-1.5) -- (0,1.5);
                    \draw[line width = 0.1mm] (0,1.5) -- (-0.2,1.4);
                    \draw[line width = 0.1mm] (0,1.3) -- (-0.2,1.2);
                    \draw[line width = 0.1mm] (0,1.1) -- (-0.2,1);
                    \draw[line width = 0.1mm] (0,0.9) -- (-0.2,0.8);
                    \draw[line width = 0.1mm] (0,0.7) -- (-0.2,0.6);
                    \draw[line width = 0.1mm] (0,0.5) -- (-0.2,0.4);
                    \draw[line width = 0.1mm] (0,0.3) -- (-0.2,0.2);
                    \draw[line width = 0.1mm] (0,0.1) -- (-0.2,0);
                    \draw[line width = 0.1mm] (0,-0.1) -- (-0.2,-0.2);
                    \draw[line width = 0.1mm] (0,-0.3) -- (-0.2,-0.4);
                    \draw[line width = 0.1mm] (0,-0.5) -- (-0.2,-0.6);
                    \draw[line width = 0.1mm] (0,-0.7) -- (-0.2,-0.8);
                    \draw[line width = 0.1mm] (0,-0.9) -- (-0.2,-1);
                    \draw[line width = 0.1mm] (0,-1.1) -- (-0.2,-1.2);
                    \draw[line width = 0.1mm] (0,-1.3) -- (-0.2,-1.4);
                    \draw[line width = 0.1mm] (0,-1.5) -- (-0.2,-1.6);

                    % Adding points 0 and 1
                    \node at (0,-0.35) {0};
                    \node at (6,-0.35) {1};

                \end{tikzpicture}
                \caption{A cantilever Timoshenko beam.}
            \end{figure}

            Boundary conditions for a cantilever beam
            \begin{eqnarray*}
                w(0,\cdot) = 0, &\phi(0,\cdot) = 0,\\
                M(1,\cdot) = 0, &V(1,\cdot) = 0.
            \end{eqnarray*}
        \end{frame}

    \subsection{Two-dimensional cantilever beam}
        \begin{frame}{Models - Two-dimensional cantilever beam}
            \begin{columns}[T] % align columns
                \begin{column}{.6\textwidth}
                    \only<1->{ % Frame 1: Equations of motion
                        Equations of motion
                        \begin{eqnarray*}
                            \partial_t^2 u & = & \textrm{div}T + Q,
                        \end{eqnarray*}
                        with
                        \begin{eqnarray*}
                            \textrm{div} T & = &
                            \begin{bmatrix}
                                \partial_1 \sigma_{11} + \partial_2 \sigma_{12} \\
                                \partial_1 \sigma_{21} + \partial_2 \sigma_{22}
                            \end{bmatrix}.
                        \end{eqnarray*}
                    }
                    \only<1->{ % Frame 1: Constitutive equations
                        Constitutive equations
                        \begin{eqnarray*}
                            T = \frac{1}{\gamma(1+\nu)}\mathcal{E} + \frac{\nu}{\gamma(1-\nu^2)}\textrm{tr}(\mathcal{E}) I
                        \end{eqnarray*}
                    }
                \end{column}

                \begin{column}{.4\textwidth}
                    \only<2->{ % Frame 2: Parameter descriptions
                        \textbf{Description:}
                        \begin{itemize}
                            \item[-] \( u \): Displacement vector in 2D
                            \item[-] \( T \): Stress tensor
                            \item[-] \( Q \): External distributed load in 2D
                            \item[-] \( \mathcal{E} \): Young's modulus
                            \item[-] \( \nu \): Poisson's ratio
                            \item[-] \( \gamma \): Dimensionless constant
                        \end{itemize}
                    }
                \end{column}
            \end{columns}
        \end{frame}

        \begin{frame}{Models - Two-dimensional cantilever beam}
            \begin{figure}[h!]
                \centering
                \begin{tikzpicture}

                    \draw[line width = 0.4mm] (0,0.5) -- (7,0.5);
                    \draw[line width = 0.4mm] (0,-0.5) -- (7,-0.5);
                    \draw[line width = 0.4mm] (7,-0.5) -- (7,0.5);
                    \draw[line width = 0.4mm] (0,-0.5) -- (0,0.5);

                    \draw[line width = 0.1mm] (0,1.5) -- (0,-1.5);

                    \draw[line width = 0.1mm] (0,1.5) -- (-0.2,1.3);
                    \draw[line width = 0.1mm] (0,1.3) -- (-0.2,1.1);
                    \draw[line width = 0.1mm] (0,1.1) -- (-0.2,0.9);
                    \draw[line width = 0.1mm] (0,0.9) -- (-0.2,0.7);
                    \draw[line width = 0.1mm] (0,0.7) -- (-0.2,0.5);
                    \draw[line width = 0.1mm] (0,0.5) -- (-0.2,0.3);
                    \draw[line width = 0.1mm] (0,0.3) -- (-0.2,0.1);
                    \draw[line width = 0.1mm] (0,0.1) -- (-0.2,-0.1);
                    \draw[line width = 0.1mm] (0,-0.1) -- (-0.2,-0.3);
                    \draw[line width = 0.1mm] (0,-0.3) -- (-0.2,-0.5);
                    \draw[line width = 0.1mm] (0,-0.5) -- (-0.2,-0.7);
                    \draw[line width = 0.1mm] (0,-0.7) -- (-0.2,-0.9);
                    \draw[line width = 0.1mm] (0,-0.9) -- (-0.2,-1.1);
                    \draw[line width = 0.1mm] (0,-1.1) -- (-0.2,-1.3);
                    \draw[line width = 0.1mm] (0,-1.3) -- (-0.2,-1.5);

                    \node at (7.3,0) {$h$};
                    \draw[-stealth] (7.3,0.2) -- (7.3,0.5);
                    \draw (7.2,0.5) -- (7.4,0.5);
                    \draw[-stealth] (7.3,-0.2) -- (7.3,-0.5);
                    \draw (7.2,-0.5) -- (7.4,-0.5);

                    \node at (3.5,-0.8) {$1$};
                    \draw[stealth-] (0,-0.8) -- (3.3,-0.8);
                    \draw[stealth-] (7,-0.8) -- (3.7,-0.8);
                    \draw (7,-0.9) -- (7,-0.7);

                    \draw[line width = 0.1mm,->] (9,1) -- (10.4,1);
                    \draw[line width = 0.1mm,->] (9,1) -- (9,2.1);
                    \node at (9.7,0.8) {$e_1$};
                    \node at (8.8,1.5) {$e_2$};

                    \node at (-0.6,0) {$(0,0)$};
                    \node at (0,0)[circle,fill,inner sep=1.5pt]{};

                \end{tikzpicture}
                \caption{A cantilever two-dimensional elastic body.}
            \end{figure}

            % Placeholder for boundary conditions of 2D cantilever
            Boundary Conditions:
            \begin{eqnarray*}
                u & = & 0 \quad \textrm{ where } x_1 = 0\\
                Tn & = & 0 \quad \textrm{ on the rest of the body }
            \end{eqnarray*} With $n$ the outward normal vector to $\Omega$.
        \end{frame}

    \subsection{Three-dimensional cantilever beam}
        \begin{frame}{Models - Three-dimensional cantilever beam}
            \begin{columns}[T] % align columns
                \begin{column}{.6\textwidth}
                    \only<1->{ % Frame 1: Equations of motion
                        Equations of motion
                        \begin{eqnarray*}
                            \partial_t^2 u & = & \textrm{div}T + Q
                        \end{eqnarray*}
                        with
                        \begin{eqnarray*}
                            \textrm{div}  T & = &
                            \begin{bmatrix}
                                \partial_1 \sigma_{11} + \partial_2 \sigma_{12} + \partial_3 \sigma_{13} \\
                                \partial_1 \sigma_{21} + \partial_2 \sigma_{22} + \partial_3 \sigma_{23} \\
                                \partial_1 \sigma_{31} + \partial_2 \sigma_{32} + \partial_3 \sigma_{33}
                            \end{bmatrix}.\label{eq:3D_Model:divT-D}
                        \end{eqnarray*}
                    }
                    \only<1->{ % Frame 1: Constitutive equations
                        Constitutive equations
                        \begin{eqnarray*}
                            T = \frac{1}{\gamma(1+\nu)} \mathcal{E} + \frac{\nu}{\gamma(1+\nu)(1-2\nu)}\textrm{Tr}(\mathcal{E})I
                        \end{eqnarray*}
                    }
                \end{column}

                \begin{column}{.4\textwidth}
                    \only<2->{ % Frame 2: Parameter descriptions
                        \textbf{Description:}
                        \begin{itemize}
                            \item[-] \( u \): Displacement vector in 2D
                            \item[-] \( T \): Stress tensor
                            \item[-] \( Q \): External distributed load in 2D
                            \item[-] \( \mathcal{E} \): Young's modulus
                            \item[-] \( \nu \): Poisson's ratio
                            \item[-] \( \gamma \): Dimensionless constant
                        \end{itemize}
                    }
                \end{column}
            \end{columns}
        \end{frame}

        \begin{frame}{Models - Three-dimensional cantilever beam}
            \begin{figure}[h!]
                \centering
                \begin{tikzpicture}[scale=0.58]

                    \draw[line width = 0.4mm] (-0.5,1) -- (6.5,1);
                    \draw[line width = 0.4mm] (-0.5,-1) -- (6.5,-1);
                    \draw[line width = 0.4mm] (6.5,-1) -- (6.5,1);
                    \draw[line width = 0.4mm] (-0.5,-1) -- (-0.5,1);

                    \draw[line width = 0.4mm] (0.5,1.5) -- (7.5,1.5);
                    \draw[line width = 0.4mm] (7.5,-0.5) -- (7.5,1.5);

                    \draw[line width = 0.4mm] (-0.5,1) -- (0.5,1.5);
                    \draw[line width = 0.4mm] (6.5,1) -- (7.5,1.5);
                    \draw[line width = 0.4mm] (6.5,-1) -- (7.5,-0.5);

                    \draw[scale=0.5, domain=-3:3, smooth, variable=\x,dotted] plot ({\x}, {0.5*\x+4});
                    \draw[scale=0.5, domain=-3:3, smooth, variable=\x,dotted] plot ({\x}, {0.5*\x+3.5});
                    \draw[scale=0.5, domain=-3:3, smooth, variable=\x,dotted] plot ({\x}, {0.5*\x+3});
                    \draw[scale=0.5, domain=-3:3, smooth, variable=\x,dotted] plot ({\x}, {0.5*\x+2.5});

                    \draw[scale=0.5, domain=-3:-1, smooth, variable=\x,dotted] plot ({\x}, {0.5*\x+2});
                    \draw[scale=0.5, domain=2:3, smooth, variable=\x,dotted] plot ({\x}, {0.5*\x+2});

                    \draw[scale=0.5, domain=-3:-1, smooth, variable=\x,dotted] plot ({\x}, {0.5*\x+1.5});
                    \draw[scale=0.5, domain=-3:-1, smooth, variable=\x,dotted] plot ({\x}, {0.5*\x+1});
                    \draw[scale=0.5, domain=-3:-1, smooth, variable=\x,dotted] plot ({\x}, {0.5*\x+0.5});
                    \draw[scale=0.5, domain=-3:-1, smooth, variable=\x,dotted] plot ({\x}, {0.5*\x});
                    \draw[scale=0.5, domain=-3:-1, smooth, variable=\x,dotted] plot ({\x}, {0.5*\x-0.5});
                    \draw[scale=0.5, domain=-3:-1, smooth, variable=\x,dotted] plot ({\x}, {0.5*\x-1});
                    \draw[scale=0.5, domain=-3:-1, smooth, variable=\x,dotted] plot ({\x}, {0.5*\x-1.5});
                    \draw[scale=0.5, domain=-3:0, smooth, variable=\x,dotted] plot ({\x}, {0.5*\x-2});
                    \draw[scale=0.5, domain=-3:1, smooth, variable=\x,dotted] plot ({\x}, {0.5*\x-2.5});
                    \draw[scale=0.5, domain=-3:2, smooth, variable=\x,dotted] plot ({\x}, {0.5*\x-3});
                    \draw[scale=0.5, domain=-3:3, smooth, variable=\x,dotted] plot ({\x}, {0.5*\x-3.5});
                    \draw[scale=0.5, domain=-3:3, smooth, variable=\x,dotted] plot ({\x}, {0.5*\x-4});

                    %\node at (6.9,1) {$b$};
                    %\node at (6.65,0) {$h$};
                    %\node at (3.2,-1.2) {$\ell = 1$};

                    \draw[line width = 0.1mm,->] (9,1) -- (10,1.6);
                    \draw[line width = 0.1mm,->] (9,1) -- (10.4,1);
                    \draw[line width = 0.1mm,->] (9,1) -- (9,2.1);
                    \node at (9.3,1.4) {$e_3$};
                    \node at (9.7,0.8) {$e_1$};
                    \node at (8.8,1.5) {$e_2$};

                    \node at (7.7,0.5) {$h$};
                    \draw[-stealth] (7.7,0.7) -- (7.7,1.5);
                    \draw (7.6,1.5) -- (7.8,1.5);
                    \draw[-stealth] (7.7,0.3) -- (7.7,-0.5);
                    \draw (7.6,-0.5) -- (7.8,-0.5);

                    \node at (3,-1.2) {$1$};
                    \draw[stealth-] (-0.5,-1.2) -- (2.8,-1.2);
                    \draw[stealth-] (6.5,-1.2) -- (3.2,-1.2);
                    \draw (6.5,-1.3) -- (6.5,-1.1);
                    \draw (-0.5,-1.3) -- (-0.5,-1.1);

                    \node at (7.2,-0.9) {$b$};
                    \draw[stealth-] (7.7,-0.5) -- (7.35,-0.75);
                    \draw[stealth-] (6.5,-1.2) -- (7,-0.95);

                    \draw[line width = 0.1mm,->] (9,1) -- (10.4,1);
                    \draw[line width = 0.1mm,->] (9,1) -- (9,2.1);
                    \node at (9.7,0.8) {$e_1$};
                    \node at (8.8,1.5) {$e_2$};

                    \node at (-0.6,0) {$(0,0)$};
                    \node at (0,0)[circle,fill,inner sep=1.5pt]{};

                    %\node at (-1.4,-1.3) {$(0,-\frac{h}{2},-\frac{b}{2})$};
                    %\node at (-1.4,1) {$(0,\frac{h}{2},-\frac{b}{2})$};
                    %\node at (0.5,1.8) {$(0,\frac{h}{2},\frac{b}{2})$};

                    %\node at (6,-1.3) {$(1,-\frac{h}{2},-\frac{b}{2})$};
                    %\node at (6.3,0.7) {$(1,\frac{h}{2},-\frac{b}{2})$};
                    %\node at (7.5,1.8) {$(1,\frac{h}{2},\frac{b}{2})$};
                    %\node at (8.4,-0.4) {$(1,-\frac{h}{2},\frac{b}{2})$};

                \end{tikzpicture}
                \caption{Cantilever Three-Dimensional Elastic Body with Rectangular Cross-Section.}
            \end{figure} 
            Boundary Conditions:
            \begin{eqnarray*}
                u & = & 0 \quad \textrm{ where } x_1 = 0\\
                Tn & = & 0 \quad \textrm{ on the rest of the body }
            \end{eqnarray*} With $n$ the outward normal vector to $\Omega$.
        \end{frame}

\section{Theory}
    \subsection{General vibration problem}
        \begin{frame}{General vibration problem}
            Let $V$, $W$ and $X$ be real Hilbert spaces such that $W$ is a linear subspace of $X$, and $V$ is a linear subspace of $W$, i.e. $V \subset W \subset X$.
            \begin{itemize}
                \item[] X is the global space with inner product $(\cdot,\cdot)_X$ and the induced norm $||\cdot||_X$.
                \item[] W is the inertia space with inner product $(\cdot,\cdot)_W$ and the induced norm $||\cdot||_W$.
                \item[] V is the energy space with inner product $(\cdot,\cdot)_V$ and the induced norm $||\cdot||_V$.
            \end{itemize}
            \only<2->{
            The bilinear forms $a$, $b$ and $c$ are defined as $a = (\cdot,\cdot)_X$,$b = (\cdot,\cdot)_V$ and $c = (\cdot,\cdot)_W$.
            }

            \only<3>{
            \textbf{General vibration problem:}

            Find a function $u \in C(J,\ X)$ such that $u'$ is continuous at $0$ with respect to $\Vert \cdot \Vert_{W}$, and for each $t \in J$, $u(t) \in V$, $u'(t) \in V$, $u''(t) \in W$, satisfying
            \begin{eqnarray}
                c(u''(t),v)+a(u'(t),v)+b(u(t),v) = (f(t),v) \ \ \ \ \textrm{for each} \ v \in V, \label{eq:existence:ProblemGVarHom}
            \end{eqnarray}
            with initial conditions $u(0) = u_0$ and $u'(0) = u_1$.}

            \only<4->{
            \textbf{General vibration problem:}

            Find a function $u \in C(J,\ X)$ such that $u'$ is continuous at $0$ with respect to $\Vert \cdot \Vert_{W}$, and for each $t \in J$, $u(t) \in V$, $u'(t) \in V$, $u''(t) \in W$, satisfying
            \begin{eqnarray}
                c(u''(t),v)+b(u(t),v) = (f(t),v) \ \ \ \ \textrm{for each} \ v \in V, \label{eq:existence:ProblemGVarHom}
            \end{eqnarray}
            with initial conditions $u(0) = u_0$ and $u'(0) = u_1$.}

            \footfullcitenonumber{VV02}
        \end{frame}

    \subsection{Existence and uniqueness of solutions}
        \begin{frame}{Existence and uniqueness of solutions}
            In [\cite{VV02}] the authors present results for the existence and uniqueness of solutions to the general vibration problem. The following assumptions are made for the existence results.

            \begin{itemize}
                \item[] \textbf{A1} - $V$ is dense in $W$ and $W$ is dense in $X$.

                \item[] \textbf{A2} - There exists a positive constant $C_{W}$ such that $\Vert w\Vert_{X} \leq C_{W}\Vert w\Vert_{W}$ for each $ w\in W$.

                \item[] \textbf{A3} - There exists a positive constant $C_{V}$ such that $\Vert v\Vert_{W} \leq C_{V}\Vert v\Vert_{V}$ for each $v \in V$.

                \item[] \textbf{A4} - The bilinear form $a$ is non-negative, symmetric and bounded on $V$, i.e. there exists a positive constant $K_a$ such that for $\displaystyle u,v \in V$, \[|a(u,v)| \leq K_a\Vert u \Vert_V \Vert v \Vert_V.\]
            \end{itemize}
            \footfullcitenonumber{VV02}
        \end{frame}

        \begin{frame}{Existence and uniqueness of solutions}
            \newtheorem{Thmx}{Theorem}
            \begin{Thmx}
                Suppose assumptions \textbf{A1}-\textbf{A4} hold. If, for $u_0 \in V$ and $u_1 \in V$, there exists some $y \in W$ such that
                \begin{eqnarray}
                    b(u_0,v) + a(u_1,v) = c(y,v) \ \ \ \ \textrm{ for all } \ v \in V, \label{bilinear_equation}
                \end{eqnarray}
                then for each $f \in C^1(J,X)$, there exists a unique solution $u \in C^1(J,V)\cap C^2(J,W)$ for Problem GVar.
            \end{Thmx}

            This theorem allows for a solution of the abstract variational problem Problem GVar, if the assumptions \textbf{A1}-\textbf{A4} is satisfied and the initial values $u_0$ and $u_1$ are admissible. It is not always easy to verify that \eqref{bilinear_equation} is satisfied.

            \footfullcitenonumber{VV02}
        \end{frame}

    \subsection{Modal Analysis}
        \begin{frame}{Modal Analysis}
            \textbf{General vibration problem:}

            Find a function $u \in C(J,\ X)$ such that $u'$ is continuous at $0$ with respect to $\Vert \cdot \Vert_{W}$, and for each $t \in J$, $u(t) \in V$, $u'(t) \in V$, $u''(t) \in W$, satisfying
            \begin{eqnarray}
                c(u''(t),v)+b(u(t),v) = 0 \ \ \ \ \textrm{for each} \ v \in V, \label{eq:existence:ProblemGVarHom}
            \end{eqnarray}
            with initial conditions $u(0) = u_0$ and $u'(0) = u_1$.

            \only<2->{
            Trial solution:

            $u(t) = T(t)x$ with $x \in V$ and $x \neq 0$}
        \end{frame}

        \begin{frame}{Modal Analysis}
           Substituting trial solution into \eqref{eq:existence:ProblemGVarHom}:
           \begin{eqnarray*}
                b(T(t)x,v) = - c(T''(t)x,v).  \label{eq:existence:ProblemGVarHom:Substitution}
            \end{eqnarray*}

            Due to the linearity of the bilinear form, this equation can be rewritten as
            \begin{eqnarray*}
                T(t)b(x,v) = - T''(t)c(x,v).
            \end{eqnarray*}

            \only<1>{
            Dividing both sides by $T(t)$ gives
            \begin{eqnarray*}
                b(x,v) = - \frac{T''(t)}{T(t)}c(x,v).
            \end{eqnarray*}}

            \only<2>{
            Dividing both sides by $T(t)$ gives
            \begin{eqnarray*}
                b(x,v) = - \frac{T''(t)}{T(t)}c(x,v) \implies  \frac{T''(t)}{T(t)} \ \ \ \textrm{must be constant.}
            \end{eqnarray*}}

            \only<3->{
                Set $\displaystyle \frac{T''(t)}{T(t)} = -\lambda$.

            At this point we do not yet know if such a $\lambda$ exists.
            }
        \end{frame}

        \begin{frame}{Modal Analysis}
            Eigenvalue Problem:

            Find a real number $\lambda$ and a $x \in V$ with $x \neq 0$ such that
            \begin{eqnarray*}
                b(x,y) = \lambda c(x,y) \ \ \ \ \textrm{ for each } \ y \in V.
            \end{eqnarray*}

            Ordinary differential equation for $T(t)$:
            \begin{eqnarray}
                T''(t)  + \lambda T(t) = 0. \label{eq:1D_Model:ModalAnalysisODE}
            \end{eqnarray}
         \end{frame}

        \begin{frame}{Modal Analysis}
            Assumptions:
            \begin{itemize}
                \item[] \textbf{A1} - $V$ is dense in $W$ and $W$ is dense in $X$.

                \item[] \textbf{A2} - There exists a positive constant $C_{W}$ such that $\Vert w\Vert_{X} \leq C_{W}\Vert w\Vert_{W}$ for each $ w\in W$.

                \item[] \textbf{A3} - There exists a positive constant $C_{V}$ such that $\Vert v\Vert_{W} \leq C_{V}\Vert v\Vert_{V}$ for each $v \in V$.

                \item[] \textbf{A4} - The bilinear form $a$ is non-negative, symmetric and bounded on $V$, i.e. there exists a positive constant $K_a$ such that for $\displaystyle u,v \in V$, \[|a(u,v)| \leq K_a\Vert u \Vert_V \Vert v \Vert_V.\]

                \item[] \alert{\textbf{A5} - The embedding of $V$ into $W$ is compact.} [\cite{CVV18}]
            \end{itemize}
            \footfullcite{CVV18}
        \end{frame}

        \begin{frame}{Modal Analysis}

            ***************Verduideliking********************

            Solutions for the Eigenvalue Problem:
            \begin{eqnarray*}
                u = \sum_{i=1}^{\infty} a_i x_i = \sum_{i=1}^{\infty} c(u, x_i)x_i.
            \end{eqnarray*}

            Solutions for the Ordinary Differential Equation:
            \begin{flalign}
                T_n(t) &=  A_n \cos(\sqrt{\lambda_n} t) + B_n \sin(\sqrt{\lambda_n} t) \quad \textrm{ if } \lambda_n > 0, \label{lambda_1}
            \end{flalign}

            Formal series solution for the general vibration problem:
            \begin{eqnarray}
                u(t) = \sum_{n=1}^{\infty} T_n(t)x_n.
            \end{eqnarray}
        \end{frame}

\section{Numerical Results}
    \subsection{Finite Element Analysis}
        \begin{frame}{Eigenvalue problem - Two and Three-dimensional models}
            General eigenvalue problem for the two- and three-dimensional models:

                For the eigenvalue problem, assume that $M_{f}Q^I = 0$ so that 
            \begin{eqnarray}
                    M\ddot{\bar{u}} & = & K\bar{u}.\label{eq:2DFEM:M2}
            \end{eqnarray}

            Not so simple to compute the eigenvalues and eigenfunctions.

            Need to use a numerical method like FEM to approximate the eigenvalues and eigenvectors.

        \end{frame}

        \begin{frame}{Eigenvalue problem - Timoshenko model}
            General eigenvalue problem for the Timoshenko beam model:

            Find functions $u$ and $\phi$ and a real number $\lambda$ satisfying the following equations
            \begin{eqnarray}
            -u'' + \phi' &=& \lambda u, \label{eq:Timo:EigenvalueProblem1}\\
            -\alpha u' + \alpha\phi - \frac{1}{\gamma}\phi'' &=& \lambda\phi.\label{eq:Timo:EigenvalueProblem2}
            \end{eqnarray}

            The solution to this problem is presented in the article [\cite{VV06}]\footfullcite{VV06}.

            \begin{itemize}
                \item[-] Provides a method to calculate the exact eigenvalues.
                \item[-] Uses a cantilever beam as an example in the article.
                \item[-] Eigenvalues can be numerically obtained to any degree of accuracy.
            \end{itemize}

        \end{frame}

        \begin{frame}{FEM}
        \end{frame}

        \begin{frame}{Accuracy of the Eigenvalues}
            \centering
            \includegraphics[scale=0.3]{Convergence.png} \\ % Adjust scale as needed
            \footnotesize Rate of convergence of the first 20 eigenvalues.

            \centering
            \includegraphics[scale=0.3]{convergence3d.png} \\ % Adjust scale as needed
            \footnotesize Rate of convergence of the first 20 eigenvalues.
        \end{frame}

    \subsection{Validity of the Timoshenko beam model}
        \begin{frame}{Validity of the Timoshenko beam model}
            \begin{center}
                \begin{minipage}[b]{0.45\linewidth}
                    \centering
                    \begin{tikzpicture}[scale=0.7]
                        \draw[line width = 0.4mm] (0,0.2) -- (6,0.2);
                                    \draw[line width = 0.4mm] (0,-0.2) -- (6,-0.2);
                                    \draw[line width = 0.4mm] (6,-0.2) -- (6,0.2);

                                    \node at (6.2,0) {$h$};
                                    \node at (2.7,-0.45) {$\ell = 1$};

                                    \draw[line width = 0.1mm] (0,-1.5) -- (0,1.5);
                                    \draw[line width = 0.1mm] (0,1.5) -- (-0.2,1.4);
                                    \draw[line width = 0.1mm] (0,1.3) -- (-0.2,1.2);
                                    \draw[line width = 0.1mm] (0,1.1) -- (-0.2,1);
                                    \draw[line width = 0.1mm] (0,0.9) -- (-0.2,0.8);
                                    \draw[line width = 0.1mm] (0,0.7) -- (-0.2,0.6);
                                    \draw[line width = 0.1mm] (0,0.5) -- (-0.2,0.4);
                                    \draw[line width = 0.1mm] (0,0.3) -- (-0.2,0.2);
                                    \draw[line width = 0.1mm] (0,0.1) -- (-0.2,0);

                                    \draw[line width = 0.1mm] (0,-0.1) -- (-0.2,-0.2);
                                    \draw[line width = 0.1mm] (0,-0.3) -- (-0.2,-0.4);
                                    \draw[line width = 0.1mm] (0,-0.5) -- (-0.2,-0.6);
                                    \draw[line width = 0.1mm] (0,-0.7) -- (-0.2,-0.8);
                                    \draw[line width = 0.1mm] (0,-0.9) -- (-0.2,-1);
                                    \draw[line width = 0.1mm] (0,-1.1) -- (-0.2,-1.2);
                                    \draw[line width = 0.1mm] (0,-1.3) -- (-0.2,-1.4);
                                    \draw[line width = 0.1mm] (0,-1.5) -- (-0.2,-1.6);
                    \end{tikzpicture}
                    \captionof{figure}{Two-Dimensional Beam}
                \end{minipage}
                \hfill
                \begin{minipage}[b]{0.45\linewidth}
                    \centering
                    \begin{tikzpicture}[scale=0.7]
                        \draw[line width = 0.4mm] (0,0) -- (6,0);
                                \node at (2.7,-0.25) {$\ell = 1$};

                                \draw[line width = 0.1mm] (0,-1.5) -- (0,1.5);
                                \draw[line width = 0.1mm] (0,1.5) -- (-0.2,1.4);
                                \draw[line width = 0.1mm] (0,1.3) -- (-0.2,1.2);
                                \draw[line width = 0.1mm] (0,1.1) -- (-0.2,1);
                                \draw[line width = 0.1mm] (0,0.9) -- (-0.2,0.8);
                                \draw[line width = 0.1mm] (0,0.7) -- (-0.2,0.6);
                                \draw[line width = 0.1mm] (0,0.5) -- (-0.2,0.4);
                                \draw[line width = 0.1mm] (0,0.3) -- (-0.2,0.2);
                                \draw[line width = 0.1mm] (0,0.1) -- (-0.2,0);

                                \draw[line width = 0.1mm] (0,-0.1) -- (-0.2,-0.2);
                                \draw[line width = 0.1mm] (0,-0.3) -- (-0.2,-0.4);
                                \draw[line width = 0.1mm] (0,-0.5) -- (-0.2,-0.6);
                                \draw[line width = 0.1mm] (0,-0.7) -- (-0.2,-0.8);
                                \draw[line width = 0.1mm] (0,-0.9) -- (-0.2,-1);
                                \draw[line width = 0.1mm] (0,-1.1) -- (-0.2,-1.2);
                                \draw[line width = 0.1mm] (0,-1.3) -- (-0.2,-1.4);
                                \draw[line width = 0.1mm] (0,-1.5) -- (-0.2,-1.6);
                    \end{tikzpicture}
                    \captionof{figure}{Timoshenko Beam}
                \end{minipage}
            \end{center}
        \end{frame}

        \begin{frame}{Comparison of two and three-dimensional models}
            \centering
            \begin{minipage}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{Beam1.jpg}
                \\ 2D Beam Type - $\lambda_6 = 21.911$
                \label{fig:minipage2}
            \end{minipage}
            \hfill
            \begin{minipage}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{1DBeam2.png}
                \\ Timoshenko - $\lambda_5 = 21.794$
                \label{fig:minipage1}
            \end{minipage}
            \vspace{1em}
            \begin{minipage}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{Beam2.jpg}
                \\ 2D Beam Type - $\lambda_7 = 45.711$
                \label{fig:minipage4}
            \end{minipage}
            \hfill
            \begin{minipage}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{1DBeam1.png}
                \\ Timoshenko - $\lambda_6 = 45.390$
                \label{fig:minipage3}
            \end{minipage}
            \\ Modal shapes of the displacement $w$ for the beam-type 2D body and the Timoshenko beam with $\alpha = 4800$ ($h = 1/20$).
        \end{frame}

        \begin{frame}{Comparison of two and three-dimensional models}
            \centering
            \begin{minipage}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{NotBeam1.jpg}
                \\ Non-Beam Type - $\lambda_4 = 7.7077$
                \label{fig:nonbeam1}
            \end{minipage}
            \hfill
            \begin{minipage}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{NotBeam2.jpg}
                \\ Non-Beam Type - $\lambda_8 = 69.344$
                \label{fig:nonbeam2}
            \end{minipage}
            \\ Modal shapes of the displacement $w$ for the non-beam type 2D body with $h = 1/20$.
        \end{frame}

        \begin{frame}{Comparison of two and three-dimensional models}
            \centering
            \footnotesize
            \renewcommand{\arraystretch}{0.9}
            Results from [LVV09] and results obtained. $0^a$ indicates a 0 due to rounding. $\alpha = 1200$

            \begin{tabular}{cccc|cccc}
                \hline
                & \multicolumn{3}{c}{[LVV09]} & & \multicolumn{3}{c}{Dissertation} \\

                & 2D & Timo & RE & & 2D & Timo & RE \\
                \hline
                1 & 0.0317 & 0.0316 & 0.00315 & 1 & 0.031713 & 0.031639 & 0.0023407 \\
                2 & 1.14 & 1.14 & $0^a$ & 2 & 1.1413 & 1.1365 & 0.0042050 \\
                3 & 7.72 & - & - & 3 & 7.7161 & - & - \\
                4 & 7.92 & 7.86 & 0.00758 & 4 & 7.918 & 7.8617 & 0.0071116 \\
                5 & 26.2 & 25.9 & 0.0115 & 5 & 26.148 & 25.869 & 0.010669 \\
                6 & 60.8 & 59.9 & 0.0148 & 6 & 60.816 & 59.946 & 0.014303 \\
                7 & 69.3 & - & - & 7 & 69.344 & - & - \\
                8 & 115 & 113 & 0.0174 & 8 & 115.28 & 113.23 & 0.017787 \\
                9 & 192 & 188 & 0.0208 & 9 & 191.57 & 187.55 & 0.020999 \\
                10 & 192 & - & - & 10 & 192.03 & - & - \\
                11 & 291 & 284 & 0.0241 & 11 & 290.76 & 283.81 & 0.023899 \\
                \hline
            \end{tabular}

        \end{frame}

        \begin{frame}{Eigenvalue problem - FEM}
            \centering
            \includegraphics[scale=0.45]{Result.png} \\ % Adjust scale as needed
            \footnotesize Rate of convergence of the first 20 eigenvalues.
            \label{fig:conv}
        \end{frame}

        \begin{frame}{Comparison of Eigenvalues across Different Parameters}
            \centering
            \footnotesize
            \renewcommand{\arraystretch}{0.8}
            \resizebox{\textwidth}{!}{%
            \begin{tabular}{|cccc||cccc||cccc||cccc|}
                \hline
                \multicolumn{16}{|c|}{Comparison of Eigenvalues} \\
                \hline\hline
                \multicolumn{4}{|c||}{ $h = 1/5$ or $\alpha = 300$} & \multicolumn{4}{c||}{$h =1/10$ or $\alpha = 1200$} & \multicolumn{4}{c||}{$h = 1/20$ or $\alpha = 4800$} & \multicolumn{4}{c|}{$h = 1/30$ or $\alpha = 10800$} \\
                \hline

                {i} & {2D} & {j} & {Timo} & {i} & {2D} & {j} & {Timo} & {i} & {2D} & {j} & {Timo} & {i} & {2D} & {j} & {Timo} \\
                \hline
                {1}&0.12151&1&0.12092&1&0.031713&1&0.031639&1&0.008013&1&0.008004&1&0.003568&1&0.003565\\
                {2}&3.5460&2&3.5071&2&1.1413&2     & 1.1365 & 2     & 0.30756 & 2     & 0.30705 & 2     & 0.13869 & 2     & 0.13855 \\
                {3} & {7.7311} &       & {-} & {3}     & {7.7161} &       & {-} & 3     & 2.3273 & 3     & 2.3213 & 3     & 1.0698 & 3     & 1.0683 \\
                {4} & 20.225 & 3     & 19.869 & 4     & 7.9180 & 3     & 7.8617 & {4}     & {7.7077} &       & {-} & 4     & 4.0140 & 4     & 4.0058 \\
                {5} & 56.109 & 4     & 54.766 & 5     & 26.148 & 4     & 25.869 & 5     & 8.5086 & 4     & 8.4762 & {5}     & {7.7047} &       & {-} \\
                {6} & {69.164} &       & {-} & 6     & 60.816 & 5     & 59.946 & 6     & 21.911 & 5     & 21.794 & 6     & 10.655 & 5     & 10.625 \\
                {7} & 114.03 & 5     & 110.75 & {7}     & {69.344} &       & {-} & 7     & 45.711 & 6     & 45.390 & 7     & 22.975 & 6     & 22.890 \\
                {8} & {189.17} &  6     & {186.50} & 8     & 115.28 & 6     & 113.23 & {8}     & {69.344} &       & {-} & 8     & 43.113 & 7     & 42.909 \\
                {9} & {192.61} &      &  & {9}     & {191.57} &   7  & {187.55} & 9     & 82.887 & 7     & 82.154 & {9}     & {69.331} &       & {-} \\
                {10} & 285.85 & 7     & 277.64 & {10}    & {192.03} &      & {-} & 10    & 136.03 & 8     & 134.58 & 10    & 73.230 & 8     & 72.803 \\
                {11} & 328.40 & 8     & 330.29 & 11    & 290.76 & 8     & 283.81 & {11}    & {192.48} &       & {-} & 11    & 115.41 & 9     & 114.61 \\
                {12} & {357.08} &       & {-} & {12}    & {374.45} &       & {-} & 12    & 207.29 & 9     & 204.69 & 12    & 171.61 & 10    & 170.20 \\
                {13} & 397.33 & 9     & 394.02 & 13    & 413.20 & 9     & 402.27 & 13    & 298.38 & 10    & 294.10 & {13}    & {192.52} &       & {-} \\
                {14} & 442.00   & 10    & 439.52 & 14    & 558.67 & 10    & 542.65 & {14}    & {376.83} &       & {-} & 14    & 243.56 & 11    & 241.26 \\
                {15} & {533.71} &       & {-} & {15}    & {614.11} &       & {-} & 15    & 410.63 & 11    & 404.01 & 15    & 332.83 & 12    & 329.28 \\
                {16} & 538.97 & 11    & 541.55 & 16    & 726.26 & 11    & 704.15 & 16    & 545.03 & 12    & 535.32 & {16}    & {377.16} &       & {-} \\
                {17} & {596.06} &       & {-} & {17}    & {906.28} &       & {-} & {17}    & {621.95} &       & {-} & 17    & 440.77 & 13    & 435.51 \\
                {18} & 602.77 & 12    & 596.09 & 18    & 913.69 & 12    & 884.92 & 18    & 702.30 & 13    & 688.64 & 18    & 568.51 & 14    & 561.04 \\
                {19} & {657.87} &       & {-} & 19    & 1113.7 & 13    & 1080.1 & 19    & 882.95 & 14    & 864.40 & {19}    & {623.05} &       & {-} \\
                {20} & 717.37 & 13    & 731.74 & {20}    & {1218.0}  &       & {-} & {20}    & {927.18} &       & {-} & 20    & 717.04 & 15    & 706.74 \\
                \hline
                \multicolumn{2}{|c}{Max RE:} & \multicolumn{2}{c||}{3.1718\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c||}{3.1486\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c||}{2.1018\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c|}{1.4361\%} \\
                \hline
            \end{tabular}}
        \end{frame}

    \subsection{Validity of the Two-dimensional beam model}
        \begin{frame}{Validity of the Two-dimensional beam model}
            \begin{center}
                \begin{minipage}[b]{0.45\linewidth}
                    \centering
                    \begin{tikzpicture}[scale=0.6]
                        \draw[line width = 0.3mm] (-0.1,1) -- (6.9,1);
                        \draw[line width = 0.3mm] (-0.1,-1) -- (6.9,-1);
                        \draw[line width = 0.3mm] (6.9,-1) -- (6.9,1);
                        \draw[line width = 0.3mm] (-0.1,-1) -- (-0.1,1);

                        \draw[line width = 0.3mm] (0.1,1.1) -- (7.1,1.1);
                        \draw[line width = 0.3mm] (7.1,-0.9) -- (7.1,1.1);

                        \draw[line width = 0.3mm] (-0.1,1) -- (0.1,1.1);
                        \draw[line width = 0.3mm] (6.9,1) -- (7.1,1.1);
                        \draw[line width = 0.3mm] (6.9,-1) -- (7.1,-0.9);

                        \draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+4});
                        \draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+3.5});
                        \draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+3});
                        \draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+2.5});

                        \draw[scale=0.5, domain=-3:-0.1, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+2});
                        \draw[scale=0.5, domain=-3:-0.1, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1.5});
                        \draw[scale=0.5, domain=-3:-0.1, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1});

                        \draw[scale=0.5, domain=0.5:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+2});
                        \draw[scale=0.5, domain=1.5:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1.5});
                        \draw[scale=0.5, domain=2.5:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1});

                        \draw[scale=0.5, domain=-3:-0.1, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+0.5});
                        \draw[scale=0.5, domain=-3:-0.1, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x});
                        \draw[scale=0.5, domain=-3:-0.1, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-0.5});
                        \draw[scale=0.5, domain=-3:-0.1, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-1});
                        \draw[scale=0.5, domain=-3:-0.1, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-1.5});
                        \draw[scale=0.5, domain=-3:0, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-2});
                        \draw[scale=0.5, domain=-3:1, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-2.5});
                        \draw[scale=0.5, domain=-3:2, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-3});
                        \draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-3.5});
                        \draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-4});

                        \node at (6.95,1.25) {$b$};
                        \node at (6.78,0) {$h$};
                        \node at (3.2,-1.2) {$\ell = 1$};
                    \end{tikzpicture}
                    \captionof{figure}{Three-Dimensional Beam}
                \end{minipage}
                \hfill
                \begin{minipage}[b]{0.45\linewidth}
                    \centering
                    \begin{tikzpicture}[scale=0.6]
                        \draw[line width = 0.3mm] (0,1) -- (7,1);~
                        \draw[line width = 0.3mm] (0,-1) -- (7,-1);
                        \draw[line width = 0.3mm] (7,-1) -- (7,1);
                        \draw[line width = 0.3mm] (0,-1) -- (0,1);

                        \draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+4});
                        \draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+3.5});
                        \draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+3});
                        \draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+2.5});
                        \draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+2});

                        \draw[scale=0.5, domain=-3:0, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1.5});
                        \draw[scale=0.5, domain=-3:0, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1});

                        \draw[scale=0.5, domain=1:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1.5});
                        \draw[scale=0.5, domain=2:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+1});

                        \draw[scale=0.5, domain=-3:0, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x+0.5});
                        \draw[scale=0.5, domain=-3:0, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x});
                        \draw[scale=0.5, domain=-3:0, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-0.5});
                        \draw[scale=0.5, domain=-3:0, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-1});
                        \draw[scale=0.5, domain=-3:0, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-1.5});
                        \draw[scale=0.5, domain=-3:0, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-2});
                        \draw[scale=0.5, domain=-3:1, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-2.5});
                        \draw[scale=0.5, domain=-3:2, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-3});
                        \draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-3.5});
                        \draw[scale=0.5, domain=-3:3, smooth, variable=\x,densely dotted] plot ({\x}, {0.5*\x-4});

                        \node at (7.2,0) {$h$};
                        \node at (3.2,-1.2) {$\ell = 1$};
                    \end{tikzpicture}
                    \captionof{figure}{Two-dimensional Beam}
                \end{minipage}
            \end{center}
        \end{frame}

        \begin{frame}{Three-Dimensional model vs Two-Dimensional model}
            \centering
            \begin{minipage}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{3D12.png}
                \\ 3D Beam Type - $\lambda_{12} = 2.3293$
                \label{fig:minipage2}
            \end{minipage}
            \hfill
            \begin{minipage}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{2D3.png}
                \\ 2D Beam Type - $\lambda_3 = 2.3273$
                \label{fig:minipage1}
            \end{minipage}

            \vspace{1em} % spacing between the rows

            \begin{minipage}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{3D22.png}
                \\ 3D Beam Type - $\lambda_{24} = 21.929$
                \label{fig:minipage4}
            \end{minipage}
            \hfill
            \begin{minipage}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{2D6.png}
                \\ 2D Beam Type - $\lambda_6 = 21.911$
                \label{fig:minipage3}
            \end{minipage}

            Mode shapes of the displacement \( w \) with \( h=1/20 \).

        \end{frame}

        \begin{frame}{Three-Dimensional model}
            \centering
            \begin{minipage}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{3D33.png}
                \\ 3D Non-Beam Type - $\lambda_{33} = 69.374$
                \label{fig:minipage6}
            \end{minipage}
            \hfill
            \begin{minipage}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{2D8.png}
                \\ 2D Non-Beam Type - $\lambda_8 = 69.344$
                \label{fig:minipage5}
            \end{minipage}

            \vspace{1em} % spacing between the rows

            \begin{minipage}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{3DNonBeam10.png}
                \\ Non-2D Type - $\lambda_{10}$
                \label{fig:minipage8}
            \end{minipage}
            \hfill
            \begin{minipage}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{3DNonBeam11.png}
                \\ Non-2D Type - $\lambda_{11}$
                \label{fig:minipage7}
            \end{minipage}

            Mode shapes of the displacement \( w \) with \( \alpha = 4800 \).
        \end{frame}

        \begin{frame}
            \begin{table}[htbp]
                \scalebox{.8}{
                \makebox[\textwidth]{
                    \begin{tabular}{|cc|cc|cc|cc||cc|}
                        \hline
                        \multicolumn{10}{|c|}{Eigenvalues for $b \leq h$} \\
                        \hline
                        \hline
                        i     & {b = h} & i     & {b = 1/2 h} & i     & {b = 1/4 h} & i     & {b = 1/8 h} & j     & 2D \\
                        \hline
                        {2} & 0.12307 & {2} & 0.12234 & {2} & 0.12198 & {3} & 0.12178 & 1     & 0.12151 \\
                        {3} & 3.5773 & {5} & 3.5630 & {6} & 3.5558 & {8} & 3.5519 & 2     & 3.5460 \\
                        \rowcolor{lightgray}{5} & 7.7799 & {6} & 7.7596 & {8} & 7.7471 & {11} & 7.7401 & 3     & 7.7311 \\
                        {8} & 20.334 & {9} & 20.283 & {11} & 20.26 & {14} & 20.247 & 4     & 20.225 \\
                        {10} & 56.247 & {12} & 56.173 & {15} & 56.156 & {22} & 56.142 & 5     & 56.109 \\
                        \rowcolor{lightgray}{11} & 69.197 & {14} & 69.319 & {17} & 69.281 & {24} & 69.238 & 6     & 69.164 \\
                        {14} & 114.03 & {16} & 114.01 & {20} & 114.05 & {29} & 114.06 & 7     & 114.03 \\
                        \rowcolor{lightgray}{17} & 187.01 & {19} & 189.14 & {25} & 189.37 & {36} & 189.34 & 8     & 189.17 \\
                        {18} & 192.21 & {20} & 192.41 & {26} & 192.58 & {37} & 192.63 & 9     & 192.61 \\
                        {21} & 284.76 & {23} & 285.43 & {31} & 285.74 & {42} & 285.84 & 10    & 285.85 \\
                        {23} & 327.57 & {26} & 328.24 & {35} & 328.37 & {46} & 328.40 & 11    & 328.40 \\
                        \rowcolor{lightgray}{25} & 347.77 & {28} & 356.44 & {36} & 357.30 & {50} & 357.33 & 12    & 357.08 \\
                        {27} & 393.69 & {30} & 396.84 & {38} & 397.28 & {53} & 397.37 & 13    & 397.33 \\
                        {30} & 434.46 & {34} & 441.05 & {41} & 441.81 & {57} & 441.99 & 14    & 442.00 \\
                        \rowcolor{lightgray}{31} & 523.65 & {36} & 534.04 & {43} & 534.17 & {63} & 534.03 & 15    & 533.71 \\
                        {34} & 550.51 & {37} & 537.82 & {44} & 538.86 & {64} & 539.06 & 16    & 538.97 \\
                        \rowcolor{lightgray}{37} & 590.86 & {41} & 587.43 & {48} & 594.17 & {65} & 595.58 & 17    & 596.06 \\
                        {39} & 590.86 & {42} & 600.52 & {49} & 602.25 & {67} & 602.69 & 18    & 602.77 \\
                        \rowcolor{lightgray}{42} & 646.21 & {44} & 657.22 & {50} & 658.04 & {71} & 658.06 & 19    & 657.87 \\
                        {44} & 711.07 & {46} & 714.62 & {53} & 717.10 & {73} & 717.51 & 20    & 717.37 \\
                        \hline
                        \hline
                        \multicolumn{2}{|c|}{Max RE: \  2.6069\%} &\multicolumn{2}{c|}{Max RE: \ 1.4469\%}  & \multicolumn{2}{c|}{Max RE: \  0.38192\%}  & \multicolumn{2}{c||}{Max RE: \ 0.22301\%}& \multicolumn{2}{c|}{-} \\
                        \hline
                    \end{tabular}%
                    \label{tab:2v3_1}%
                }}
            \end{table}%
        \end{frame}

        \begin{frame}
            \begin{table}[htbp]
                \centering
                \begin{tabular}{|c|cccc|}
                    \hline
                    \multicolumn{5}{|c|}{Maximum relative error for beam type and non-beam type eigenvalues for $h = 1/5$.} \\
                    \hline
                    \hline
                    & {b = h} & {b = 1/2h} & {b = 1/4h} & {b = 1/8h} \\
                    \hline
                    Beam Type & 2.1420 \% & 0.6804 \% & 0.38192 \% & 0.22301 \% \\
                    Non-Beam Type & 2.6069 \% & 1.4469 \% & 0.31546 \% & 0.11680 \% \\
                    \hline
                \end{tabular}%
                \label{tab:2Dv3D_1_breakup}%
            \end{table}%
        \end{frame}

        \begin{frame}
            \begin{table}[ht]
                \scalebox{.8}{
                \makebox[\textwidth]{
                \begin{tabular}{|cc|cc|cc|cc||cc|}
                    \hline
                    \multicolumn{10}{|c|}{Comparsion of Eigenvalues with $h = 1/20$, with decreasing $b$ and $b < h$.} \\
                    \hline
                    \hline
                    i     & {b = h} & i     & {b = 1/2 h} & i     & {b = 1/4 h} & i     & {b = 1/8 h} & j     & 2D \\
                    \hline
                    {2} & 0.008043 & {2} & 0.008029 & {2} & 0.008023 & {3} & 0.00802 & {1} & {0.008013} \\
                    {3} & 0.3087 & {4} & 0.30816 & {5} & 0.30794 & {7} & 0.30785 & {2} & {0.30757} \\
                    {5} & 2.3357 & {8} & 2.3316 & {9} & 2.3300  & {12} & 2.3293 & {3} & {2.3273} \\
                    \rowcolor{lightgray}{8} & 7.7217 & {10} & 7.7156 & {13} & 7.7124 & {16} & 7.7111 & {4} & {7.7077} \\
                    {10} & 8.5387 & {11} & 8.5238 & {14} & 8.5182 & {18} & 8.516 & {5} & {8.5086} \\
                    {11} & 21.986 & {14} & 21.948 & {18} & 21.934 & {24} & 21.929 & {6} & {21.911} \\
                    {14} & 45.863 & {18} & 45.781 & {21} & 45.756 & {30} & 45.746 & {7} & {45.712} \\
                    \rowcolor{lightgray}{17} & 69.444 & {19} & 69.408 & {25} & 69.385 & {33} & 69.374 & {8} & {69.344} \\
                    {18} & 83.149 & {22} & 82.999 & {27} & 82.960 & {36} & 82.944 & {9} & {82.887} \\
                    {21} & 136.44 & {25} & 136.19 & {31} & 136.14 & {42} & 136.12 & {10} & {136.03} \\
                    \rowcolor{lightgray}{23} & 192.62 & {27} & 192.62 & {35} & 192.58 & {47} & 192.56 & {11} & {192.48} \\
                    {25} & 207.87 & {29} & 207.5 & {36} & 207.43 & {48} & 207.41 & {12} & {207.29} \\
                    {27} & 299.14 & {33} & 298.63 & {41} & 298.56 & {55} & 298.53 & {13} & {298.38} \\
                    \rowcolor{lightgray}{30} & 376.68 & {35} & 377.01 & {44} & 377.01 & {58} & 376.98 & {14} & {376.83} \\
                    {31} & 411.58 & {37} & 410.89 & {46} & 410.83 & {61} & 410.81 & {15} & {410.63} \\
                    {34} & 546.15 & {40} & 545.27 & {50} & 545.24 & {66} & 545.23 & {16} & {545.03} \\
                    \rowcolor{lightgray}{37} & 620.77 & {42} & 622.02 & {53} & 622.19 & {69} & 622.18 & {17} & {621.95} \\
                    {39} & 703.55 & {44} & 702.49 & {54} & 702.29 & {71} & 702.53 & {18} & {702.31} \\
                    {42} & 884.27 & {47} & 883.02 & {59} & 883.14 & {82} & 883.19 & {19} & {882.96} \\
                    \rowcolor{lightgray}{44} & 923.68 & {49} & 926.88 & {60} & 927.43 & {86} & 927.50 & {20} & {927.18} \\
                    \hline
                    \hline
                    \multicolumn{2}{|c|}{Max RE: \  0.37701\%} &\multicolumn{2}{c|}{Max RE: \ 0.19893\%}  & \multicolumn{2}{c|}{Max RE: \  0.12393\%}  & \multicolumn{2}{c||}{Max RE: \ 0.092843\%}& \multicolumn{2}{c|}{-} \\
                    \hline
                \end{tabular}%
                \label{tab:2v3_2}%
            }}
            \end{table}%
        \end{frame}

        \begin{frame}
            \begin{table}[htbp]
                \centering
                \begin{tabular}{|c|cccc|}
                    \hline
                    \multicolumn{5}{|c|}{Maximum relative error for beam type and non-beam type eigenvalues for $h = 1/20$} \\
                    \hline
                    \hline
                    & {b = h} & {b = 1/2h} & {b = 1/4h} & {b = 1/8h} \\
                    \hline
                    Beam Type & 0.37701 \% & 0.19893 \% & 0.12393 \% & 0.092843 \% \\
                    Non-Beam Type & 0.37682 \% & 0.10218 \% & 0.061213 \% & 0.043601 \% \\
                    \hline
                \end{tabular}%
                \label{tab:2v3_2_split}%
            \end{table}%
        \end{frame}

        \begin{frame}
            \begin{table}[htbp]
                \scalebox{.8}{
                \makebox[\textwidth]{
                    \begin{tabular}{|cc|cc|cc||cc|}
                        \hline
                        \multicolumn{8}{|c|}{Comparsion of Eigenvalues with $h = 1/20$, with increasing $b$ and $b > h$} \\
                        \hline
                        \hline
                        i     & {b = 2h} & i     & {b = 4h} & i     & {b = 8h} & j     & 2D \\
                        \hline
                        {2} & 0.008076 & {1} & 0.008162 & {1} & 0.008324 & 1     & 0.008013 \\
                        {3} & 0.30995 & {3} & 0.31298 & {3} & 0.31738 & 2     & 0.30757 \\
                        {5} & 2.3462 & {5} & 2.3737 & {6} & 2.4116 & 3     & 2.3273 \\
                        \rowcolor{lightgray}{8} & 7.7312 & {8} & 7.7471 & {9} & 7.7711 & 4     & 7.7077 \\
                        {10} & 8.5841 & {9} & 8.7082 & {10} & 8.7929 & 5     & 8.5086 \\
                        {11} & 22.124 & {12} & 22.491 & {14} & 23.222 & 6     & 21.911 \\
                        {14} & 46.195 & {14} & 47.003 & {18} & 47.921 & 7     & 45.712 \\
                        \rowcolor{lightgray}{17} & 69.454 & {17} & 69.281 & {22} & 72.307 & 8     & 69.344 \\
                        {18} & 83.822 & {18} & 85.218 & {24} & 80.607 & 9     & 82.887 \\
                        {21} & 137.43 & {21} & 138.58 & {32} & 140.97 & 10    & 136.03 \\
                        \hline
                        \hline
                        \multicolumn{2}{|c|}{Max RE: \  1.1289\%} &\multicolumn{2}{c|}{Max RE: \ 2.8261\%}  & \multicolumn{2}{c|}{Max RE: \  5.9821\%} & \multicolumn{2}{c|}{-} \\
                        \hline
                    \end{tabular}%
                    \label{tab:b>h_20}%
                }}
            \end{table}
        \end{frame}

        \begin{frame}
            \begin{table}[htbp]
                \centering
                \begin{tabular}{|c|ccc|}
                    \hline
                    \multicolumn{4}{|c|}{Maximum relative error for beam type and non-beam type eigenvalues for $h = 1/20$} \\
                    \hline
                    \hline
                    & {b = 2h} & {b = 4h} & {b = 8h} \\
                    \hline
                    Beam Type & 1.1289 \% & 2.8261 \% & 5.9821 \% \\
                    Non-Beam Type & 0.30521 \% & 0.51096 \% & 4.2734 \% \\
                    \hline
                \end{tabular}%
                \label{tab:b>h-split_20}%
            \end{table}%
        \end{frame}

\section{Conclusion}
    \begin{frame}{Conclusion}
    \end{frame}

    \begin{frame}{References}
        \printbibliography[heading=bibintoc, title={References}]
    \end{frame}

\end{document}